### 2022-01-13 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_metaweb_perspectives@0688247d1354ca4f0fd4214e851dd5f46c103a9e 🚀

%!TEX TS-program = xelatex
\documentclass[11pt]{article}

\usepackage[english]{babel}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{stix2}
\usepackage[scaled]{helvet}
\usepackage[scaled]{inconsolata}

\usepackage{lastpage}

\usepackage{setspace}

\usepackage{ccicons}

\usepackage[hang,flushmargin]{footmisc}

\usepackage{geometry}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}

\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\makeatletter
\newcounter{tableno}
\newenvironment{tablenos:no-prefix-table-caption}{
  \caption@ifcompatibility{}{
    \let\oldthetable\thetable
    \let\oldtheHtable\theHtable
    \renewcommand{\thetable}{tableno:\thetableno}
    \renewcommand{\theHtable}{tableno:\thetableno}
    \stepcounter{tableno}
    \captionsetup{labelformat=empty}
  }
}{
  \caption@ifcompatibility{}{
    \captionsetup{labelformat=default}
    \let\thetable\oldthetable
    \let\theHtable\oldtheHtable
    \addtocounter{table}{-1}
  }
}
\makeatother

\usepackage{array}
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\let\PBS=\PreserveBackslash

\usepackage[breaklinks=true]{hyperref}
\hypersetup{colorlinks,%
citecolor=blue,%
filecolor=blue,%
linkcolor=blue,%
urlcolor=blue}
\usepackage{url}

\usepackage{caption}
\setcounter{secnumdepth}{0}
\usepackage{cleveref}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}

\usepackage{longtable}
\usepackage{booktabs}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for \widthof, \maxof
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\maxof{\widthof{#1}}{\csllabelwidth}}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\geometry{verbose,letterpaper,tmargin=2.2cm,bmargin=2.2cm,lmargin=2.2cm,rmargin=2.2cm}

\usepackage{lineno}
\usepackage[nolists,noheads]{endfloat}

\pagestyle{plain}

\tolerance=1
\emergencystretch=\maxdimen
\hyphenpenalty=10000
\hbadness=10000

\doublespacing

\fancypagestyle{normal}
{
  \fancyhf{}
  \fancyfoot[R]{\footnotesize\sffamily\thepage\ of \pageref*{LastPage}}
}
\begin{document}
\raggedright
\thispagestyle{empty}
{\Large\bfseries\sffamily Predicting metawebs: transfer of graph
embeddings can help alleviate spatial data deficiencies}
\vskip 5em

%
\href{https://orcid.org/0000-0001-6067-1349}{Tanya\,Strydom}%
%
\,\textsuperscript{1,2,‡}\quad %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{1,2,‡}

\textsuperscript{1}\,Département de Sciences Biologiques, Université de
Montréal, Montréal, Canada\quad \textsuperscript{2}\,Quebec Centre for
Biodiversity Science, Montréal, Canada

\textsuperscript{‡}\,These authors contributed equally to the work\\

\textbf{Correspondance to:}\\
Timothée Poisot --- \texttt{timothee.poisot@umontreal.ca}\\

\vfill
This work is released by its authors under a CC-BY 4.0 license\hfill\ccby\\
Last revision: \emph{\today}

\clearpage
\thispagestyle{empty}

\vfill

\begin{enumerate}
    \item Metawebs, i.e.~networks of potential interactions within a
species pool, are a powerful abstraction to understand how large-scales
species interaction networks are structured.%
    \item Because metawebs are typically expressed at large spatial and
taxonomic scales, assembling them is a tedious and costly process;
predictive methods can help circumvent the limitations in data
deficiencies, by providing `draft' metawebs.%
    \item One way to improve the predictive ability is to maximize the
information used for prediction, by using graph embeddings rather than
the list of species interactions. Graph embedding is an emerging field
in machine learning that holds great potential for ecoloigcal problems.%
    \item In this perspective, we outline how the challenges associated
with infering metawebs line-up with the advantages of graph embeddings;
furthermore, because metawebs are inherently spatial objects, we discuss
how the choice of the species pool has consequences on the reconstructed
network, but also embeds hypotheses about which human-made boundaries
are ecologically meaningful.%
\end{enumerate}

\vfill

\clearpage
\linenumbers
\pagestyle{normal}

Being able to infer \emph{potential} interactions could be the catalyst
for significant breakthroughs in our ability to start thinking about
species interaction networks over large spatial scales (Hortal et al.,
2015). In a recent overview of the field of ecological network
prediction, Strydom, Catchen, et al. (2021) identified two challenges of
interest to the prediction of interactions at large scales. First, there
is a relative scarcity of relevant data in most places globally --
paradoxically, this restricts our ability to infer interactions to
locations where inference is perhaps the least required; second,
accurate predictions often demand accurate predictors, and the lack of
methods that can leverage small amount of data is a serious impediment
to our global predictive ability. In most places, our most reliable
biodiversity knowledge is that of a species pool: through the analysis
of databases like GBIF or IUCN, it is possible to establish a list of
species in a region of interest; but establishing the interactions
between these species is difficult.

Following the definition of Dunne (2006), a metaweb is the ecological
network analogue to the species pool; specifically, it iventories
\emph{potential} interactions between species from a spatially delimited
area (and so captures the \(\gamma\) diversity of interactions). The
metaweb is not a prediction of the network at a specific point within
the spatial area it covers: it will have a different structure, notably
by having a larger connectance (see \emph{e.g.} Wood et al., 2015), from
any of these local networks. These local networks (capturing the
\(\alpha\) diversity of interactions) are a subset of the metaweb's
species and interactions, and have been called ``metaweb realizations''
(Poisot et al., 2015). Differences between local networks and their
metawebs are due to chance, species abundance and co-occurrence, local
environmental conditions, and local distribution of functional traits,
among others.

Because the metaweb represents the joint effect of functional,
phylogenetic, and macroecological processes (Morales-Castilla et al.,
2015), it holds valuable ecological information. Specifically, it is the
``upper bounds'' on what the composition of the local networks can be
(see \emph{e.g.} McLeod et al., 2021). These local networks may be
reconstructed given appropriate knowledge of local species composition,
providing information on structure of food webs at finer spatial scales.
This has been done for example for tree-galler-parasitoid systems
(Gravel et al., 2018), fish trophic interactions (Albouy et al., 2019),
tetrapod trophic interactions (O'Connor et al., 2020), and crop-pest
networks (Grünig et al., 2020). In this contribution, we highlight the
\emph{probabilistic} nature of metawebs, discuss how a family of machine
learning tools (graph embeddings and transfer learning) can be used to
overcome data limitations to metaweb inference, and highlight how the
use of metawebs introduces important questions for the field of network
ecology.

\hypertarget{the-metaweb-is-an-inherently-probabilistic-object}{%
\section{The metaweb is an inherently probabilistic
object}\label{the-metaweb-is-an-inherently-probabilistic-object}}

Dallas et al. (2017) suggested that most links in ecological networks
are cryptic, \emph{i.e.} uncommon or hard to observe. This argument
echoes Jordano (2016): sampling ecological interactions is difficult
because it requires first the joint observation of two species, and then
the observation of their interaction. In addition, it is generally
expected that weak or rare links would be more common in networks
(Csermely, 2004), compared to strong, persistent links; this is notably
the case in food chains, wherein many weaker links are key to the
stability of a system (Neutel et al., 2002). In the light of these
observations, we expect to see an over-representation of low-probability
interactions under a model that accurately predicts interaction
probabilities. Yet the original metaweb definition, and indeed most past
uses of metawebs, was based on the presence/absence of interactions.
Moving towards \emph{probabilistic} metawebs, by represent interactions
as Bernoulli events (see \emph{e.g.} Poisot et al., 2016), offers the
opportunity to weigh these rare interactions appropriately. The inherent
plasticity of interactions is important to capture: there have been
documented instances of food webs undergoing rapid collapse/recovery
cycles over short periods of time (\emph{e.g.} Pedersen et al., 2017).
These considerations emphasize why metaweb predictions should focus on
quantitative (preferentially probabilistic) predictions; this should
constrain the suite of appropriate models.

Yet it is important to recall that a metaweb is intended as a catalogue
of all potential interactions, which is then filtered (Morales-Castilla
et al., 2015). In a sense, that most ecological interactions are elusive
can call for a slightly different approach to sampling: once the common
interactions are documented, the effort required in documenting each
rare interaction will increase exponentially, and will do so for each
undocumented interaction. Recent proposals suggest that machine learning
algorithms, in these situations, can act as data generators (Hoffmann et
al., 2019): high quality observational data can generate the core rules
underpinning the network structure, and be supplemented with synthetic
data coming from predictive models, increasing the volume of information
available for inference. Indeed, Strydom, Catchen, et al. (2021)
suggested that knowing the metaweb may render the prediction of local
networks easier, because it fixes an ``upper bound'' on which
interactions can exist. A probabilistic metaweb represents an
aggregation of informative priors on the interactions, an elusive
information with the potential to boost our predictive ability
(\textbf{Bartomeus2016ComFra?}).

\begin{figure}
\hypertarget{fig:embedding}{%
\centering
\includegraphics{figures/conceptual_embedding.jpg}
\caption{Overview of the embedding process. A network (\emph{A}),
possibbly represented as its adjacency matrix (\emph{B}), is converted
into a lower-dimensional object (\emph{C}) where nodes, subgraphs, or
edges have specific values (see tbl.~\ref{tbl:methods}). For the
purposes of prediction, this low-dimensional object encodes feature
vectors for \emph{e.g.} the nodes. Embedding also allows to visualize
the structure in the data differently (\emph{D}), much like with a
principal component analysis.}\label{fig:embedding}
}
\end{figure}

\hypertarget{graph-embedding-offers-promises-for-the-inference-of-potential-interactions}{%
\section{Graph embedding offers promises for the inference of potential
interactions}\label{graph-embedding-offers-promises-for-the-inference-of-potential-interactions}}

Graph embedding fig.~\ref{fig:embedding} is a varied family of machine
learning techniques aiming to transform nodes and edges into a vector
space (Arsov \& Mirceva, 2019), usually of a lower dimension, whilst
maximally retaining key properties of the graph (Yan et al., 2005).
Ecological networks are an interesting candidate for the widespread
application of embeddings, as they tend to posess a shared sstructural
backbone (Mora et al., 2018), which hints at structural invariants that
can be revealed a lower dimensions. Indeed, previous work by Eklöf et
al. (2013) suggests that food webs are inherently low-dimensional
objects, and can be adequately represented with less than ten
dimensions. Simulation results by Botella et al. (2022) suggest that
there is no best method to identify architectural similarities between
networks, and that multiple approaches need to be tested and compared to
the network descriptor of interest. This matches with previous, more
general results on graph embedding, which suggest that embedding
algorithm choice matters for the results (Goyal \& Ferrara, 2018).

\hypertarget{tbl:methods}{}
\begin{longtable}[]{@{}llrr@{}}
\caption{\label{tbl:methods}Overview of some common graph embedding
approaches, by time of publication, alongside examples of their use in
the prediction of species interactions. Surprisingly, these methods have
not yet been used routinely to predict species interactions; most of the
examples we identified were either statistical associations, or
analogues to joint species distribution models. \(^a\): statistical
interactions; \(^b\): joint-SDM-like approach.}\tabularnewline
\toprule
\begin{minipage}[b]{0.11\columnwidth}\raggedright
Method\strut
\end{minipage} & \begin{minipage}[b]{0.32\columnwidth}\raggedright
Embedding approach\strut
\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\raggedleft
Reference\strut
\end{minipage} & \begin{minipage}[b]{0.28\columnwidth}\raggedleft
Species interactions application\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.11\columnwidth}\raggedright
Method\strut
\end{minipage} & \begin{minipage}[b]{0.32\columnwidth}\raggedright
Embedding approach\strut
\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\raggedleft
Reference\strut
\end{minipage} & \begin{minipage}[b]{0.28\columnwidth}\raggedleft
Species interactions application\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.11\columnwidth}\raggedright
RDPG\strut
\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedright
graphs through SVD\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedleft
Young \& Scheinerman (2007)\strut
\end{minipage} & \begin{minipage}[t]{0.28\columnwidth}\raggedleft
Poisot et al. (2021)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright
tSNE\strut
\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedright
nodes through statistical divergence\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedleft
Hinton \& Roweis (2002)\strut
\end{minipage} & \begin{minipage}[t]{0.28\columnwidth}\raggedleft
Cieslak et al. (2020) \(^a\)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright
DeepWalk\strut
\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedright
graph walk\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedleft
Perozzi et al. (2014)\strut
\end{minipage} & \begin{minipage}[t]{0.28\columnwidth}\raggedleft
Wardeh et al. (2021)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright
FastEmbed\strut
\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedright
graph through PCA/SVD analogue\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedleft
Ramasamy \& Madhow (2015)\strut
\end{minipage} & \begin{minipage}[t]{0.28\columnwidth}\raggedleft
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright
LINE\strut
\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedright
nodes through statistical divergence\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedleft
Tang et al. (2015)\strut
\end{minipage} & \begin{minipage}[t]{0.28\columnwidth}\raggedleft
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright
SDNE\strut
\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedright
nodes through auto-encoding\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedleft
D. Wang et al. (2016)\strut
\end{minipage} & \begin{minipage}[t]{0.28\columnwidth}\raggedleft
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright
node2vec\strut
\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedright
node embedding\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedleft
Grover \& Leskovec (2016)\strut
\end{minipage} & \begin{minipage}[t]{0.28\columnwidth}\raggedleft
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright
graph2vec\strut
\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedright
sub-graph embedding\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedleft
Narayanan et al. (2017)\strut
\end{minipage} & \begin{minipage}[t]{0.28\columnwidth}\raggedleft
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright
DMSE\strut
\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedright
joint node embedding\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedleft
D. Chen et al. (2017)\strut
\end{minipage} & \begin{minipage}[t]{0.28\columnwidth}\raggedleft
D. Chen et al. (2017) \(^b\)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright
HARP\strut
\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedright
nodes through a meta-strategy\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedleft
H. Chen et al. (2017)\strut
\end{minipage} & \begin{minipage}[t]{0.28\columnwidth}\raggedleft
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright
GraphKKE\strut
\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedright
graph embedding\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedleft
Melnyk et al. (2020)\strut
\end{minipage} & \begin{minipage}[t]{0.28\columnwidth}\raggedleft
Melnyk et al. (2020) \(^a\)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright
Joint methods\strut
\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedright
multiple graphs\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedleft
S. Wang et al. (2021)\strut
\end{minipage} & \begin{minipage}[t]{0.28\columnwidth}\raggedleft
\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

But the popularity of graph embedding techniques in machine learning is
rather more intuitive than the search for structural invariants: while
graphs are discrete objects, machine learning techniques tend to handle
continuous data better. Therefore, bringing a sparse graph into a
continuous, dense vector space (Xu, 2020) opens up a broader variety of
predictive algorithms.

\textbf{TK} Transfer + embedding graf

Rather than directly predicting biological rules (see \emph{e.g.}
Pichler et al., 2020 for an overview), which may be confounded by the
sparse nature of graph data, learning embeddings works in the
low-dimensional space that maximizes information about the network
structure.

The embedding of a network, in a sense, embeds its macro-evolutionary
history, especially as RDPG captures ecological signal (Dalla Riva \&
Stouffer, 2016); at this point,

\begin{figure}
\hypertarget{fig:prediction}{%
\centering
\includegraphics{figures/conceptual_prediction.jpg}
\caption{From a low-dimensional feature vector (see
fig.~\ref{fig:embedding}), it is possible to develop predictive
approaches. Nodes in an ecological network are species, for which we can
leverage phylogenetic relatedness (\emph{e.g.} Strydom, Bouskila, et
al., 2021) or functional traits to fill the values of additional species
we would like to project in this space (here, I, J, K, and L) from the
embedding of known species (here, A, B, C, and D). Because embeddings
can be projected back to a graph, this allows to reconstruct a network
with these new species. This approach constitutes an instance of
transfer learning.}\label{fig:prediction}
}
\end{figure}

\hypertarget{the-metaweb-embeds-strong-ecological-hypotheses}{%
\section{The metaweb embeds strong ecological
hypotheses}\label{the-metaweb-embeds-strong-ecological-hypotheses}}

As Herbert (1965) rightfully pointed out, ``{[}y{]}ou can't draw neat
lines around planet-wide problems''; in this regard, any inference of a
metaweb at large scales must contend with several interesting and
interwoven families of problems.

The first is the spatial and taxonomic limit of the metaweb to embed and
transfer. If the initial metaweb is too narrow in scope, notably from a
taxonomic point of view, the chances of finding another area with enough
related species (through phylogenetic relatedness or similarity of
functional traits) to make a reliable inference decreases; this would
likely be indicated by large confidence intervals during estimation of
the values in the low-rank space, but the lack of well documented
metawebs is currently preventing the development of more concrete
guidelines. The question of phylogenetic relatedness and dispersal is
notably true if the metaweb is assembled in an area with mostly endemic
species, and as with every predictive algorithm, there is room for the
application of our best ecological judgement. Conversely, the metaweb
should be reliably filled, which assumes that the \(S^2\) interactions
in a pool of \(S\) species have been examined, either through literature
surveys or expert elicitation.

\textbf{TK} Supp. Mat. 1 provides some guidance as to the type of
sampling effort that should be prioritized. While RDPG was able to
maintain very high predictive power when interactions were missing, the
addition of false positive interactions was immediately detected; this
suggests that it may be appropriate to err on the side of ``too many''
interactions when constructing the initial metaweb to be transferred.

The second series of problems are related to determining which area
should be used to infer the new metaweb in, as this determines the
species pool that must be used.

\textbf{TK} In our application, we focused on the mammals of Canada. The
upside of this approach is that information at the country level is
likely to be required by policy makers and stakeholders for their
biodiversity assessment, as each country tends to set goals at the
national level (Buxton et al., 2021) for which quantitative instruments
are designed (Turak et al., 2017), with specific strategies often
enacted at smaller scales (Ray et al., 2021). And yet, we do not really
have a satisfying answer to the question of ``where does a food web
stop?''; the current most satisfying solutions involve examining the
spatial consistency of network area relationships (Fortin et al., 2021;
see \emph{e.g.} Galiana et al., 2018, 2019, 2021), which is of course
impossible in the absence of enough information about the network
itself. This suggests that an \emph{a posteriori} refinement of the
results may be required, based on a downscaling of the metaweb.

The final family of problems relates less to the availability of data or
quantitative tools, and more to the praxis of spatial ecology. Operating
under the context of national divisions, in large parts of the world,
reflects nothing more than the legacy of settler colonialism. Indeed,
the use of ecological data is not an apolitical act (Nost \& Goldstein,
2021), as data infrastructures tend to be designed to answer questions
within national boundaries, and their use both draws upon and reinforces
territorial statecraft; as per Machen \& Nost (2021), this is
particularly true when the output of ``algorithmic thinking''
(\emph{e.g.} relying on machine learning to generate knowledge) can be
re-used for governance (\emph{e.g.} enacting conservation decisions at
the national scale). We therefore recognize that methods such as we
propose operate under the framework that contributed to the ongoing
biodiversity crisis (Adam, 2014), reinforced environmental injustice
(Choudry, 2013; Domínguez \& Luoma, 2020), and on Turtle Island
especially, should be replaced by Indigenous principles of land
management (Eichhorn et al., 2019; No'kmaq et al., 2021). As we see
AI/ML being increasingly mobilized to generate knowledge that is lacking
for conservation decisions (\emph{e.g.} Lamba et al., 2019; Mosebo
Fernandes et al., 2020), our discussion of these tools need to go beyond
the technical, and into the governance consequences they can have.

\textbf{Acknowledgements:} We acknowledge that this study was conducted
on land within the traditional unceded territory of the Saint Lawrence
Iroquoian, Anishinabewaki, Mohawk, Huron-Wendat, and Omàmiwininiwak
nations. TP, TS, DC, and LP received funding from the Canadian Institue
for Ecology \& Evolution. FB is funded by the Institute for Data
Valorization (IVADO). TS, SB, and TP are funded by a donation from the
Courtois Foundation. CB was awarded a Mitacs Elevate Fellowship no.
IT12391, in partnership with fRI Research, and also acknowledges funding
from Alberta Innovates and the Forest Resources Improvement Association
of Alberta. M-JF acknowledges funding from NSERC Discovery Grant and
NSERC CRC. RR is funded by New Zealand's Biological Heritage Ngā Koiora
Tuku Iho National Science Challenge, administered by New Zealand
Ministry of Business, Innovation, and Employment. BM is funded by the
NSERC Alexander Graham Bell Canada Graduate Scholarship and the FRQNT
master's scholarship. LP acknowledges funding from NSERC Discovery Grant
(NSERC RGPIN-2019-05771). TP acknowledges financial support from NSERC
through the Discovery Grants and Discovery Accelerator Supplement
programs.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Adam2014EleTre}{}%
Adam, R. (2014). \emph{Elephant treaties: The Colonial legacy of the
biodiversity crisis}. UPNE.

\leavevmode\hypertarget{ref-Albouy2019MarFis}{}%
Albouy, C., Archambault, P., Appeltans, W., Araújo, M. B., Beauchesne,
D., Cazelles, K., Cirtwill, A. R., Fortin, M.-J., Galiana, N., Leroux,
S. J., Pellissier, L., Poisot, T., Stouffer, D. B., Wood, S. A., \&
Gravel, D. (2019). The marine fish food web is globally connected.
\emph{Nature Ecology \& Evolution}, \emph{3}(8, 8), 1153--1161.
\url{https://doi.org/10.1038/s41559-019-0950-y}

\leavevmode\hypertarget{ref-Arsov2019NetEmb}{}%
Arsov, N., \& Mirceva, G. (2019, November 26). \emph{Network Embedding:
An Overview}. \url{http://arxiv.org/abs/1911.11726}

\leavevmode\hypertarget{ref-Botella2022AppGra}{}%
Botella, C., Dray, S., Matias, C., Miele, V., \& Thuiller, W. (2022). An
appraisal of graph embeddings for comparing trophic network
architectures. \emph{Methods in Ecology and Evolution}, \emph{13}(1),
203--216. \url{https://doi.org/10.1111/2041-210X.13738}

\leavevmode\hypertarget{ref-Buxton2021KeyInf}{}%
Buxton, R. T., Bennett, J. R., Reid, A. J., Shulman, C., Cooke, S. J.,
Francis, C. M., Nyboer, E. A., Pritchard, G., Binley, A. D., Avery-Gomm,
S., Ban, N. C., Beazley, K. F., Bennett, E., Blight, L. K., Bortolotti,
L. E., Camfield, A. F., Gadallah, F., Jacob, A. L., Naujokaitis-Lewis,
I., \ldots{} Smith, P. A. (2021). Key information needs to move from
knowledge to action for biodiversity conservation in Canada.
\emph{Biological Conservation}, \emph{256}, 108983.
\url{https://doi.org/10.1016/j.biocon.2021.108983}

\leavevmode\hypertarget{ref-Chen2017DeeMul}{}%
Chen, D., Xue, Y., Fink, D., Chen, S., \& Gomes, C. P. (2017).
\emph{Deep Multi-species Embedding}. 3639--3646.
\url{https://www.ijcai.org/proceedings/2017/509}

\leavevmode\hypertarget{ref-Chen2017HarHie}{}%
Chen, H., Perozzi, B., Hu, Y., \& Skiena, S. (2017, November 16).
\emph{HARP: Hierarchical Representation Learning for Networks}.
\url{http://arxiv.org/abs/1706.07845}

\leavevmode\hypertarget{ref-Choudry2013SavBio}{}%
Choudry, A. (2013). Saving biodiversity, for whom and for what?
Conservation NGOs, complicity, colonialism and conquest in an era of
capitalist globalization. In \emph{NGOization: Complicity,
contradictions and prospects} (pp. 24--44). Bloomsbury Publishing.

\leavevmode\hypertarget{ref-Cieslak2020TdiSto}{}%
Cieslak, M. C., Castelfranco, A. M., Roncalli, V., Lenz, P. H., \&
Hartline, D. K. (2020). T-Distributed Stochastic Neighbor Embedding
(t-SNE): A tool for eco-physiological transcriptomic analysis.
\emph{Marine Genomics}, \emph{51}, 100723.
\url{https://doi.org/10.1016/j.margen.2019.100723}

\leavevmode\hypertarget{ref-Csermely2004StrLin}{}%
Csermely, P. (2004). Strong links are important, but weak links
stabilize them. \emph{Trends in Biochemical Sciences}, \emph{29}(7),
331--334. \url{https://doi.org/10.1016/j.tibs.2004.05.004}

\leavevmode\hypertarget{ref-DallaRiva2016ExpEvo}{}%
Dalla Riva, G. V., \& Stouffer, D. B. (2016). Exploring the evolutionary
signature of food webs' backbones using functional traits. \emph{Oikos},
\emph{125}(4), 446--456. \url{https://doi.org/10.1111/oik.02305}

\leavevmode\hypertarget{ref-Dallas2017PreCry}{}%
Dallas, T., Park, A. W., \& Drake, J. M. (2017). Predicting cryptic
links in host-parasite networks. \emph{PLOS Computational Biology},
\emph{13}(5), e1005557.
\url{https://doi.org/10.1371/journal.pcbi.1005557}

\leavevmode\hypertarget{ref-Dominguez2020DecCon}{}%
Domínguez, L., \& Luoma, C. (2020). Decolonising Conservation Policy:
How Colonial Land and Conservation Ideologies Persist and Perpetuate
Indigenous Injustices at the Expense of the Environment. \emph{Land},
\emph{9}(3, 3), 65. \url{https://doi.org/10.3390/land9030065}

\leavevmode\hypertarget{ref-Dunne2006NetStr}{}%
Dunne, J. A. (2006). The Network Structure of Food Webs. In J. A. Dunne
\& M. Pascual (Eds.), \emph{Ecological networks: Linking structure and
dynamics} (pp. 27--86). Oxford University Press.

\leavevmode\hypertarget{ref-Eichhorn2019SteDec}{}%
Eichhorn, M. P., Baker, K., \& Griffiths, M. (2019). Steps towards
decolonising biogeography. \emph{Frontiers of Biogeography},
\emph{12}(1), 1--7. \url{https://doi.org/10.21425/F5FBG44795}

\leavevmode\hypertarget{ref-Eklof2013DimEco}{}%
Eklöf, A., Jacob, U., Kopp, J., Bosch, J., Castro-Urgal, R., Chacoff, N.
P., Dalsgaard, B., de Sassi, C., Galetti, M., Guimarães, P. R.,
Lomáscolo, S. B., Martín González, A. M., Pizo, M. A., Rader, R.,
Rodrigo, A., Tylianakis, J. M., Vázquez, D. P., \& Allesina, S. (2013).
The dimensionality of ecological networks. \emph{Ecology Letters},
\emph{16}(5), 577--583. \url{https://doi.org/10.1111/ele.12081}

\leavevmode\hypertarget{ref-Fortin2021NetEco}{}%
Fortin, M.-J., Dale, M. R. T., \& Brimacombe, C. (2021). Network ecology
in dynamic landscapes. \emph{Proceedings of the Royal Society B:
Biological Sciences}, \emph{288}(1949), rspb.2020.1889, 20201889.
\url{https://doi.org/10.1098/rspb.2020.1889}

\leavevmode\hypertarget{ref-Galiana2021SpaSca}{}%
Galiana, N., Barros, C., Braga, J., Ficetola, G. F., Maiorano, L.,
Thuiller, W., Montoya, J. M., \& Lurgi, M. (2021). The spatial scaling
of food web structure across European biogeographical regions.
\emph{Ecography}, \emph{n/a}(n/a).
\url{https://doi.org/10.1111/ecog.05229}

\leavevmode\hypertarget{ref-Galiana2019GeoVar}{}%
Galiana, N., Hawkins, B. A., \& Montoya, J. M. (2019). The geographical
variation of network structure is scale dependent: Understanding the
biotic specialization of host--parasitoid networks. \emph{Ecography},
\emph{42}(6), 1175--1187. \url{https://doi.org/10.1111/ecog.03684}

\leavevmode\hypertarget{ref-Galiana2018SpaSca}{}%
Galiana, N., Lurgi, M., Claramunt-López, B., Fortin, M.-J., Leroux, S.,
Cazelles, K., Gravel, D., \& Montoya, J. M. (2018). The spatial scaling
of species interaction networks. \emph{Nature Ecology \& Evolution},
\emph{2}(5), 782--790. \url{https://doi.org/10.1038/s41559-018-0517-3}

\leavevmode\hypertarget{ref-Goyal2018GraEmb}{}%
Goyal, P., \& Ferrara, E. (2018). Graph embedding techniques,
applications, and performance: A survey. \emph{Knowledge-Based Systems},
\emph{151}, 78--94. \url{https://doi.org/10.1016/j.knosys.2018.03.022}

\leavevmode\hypertarget{ref-Gravel2018BriElt}{}%
Gravel, D., Baiser, B., Dunne, J. A., Kopelke, J.-P., Martinez, N. D.,
Nyman, T., Poisot, T., Stouffer, D. B., Tylianakis, J. M., Wood, S. A.,
\& Roslin, T. (2018). Bringing Elton and Grinnell together: A
quantitative framework to represent the biogeography of ecological
interaction networks. \emph{Ecography}, \emph{0}(0).
\url{https://doi.org/10.1111/ecog.04006}

\leavevmode\hypertarget{ref-Grover2016NodSca}{}%
Grover, A., \& Leskovec, J. (2016). Node2vec: Scalable Feature Learning
for Networks. \emph{Proceedings of the 22nd ACM SIGKDD International
Conference on Knowledge Discovery and Data Mining}, 855--864.
\url{https://doi.org/10.1145/2939672.2939754}

\leavevmode\hypertarget{ref-Grunig2020CroFor}{}%
Grünig, M., Mazzi, D., Calanca, P., Karger, D. N., \& Pellissier, L.
(2020). Crop and forest pest metawebs shift towards increased linkage
and suitability overlap under climate change. \emph{Communications
Biology}, \emph{3}(1, 1), 1--10.
\url{https://doi.org/10.1038/s42003-020-0962-9}

\leavevmode\hypertarget{ref-Herbert1965Dun}{}%
Herbert, F. (1965). \emph{Dune} (1st ed.). Chilton Book Company.

\leavevmode\hypertarget{ref-Hinton2002StoNei}{}%
Hinton, G., \& Roweis, S. T. (2002). Stochastic neighbor embedding.
\emph{NIPS}, \emph{15}, 833--840.

\leavevmode\hypertarget{ref-Hoffmann2019MacLea}{}%
Hoffmann, J., Bar-Sinai, Y., Lee, L. M., Andrejevic, J., Mishra, S.,
Rubinstein, S. M., \& Rycroft, C. H. (2019). Machine learning in a
data-limited regime: Augmenting experiments with synthetic data uncovers
order in crumpled sheets. \emph{Science Advances}, \emph{5}(4),
eaau6792. \url{https://doi.org/10.1126/sciadv.aau6792}

\leavevmode\hypertarget{ref-Hortal2015SevSho}{}%
Hortal, J., de Bello, F., Diniz-Filho, J. A. F., Lewinsohn, T. M., Lobo,
J. M., \& Ladle, R. J. (2015). Seven Shortfalls that Beset Large-Scale
Knowledge of Biodiversity. \emph{Annual Review of Ecology, Evolution,
and Systematics}, \emph{46}(1), 523--549.
\url{https://doi.org/10.1146/annurev-ecolsys-112414-054400}

\leavevmode\hypertarget{ref-Jordano2016SamNet}{}%
Jordano, P. (2016). Sampling networks of ecological interactions.
\emph{Functional Ecology}, \emph{30}(12), 1883--1893.
\url{https://doi.org/10.1111/1365-2435.12763}

\leavevmode\hypertarget{ref-Lamba2019DeeLea}{}%
Lamba, A., Cassey, P., Segaran, R. R., \& Koh, L. P. (2019). Deep
learning for environmental conservation. \emph{Current Biology},
\emph{29}(19), R977--R982.
\url{https://doi.org/10.1016/j.cub.2019.08.016}

\leavevmode\hypertarget{ref-Machen2021ThiAlg}{}%
Machen, R., \& Nost, E. (2021). Thinking algorithmically: The making of
hegemonic knowledge in climate governance. \emph{Transactions of the
Institute of British Geographers}, \emph{46}(3), 555--569.
\url{https://doi.org/10.1111/tran.12441}

\leavevmode\hypertarget{ref-McLeod2021SamAsy}{}%
McLeod, A., Leroux, S. J., Gravel, D., Chu, C., Cirtwill, A. R., Fortin,
M.-J., Galiana, N., Poisot, T., \& Wood, S. A. (2021). Sampling and
asymptotic network properties of spatial multi-trophic networks.
\emph{Oikos}, \emph{n/a}(n/a). \url{https://doi.org/10.1111/oik.08650}

\leavevmode\hypertarget{ref-Melnyk2020GraGra}{}%
Melnyk, K., Klus, S., Montavon, G., \& Conrad, T. O. F. (2020).
GraphKKE: Graph Kernel Koopman embedding for human microbiome analysis.
\emph{Applied Network Science}, \emph{5}(1), 96.
\url{https://doi.org/10.1007/s41109-020-00339-2}

\leavevmode\hypertarget{ref-Mora2018IdeCom}{}%
Mora, B. B., Gravel, D., Gilarranz, L. J., Poisot, T., \& Stouffer, D.
B. (2018). Identifying a common backbone of interactions underlying food
webs from different ecosystems. \emph{Nature Communications},
\emph{9}(1), 2603. \url{https://doi.org/10.1038/s41467-018-05056-0}

\leavevmode\hypertarget{ref-Morales-Castilla2015InfBio}{}%
Morales-Castilla, I., Matias, M. G., Gravel, D., \& Araújo, M. B.
(2015). Inferring biotic interactions from proxies. \emph{Trends in
Ecology \& Evolution}, \emph{30}(6), 347--356.
\url{https://doi.org/10.1016/j.tree.2015.03.014}

\leavevmode\hypertarget{ref-MoseboFernandes2020MacLea}{}%
Mosebo Fernandes, A. C., Quintero Gonzalez, R., Lenihan-Clarke, M. A.,
Leslie Trotter, E. F., \& Jokar Arsanjani, J. (2020). Machine Learning
for Conservation Planning in a Changing Climate. \emph{Sustainability},
\emph{12}(18, 18), 7657. \url{https://doi.org/10.3390/su12187657}

\leavevmode\hypertarget{ref-Narayanan2017GraLea}{}%
Narayanan, A., Chandramohan, M., Venkatesan, R., Chen, L., Liu, Y., \&
Jaiswal, S. (2017, July 17). \emph{Graph2vec: Learning Distributed
Representations of Graphs}. \url{http://arxiv.org/abs/1707.05005}

\leavevmode\hypertarget{ref-Neutel2002StaRea}{}%
Neutel, A.-M., Heesterbeek, J. A. P., \& de Ruiter, P. C. (2002).
Stability in Real Food Webs: Weak Links in Long Loops. \emph{Science},
\emph{296}(5570), 1120--1123.
\url{https://doi.org/10.1126/science.1068326}

\leavevmode\hypertarget{ref-Nokmaq2021AwaSle}{}%
No'kmaq, M., Marshall, A., Beazley, K. F., Hum, J., joudry, shalan,
Papadopoulos, A., Pictou, S., Rabesca, J., Young, L., \& Zurba, M.
(2021). {``Awakening the sleeping giant''}: Re-Indigenization principles
for transforming biodiversity conservation in Canada and beyond.
\emph{FACETS}, \emph{6}(1), 839--869.

\leavevmode\hypertarget{ref-Nost2021PolEco}{}%
Nost, E., \& Goldstein, J. E. (2021). A political ecology of data.
\emph{Environment and Planning E: Nature and Space}, 25148486211043503.
\url{https://doi.org/10.1177/25148486211043503}

\leavevmode\hypertarget{ref-OConnor2020UnvFoo}{}%
O'Connor, L. M. J., Pollock, L. J., Braga, J., Ficetola, G. F.,
Maiorano, L., Martinez‐Almoyna, C., Montemaggiori, A., Ohlmann, M., \&
Thuiller, W. (2020). Unveiling the food webs of tetrapods across Europe
through the prism of the Eltonian niche. \emph{Journal of Biogeography},
\emph{47}(1), 181--192. \url{https://doi.org/10.1111/jbi.13773}

\leavevmode\hypertarget{ref-Pedersen2017SigCol}{}%
Pedersen, E. J., Thompson, P. L., Ball, R. A., Fortin, M.-J., Gouhier,
T. C., Link, H., Moritz, C., Nenzen, H., Stanley, R. R. E., Taranu, Z.
E., Gonzalez, A., Guichard, F., \& Pepin, P. (2017). Signatures of the
collapse and incipient recovery of an overexploited marine ecosystem.
\emph{Royal Society Open Science}, \emph{4}(7), 170215.
\url{https://doi.org/10.1098/rsos.170215}

\leavevmode\hypertarget{ref-Perozzi2014DeeOnl}{}%
Perozzi, B., Al-Rfou, R., \& Skiena, S. (2014). DeepWalk: Online
learning of social representations. \emph{Proceedings of the 20th ACM
SIGKDD International Conference on Knowledge Discovery and Data Mining},
701--710. \url{https://doi.org/10.1145/2623330.2623732}

\leavevmode\hypertarget{ref-Pichler2020MacLea}{}%
Pichler, M., Boreux, V., Klein, A.-M., Schleuning, M., \& Hartig, F.
(2020). Machine learning algorithms to infer trait-matching and predict
species interactions in ecological networks. \emph{Methods in Ecology
and Evolution}, \emph{11}(2), 281--293.
\url{https://doi.org/10.1111/2041-210X.13329}

\leavevmode\hypertarget{ref-Poisot2016StrPro}{}%
Poisot, T., Cirtwill, A. R., Cazelles, K., Gravel, D., Fortin, M.-J., \&
Stouffer, D. B. (2016). The structure of probabilistic networks.
\emph{Methods in Ecology and Evolution}, \emph{7}(3), 303--312.
\url{https://doi.org/10.1111/2041-210X.12468}

\leavevmode\hypertarget{ref-Poisot2021ImpMam}{}%
Poisot, T., Ouellet, M.-A., Mollentze, N., Farrell, M. J., Becker, D.
J., Albery, G. F., Gibb, R. J., Seifert, S. N., \& Carlson, C. J. (2021,
May 31). \emph{Imputing the mammalian virome with linear filtering and
singular value decomposition}. \url{http://arxiv.org/abs/2105.14973}

\leavevmode\hypertarget{ref-Poisot2015SpeWhy}{}%
Poisot, T., Stouffer, D. B., \& Gravel, D. (2015). Beyond species: Why
ecological interaction networks vary through space and time.
\emph{Oikos}, \emph{124}(3), 243--251.
\url{https://doi.org/10.1111/oik.01719}

\leavevmode\hypertarget{ref-Ramasamy2015ComSpe}{}%
Ramasamy, D., \& Madhow, U. (2015). Compressive spectral embedding:
Sidestepping the SVD. In C. Cortes, N. Lawrence, D. Lee, M. Sugiyama, \&
R. Garnett (Eds.), \emph{Advances in neural information processing
systems} (Vol. 28). Curran Associates, Inc.
\url{https://proceedings.neurips.cc/paper/2015/file/4f6ffe13a5d75b2d6a3923922b3922e5-Paper.pdf}

\leavevmode\hypertarget{ref-Ray2021BioCri}{}%
Ray, J. C., Grimm, J., \& Olive, A. (2021). The biodiversity crisis in
Canada: Failures and challenges of federal and sub-national strategic
and legal frameworks. \emph{FACETS}, \emph{6}, 1044--1068.
\url{https://doi.org/10.1139/facets-2020-0075}

\leavevmode\hypertarget{ref-Strydom2021FooWeb}{}%
Strydom, T., Bouskila, S., Banville, F., Barros, C., Caron, D., Farrell,
M. J., Fortin, M.-J., Hemming, V., Mercier, B., Pollock, L., Runghen,
R., Riva, G. V. D., \& Poisot, T. (2021). \emph{Food web reconstruction
through phylogenetic transfer of low-rank network representation}.
\url{https://doi.org/10.32942/osf.io/y7sdz}

\leavevmode\hypertarget{ref-Strydom2021RoaPre}{}%
Strydom, T., Catchen, M. D., Banville, F., Caron, D., Dansereau, G.,
Desjardins-Proulx, P., Forero-Muñoz, N. R., Higino, G., Mercier, B.,
Gonzalez, A., Gravel, D., Pollock, L., \& Poisot, T. (2021). A roadmap
towards predicting species interaction networks (across space and time).
\emph{Philosophical Transactions of the Royal Society B: Biological
Sciences}, \emph{376}(1837), 20210063.
\url{https://doi.org/10.1098/rstb.2021.0063}

\leavevmode\hypertarget{ref-Tang2015LinLar}{}%
Tang, J., Qu, M., Wang, M., Zhang, M., Yan, J., \& Mei, Q. (2015). LINE:
Large-scale Information Network Embedding. \emph{Proceedings of the 24th
International Conference on World Wide Web}, 1067--1077.
\url{https://doi.org/10.1145/2736277.2741093}

\leavevmode\hypertarget{ref-Turak2017UsiEss}{}%
Turak, E., Brazill-Boast, J., Cooney, T., Drielsma, M., DelaCruz, J.,
Dunkerley, G., Fernandez, M., Ferrier, S., Gill, M., Jones, H., Koen,
T., Leys, J., McGeoch, M., Mihoub, J.-B., Scanes, P., Schmeller, D., \&
Williams, K. (2017). Using the essential biodiversity variables
framework to measure biodiversity change at national scale.
\emph{Biological Conservation}, \emph{213}, 264--271.
\url{https://doi.org/10.1016/j.biocon.2016.08.019}

\leavevmode\hypertarget{ref-Wang2016StrDee}{}%
Wang, D., Cui, P., \& Zhu, W. (2016). Structural Deep Network Embedding.
\emph{Proceedings of the 22nd ACM SIGKDD International Conference on
Knowledge Discovery and Data Mining}, 1225--1234.
\url{https://doi.org/10.1145/2939672.2939753}

\leavevmode\hypertarget{ref-Wang2021JoiEmb}{}%
Wang, S., Arroyo, J., Vogelstein, J. T., \& Priebe, C. E. (2021). Joint
Embedding of Graphs. \emph{IEEE Transactions on Pattern Analysis and
Machine Intelligence}, \emph{43}(4), 1324--1336.
\url{https://doi.org/10.1109/TPAMI.2019.2948619}

\leavevmode\hypertarget{ref-Wardeh2021PreMam}{}%
Wardeh, M., Baylis, M., \& Blagrove, M. S. C. (2021). Predicting
mammalian hosts in which novel coronaviruses can be generated.
\emph{Nature Communications}, \emph{12}(1, 1), 780.
\url{https://doi.org/10.1038/s41467-021-21034-5}

\leavevmode\hypertarget{ref-Wood2015EffSpa}{}%
Wood, S. A., Russell, R., Hanson, D., Williams, R. J., \& Dunne, J. A.
(2015). Effects of spatial scale of sampling on food web structure.
\emph{Ecology and Evolution}, \emph{5}(17), 3769--3782.
\url{https://doi.org/10.1002/ece3.1640}

\leavevmode\hypertarget{ref-Xu2020UndGra}{}%
Xu, M. (2020, December 14). \emph{Understanding graph embedding methods
and their applications}. \url{http://arxiv.org/abs/2012.08019}

\leavevmode\hypertarget{ref-Yan2005GraEmb}{}%
Yan, S., Xu, D., Zhang, B., \& Zhang, H.-J. (2005). Graph embedding: A
general framework for dimensionality reduction. \emph{2005 IEEE Computer
Society Conference on Computer Vision and Pattern Recognition
(CVPR'05)}, \emph{2}, 830--837 vol. 2.
\url{https://doi.org/10.1109/CVPR.2005.170}

\leavevmode\hypertarget{ref-Young2007RanDot}{}%
Young, S. J., \& Scheinerman, E. R. (2007). Random Dot Product Graph
Models for Social Networks. In A. Bonato \& F. R. K. Chung (Eds.),
\emph{Algorithms and Models for the Web-Graph} (pp. 138--149). Springer.
\url{https://doi.org/10.1007/978-3-540-77004-6_11}

\end{CSLReferences}

\end{document}