### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_metaweb_perspectives@1d38a4d0137defc303e636439c8f51544298a89c 🚀

%!TEX TS-program = xelatex
\documentclass[11pt]{article}

\usepackage[english]{babel}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{stix2}
\usepackage[scaled]{helvet}
\usepackage[scaled]{inconsolata}

\usepackage{lastpage}

\usepackage{setspace}

\usepackage{ccicons}

\usepackage[hang,flushmargin]{footmisc}

\usepackage{geometry}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}

\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\makeatletter
\newcounter{tableno}
\newenvironment{tablenos:no-prefix-table-caption}{
  \caption@ifcompatibility{}{
    \let\oldthetable\thetable
    \let\oldtheHtable\theHtable
    \renewcommand{\thetable}{tableno:\thetableno}
    \renewcommand{\theHtable}{tableno:\thetableno}
    \stepcounter{tableno}
    \captionsetup{labelformat=empty}
  }
}{
  \caption@ifcompatibility{}{
    \captionsetup{labelformat=default}
    \let\thetable\oldthetable
    \let\theHtable\oldtheHtable
    \addtocounter{table}{-1}
  }
}
\makeatother

\usepackage{array}
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\let\PBS=\PreserveBackslash

\usepackage[breaklinks=true]{hyperref}
\hypersetup{colorlinks,%
citecolor=blue,%
filecolor=blue,%
linkcolor=blue,%
urlcolor=blue}
\usepackage{url}

\usepackage{caption}
\setcounter{secnumdepth}{0}
\usepackage{cleveref}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}

\usepackage{longtable}
\usepackage{booktabs}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for \widthof, \maxof
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\maxof{\widthof{#1}}{\csllabelwidth}}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\geometry{verbose,letterpaper,tmargin=2.2cm,bmargin=2.2cm,lmargin=2.2cm,rmargin=2.2cm}

\usepackage{lineno}
\usepackage[nolists,noheads]{endfloat}

\DeclareDelayedFloatFlavour*{longtable}{table}

\pagestyle{plain}

\tolerance=1
\emergencystretch=\maxdimen
\hyphenpenalty=10000
\hbadness=10000

\doublespacing

\fancypagestyle{normal}
{
  \fancyhf{}
  \fancyfoot[R]{\footnotesize\sffamily\thepage\ of \pageref*{LastPage}}
}
\begin{document}
\raggedright
\thispagestyle{empty}
{\Large\bfseries\sffamily Graph embedding and transfer learning can help
predict species interaction networks despite data limitations}
\vskip 5em

%
\href{https://orcid.org/0000-0001-6067-1349}{Tanya\,Strydom}%
%
\,\textsuperscript{1,2,‡}\quad %
\href{https://orcid.org/0000-0003-0193-5441}{Salomé\,Bouskila}%
%
\,\textsuperscript{1,‡}\quad %
\href{https://orcid.org/0000-0001-9051-0597}{Francis\,Banville}%
%
\,\textsuperscript{1,3,2}\quad %
\href{https://orcid.org/0000-0003-4036-977X}{Ceres\,Barros}%
%
\,\textsuperscript{4}\quad %
\href{https://orcid.org/0000-0002-2151-6693}{Dominique\,Caron}%
%
\,\textsuperscript{5,2}\quad %
\href{https://orcid.org/0000-0003-0452-6993}{Maxwell J\,Farrell}%
%
\,\textsuperscript{6}\quad %
\href{https://orcid.org/0000-0002-9935-1366}{Marie-Josée\,Fortin}%
%
\,\textsuperscript{6}\quad %
\href{https://orcid.org/0000-0003-3220-6161}{Victoria\,Hemming}%
%
\,\textsuperscript{7}\quad %
\href{https://orcid.org/0000-0002-4104-9463}{Benjamin\,Mercier}%
%
\,\textsuperscript{3,2}\quad %
\href{https://orcid.org/0000-0002-6004-4027}{Laura J.\,Pollock}%
%
\,\textsuperscript{5,2}\quad %
\href{https://orcid.org/0000-0001-5785-8321}{Rogini\,Runghen}%
%
\,\textsuperscript{8}\quad %
\href{https://orcid.org/0000-0002-3454-0633}{Giulio V.\,Dalla Riva}%
%
\,\textsuperscript{9}\quad %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{1,2,‡}

\textsuperscript{1}\,Département de Sciences Biologiques, Université de
Montréal, Montréal, Canada\quad \textsuperscript{2}\,Quebec Centre for
Biodiversity Science, Montréal,
Canada\quad \textsuperscript{3}\,Département de Biologie, Université de
Sherbrooke, Sherbrooke, Canada\quad \textsuperscript{4}\,Department of
Forest Resources Management, University of British Columbia, Vancouver,
B.C., Canada\quad \textsuperscript{5}\,Department of Biology, McGill
University, Montréal, Canada\quad \textsuperscript{6}\,Department of
Ecology \& Evolutionary Biology, University of Toronto, Toronto,
Canada\quad \textsuperscript{7}\,Department of Forest and Conservation
Sciences, University of British Columbia, Vancouver,
Canada\quad \textsuperscript{8}\,Centre for Integrative Ecology, School
of Biological Sciences, University of Canterbury, Canterbury, New
Zealand\quad \textsuperscript{9}\,School of Mathematics and Statistics,
University of Canterbury, Canterbury, New Zealand

\textsuperscript{‡}\,These authors contributed equally to the work\\

\textbf{Correspondance to:}\\
Timothée Poisot --- \texttt{timothee.poisot@umontreal.ca}\\

\vfill
This work is released by its authors under a CC-BY 4.0 license\hfill\ccby\\
Last revision: \emph{\today}

\clearpage
\thispagestyle{empty}

\vfill

\begin{enumerate}
    \item Metawebs, (networks of potential interactions within a species
pool) are a powerful abstraction to understand how large-scale species
interaction networks are structured.%
    \item Because metawebs are typically expressed at large spatial and
taxonomic scales, assembling them is a tedious and costly process;
predictive methods can help circumvent the limitations in data
deficiencies, by providing `draft' metawebs.%
    \item One way to improve our ability to predict metawebs is to
maximize available information by using graph embeddings, instead of the
list of species interactions. Graph embedding is an emerging field in
machine learning that holds great potential for ecological problems.%
    \item In this perspective, we outline how the challenges associated
with inferring metawebs line-up with the advantages of graph embeddings;
as well as discuss how the choice of the species pool has consequences
on the reconstructed network, but also embeds hypotheses about which
human-made boundaries are ecologically meaningful.%
\end{enumerate}

\vfill

\clearpage
\linenumbers
\pagestyle{normal}

Being able to infer \emph{potential} interactions could serve as a
significant breakthrough in our ability to start thinking about species
interaction networks over large spatial scales (Hortal et al., 2015).
Understanding species interactions holds enormous potential to not only
understand and more rapidly learn about species interactions and
metawebs, but also how changes in management of a single species may
impact non-target species. In a recent overview of the field of
ecological network prediction, Strydom, Catchen, et al. (2021)
identified two challenges of interest to the prediction of interactions
at large scales. First, there is a relative scarcity of relevant data in
most places globally -- paradoxically, this restricts our ability to
infer interactions for locations where inference is perhaps the least
required (and leaves us unable to make inference in regions without
interaction data); second, accurate predictors are important for
accurate predictions, and the lack of methods that can leverage a small
amount of \emph{accurate} data is a serious impediment to our predictive
ability. In most places, our most reliable biodiversity knowledge is
that of a species pool (\emph{i.e.} a set of potentially interacting
species in a given area): through the analysis of databases like GBIF or
IUCN, it is possible to construct a list of species in a region of
interest; but inferring the potential interactions between these species
is difficult.

Following the definition of Dunne (2006), a metaweb is the ecological
network analogue to the species pool; specifically, it inventories all
\emph{potential} interactions between species for a spatially delimited
area (and so captures the \(\gamma\) diversity of interactions). The
metaweb is not a prediction of the network at a specific point within
the spatial area it covers: it will have a different structure, notably
by having a larger connectance (see \emph{e.g.} Wood et al., 2015) and
complexity (see \emph{e.g.} Galiana et al., 2022), from any of these
local networks. These local networks (which capture the \(\alpha\)
diversity of interactions) are a subset of the metaweb's species and
realized interactions, and have been called ``metaweb realizations''
(Poisot et al., 2015). Differences between local networks and their
metawebs are due to chance, species abundance and co-occurrence, local
environmental conditions, and local distribution of functional traits,
among others. Yet, recent results by Saravia et al. (2021) strongly
suggest that the local (metaweb) realizations only respond weakly to
local conditions: instead, they reflect constraints inherited by the
structure of their metaweb. This establishes the metaweb structure as
the core goal of predictive network ecology, as it is a required
information to accurately produce downscaled, local predictions.

Because the metaweb represents the joint effect of functional,
phylogenetic, and macroecological processes (Morales-Castilla et al.,
2015), it holds valuable ecological information. Specifically, it
represents the ``upper bounds'' on what the composition of the local
networks, given a local species pool, can be (see \emph{e.g.} McLeod et
al., 2021); this information can help evaluate the ability of ecological
assemblages to withstand the effects of, for example, climate change
(Fricke et al., 2022). These local networks may be reconstructed given
an appropriate knowledge of local species composition and provide
information on the structure of food webs at finer spatial scales. This
has been done for example for tree-galler-parasitoid systems (Gravel et
al., 2018), fish trophic interactions (Albouy et al., 2019), tetrapod
trophic interactions (Braga et al., 2019; O'Connor et al., 2020), and
crop-pest networks (Grünig et al., 2020). In this contribution, we
highlight the power in viewing (and constructing) metawebs as
\emph{probabilistic} objects in the context of rare interactions,
discuss how a family of machine learning tools (graph embeddings and
transfer learning) can be used to overcome data limitations to metaweb
inference, and highlight how the use of metawebs introduces important
questions for the field of network ecology.

\hypertarget{a-metaweb-is-an-inherently-probabilistic-object}{%
\section{A metaweb is an inherently probabilistic
object}\label{a-metaweb-is-an-inherently-probabilistic-object}}

Treating interactions as probabilistic (as opposed to binary) events is
a more nuanced and realistic way to represent them. Dallas et al. (2017)
suggested that most links in ecological networks are cryptic,
\emph{i.e.} uncommon or hard to observe. This argument echoes Jordano
(2016): sampling ecological interactions is difficult because it
requires first the joint observation of two species, and then the
observation of their interaction. In addition, it is generally expected
weak or rare links to be more prevalent in networks than common or
strong links (Csermely, 2004), compared to strong, persistent links;
this is notably the case in food chains, wherein many weaker links are
key to the stability of a system (Neutel et al., 2002). In the light of
these observations, we expect to see an over-representation of
low-probability (rare) interactions under a model that accurately
predicts interaction probabilities. Yet the original metaweb definition,
and indeed most past uses of metawebs, was based on the presence/absence
of interactions. Moving towards \emph{probabilistic} metawebs, by
representing interactions as Bernoulli events (see \emph{e.g.} Poisot et
al., 2016), offers the opportunity to weigh these rare interactions
appropriately. The inherent plasticity of interactions is important to
capture: there have been documented instances of food webs undergoing
rapid collapse/recovery cycles over short periods of time (\emph{e.g.}
Pedersen et al., 2017). These considerations emphasize why metaweb
predictions should focus on quantitative (preferentially probabilistic)
predictions, and this should constrain the suite of appropriate models
used to predict them.

It is important to recall that a metaweb is intended as a catalogue of
all potential interactions, which is then filtered for a given
application (Morales-Castilla et al., 2015). In a sense, that most
ecological interactions are elusive can call for a slightly different
approach to sampling: once the common interactions are documented, the
effort required in documenting each rare interaction will increase
exponentially. Recent proposals suggest that machine learning
algorithms, in these situations, can act as data generators (Hoffmann et
al., 2019): high quality observational data can generate the core rules
underpinning the network structure, and be supplemented with synthetic
data coming from predictive models, increasing the volume of information
available for inference. Indeed, Strydom, Catchen, et al. (2021)
suggested that knowing the metaweb may render the prediction of local
networks easier, because it fixes an ``upper bound'' on which
interactions can exist. In this context, a probabilistic metaweb
represents an aggregation of informative priors on the interactions,
elusive information with the potential to boost our predictive ability
(Bartomeus et al., 2016).

\begin{figure}
\hypertarget{fig:embedding}{%
\centering
\includegraphics{figures/conceptual_2.png}
\caption{Overview of the embedding process. A network (\textbf{A}),
represented here as its adjacency matrix, is converted into a
lower-dimensional object (\textbf{B}) where nodes, subgraphs, or edges
have specific values (see tbl.~\ref{tbl:methods} for an overview of
methods and their use for species interactions). For the purposes of
prediction, this low-dimensional object encodes feature vectors for
\emph{e.g.} the nodes. Embedding also allows to visualize the structure
in the data differently (see fig.~\ref{fig:illustration}), much like
with a principal component analysis. From a low-dimensional feature
vector, it is possible to develop predictive approaches. Nodes in an
ecological network are usually species (\textbf{C}), for which we can
leverage phylogenetic relatedness (\emph{e.g.} Strydom et al., 2022) or
functional traits to fill the values of additional species we would like
to project in this space (here for nodes I, J, K, and L) from the
embedding of known species (here, nodes A, B, C, and D). Because
embeddings can be projected back to a graph, this allows us to
reconstruct a network with these new species (\textbf{D}). This entire
cycle constitutes an instance of transfer learning, where the transfered
information is the representation of graph \textbf{A} through its
embedding.}\label{fig:embedding}
}
\end{figure}

\hypertarget{graph-embedding-offers-promises-for-the-inference-of-potential-interactions}{%
\section{Graph embedding offers promises for the inference of potential
interactions}\label{graph-embedding-offers-promises-for-the-inference-of-potential-interactions}}

Graph (or Network) embedding (fig.~\ref{fig:embedding}) is a family of
machine learning techniques, whose main task is to learn a mapping
function from a discrete graph to a continuous domain (Arsov \& Mirceva,
2019; Chami et al., 2022). Their main goal is to learn a low dimensional
vector representations for the nodes of the graph (embeddings), such
that key properties of the graph (e.g.~local or global structures) are
retained in the embedding space(Yan et al., 2005). Ecological networks
are an interesting candidate for the widespread application of
embeddings, as they tend to possess a shared structural backbone (Bramon
Mora et al., 2018), which hints at structural invariants that can be
revealed at lower dimensions. Indeed, food webs are inherently
low-dimensional objects, and can be adequately represented with less
than ten dimensions (Braga et al., 2019; Eklöf et al., 2013). Simulation
results by Botella et al. (2022) suggest that there is no best method to
identify architectural similarities between networks, and that multiple
approaches need to be tested and compared to the network descriptor of
interest. This matches previous, more general results on graph
embedding, which suggest that the choice of embedding algorithm matters
for the results (Goyal \& Ferrara, 2018). In tbl.~\ref{tbl:methods}, we
present a selection of common graph and node embedding methods,
alongside examples of their use to predict species interactions; most of
these methods rely either on linear algebra, or on pseudo-random walks
on graphs.

One prominent family of approaches we do not discuss in the present
manuscript is Graph Neural Networks (GNN; Zhou et al., 2020). GNN are,
in a sense, a method to embed a graph into a dense subspace, but belong
to the family of deep learning methods, which has its own set of
practices (see \emph{e.g.} Goodfellow et al., 2016). An important issue
with methods based on deep learning is that, because their parameter
space is immense, the sample size of the data fed into them must be
similarly large (typically thousands of instances). This is a
requirement for the model to converge correctly during training, but
this assumption is unlikely to be met given the size of datasets
currently available for metawebs (or single time/location species
interaction networks). This data volume requirement is mostly absent
from the techniques we list below. Furthermore, GNN still have some
challenges related to their shallow structure, and concerns related to
scalability (see Gupta et al., 2021 for a review), which are mostly
absent from the methods listed in tbl.~\ref{tbl:methods}. Assuming that
the uptake of next-generation biomonitoring techniques does indeed
deliver larger datasets on species interactions (Bohan et al., 2017),
there is a potential for GNN to become an applicable
embedding/predictive technique in the coming years.

\hypertarget{tbl:methods}{}
\begin{longtable}[]{@{}lllll@{}}
\caption{\label{tbl:methods}Overview of some common graph embedding
approaches, by type of embedded objects, alongside examples of their use
in the prediction of species interactions. These methods have not yet
been routinely used to predict species interactions; most examples that
we identified were either statistical associations, or analogues to
joint species distribution models. \(^a\): statistical interactions;
\(^b\): joint-SDM-like approach. Note that the row for PCA also applies
to kernel/probabilistic PCA, which are variations on the more general
method of SVD. Note further that tSNE has been included because it is
frequently used to embed graphs, including of species
associations/interactions, despite not being strictly speaking, a graph
embedding technique (see \emph{e.g.} Chami et al., 2022)}\tabularnewline
\toprule
\begin{minipage}[b]{0.09\columnwidth}\raggedright
Method\strut
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedright
Object\strut
\end{minipage} & \begin{minipage}[b]{0.23\columnwidth}\raggedright
Technique\strut
\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedright
Reference\strut
\end{minipage} & \begin{minipage}[b]{0.29\columnwidth}\raggedright
Application\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.09\columnwidth}\raggedright
Method\strut
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedright
Object\strut
\end{minipage} & \begin{minipage}[b]{0.23\columnwidth}\raggedright
Technique\strut
\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedright
Reference\strut
\end{minipage} & \begin{minipage}[b]{0.29\columnwidth}\raggedright
Application\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.09\columnwidth}\raggedright
tSNE\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
nodes\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
statistical divergence\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
Hinton \& Roweis (2002)\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright
Gibb et al. (2021); Cieslak et al. (2020) \(^a\)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright
LINE\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
nodes\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
stochastic gradient descent\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
Tang et al. (2015)\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright
SDNE\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
nodes\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
gradient descent\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
D. Wang et al. (2016)\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright
node2vec\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
nodes\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
stochastic gradient descent\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
Grover \& Leskovec (2016)\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright
HARP\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
nodes\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
meta-strategy\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
H. Chen et al. (2017)\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright
DMSE\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
joint nodes\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
deep neural network\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
D. Chen et al. (2017)\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright
D. Chen et al. (2017) \(^b\)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright
graph2vec\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
sub-graph\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
skipgram network\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
Narayanan et al. (2017)\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright
RDPG\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
graph\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
SVD\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
Young \& Scheinerman (2007)\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright
Poisot et al. (2021); Dalla Riva \& Stouffer (2016)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright
GLEE\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
graph\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
Laplacian eigenmap\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
(\textbf{Torres2020GleGeo?})\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright
DeepWalk\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
graph\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
stochastic gradient descent\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
Perozzi et al. (2014)\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright
Wardeh et al. (2021)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright
GraphKKE\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
graph\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
stochastic differential equation\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
Melnyk et al. (2020)\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright
Melnyk et al. (2020) \(^a\)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright
FastEmbed\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
graph\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
eigen decomposition\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
Ramasamy \& Madhow (2015)\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright
PCA\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
graph\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
eigen decomposition\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
(\textbf{S2013GraEmb?})\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright
Strydom, Catchen, et al. (2021)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright
Joint methods\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
multiple graphs\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
multiple strategies\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
S. Wang et al. (2021)\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

The popularity of graph embedding techniques in machine learning is more
than the search for structural invariants: graphs are discrete objects,
and machine learning techniques tend to handle continuous data better.
Bringing a sparse graph into a continuous, dense vector space (Xu, 2020)
opens up a broader variety of predictive algorithms, notably of the sort
that are able to predict events as probabilities (Murphy, 2022).
Furthermore, the projection of the graph itself is a representation that
can be learned; Runghen et al. (2021), for example, used a neural
network to learn the embedding of a network in which not all
interactions were known, based on the nodes' metadata. This example has
many parallels in ecology (see fig.~\ref{fig:embedding} \textbf{C}), in
which node metadata can be given by phylogeny or functional traits.
Rather than directly predicting biological rules (see \emph{e.g.}
Pichler et al., 2020 for an overview), which may be confounded by the
sparse nature of graph data, learning embeddings works in the
low-dimensional space that maximizes information about the network
structure. This approach is further justified by the observation, for
example, that the macro-evolutionary history of a network is adequately
represented by some graph embeddings (Random dot product graphs (RDPG);
see Dalla Riva \& Stouffer, 2016). In a recent publication, Strydom et
al. (2022) have used an embedding (based on RDPG) to project a metaweb
of trophic interactions between European mammals, and transferred this
information to mammals of Canada, using the phylogenetic distance
between related clades to infer the values in the latent sub-space into
which the European metaweb was projected. By performing the RDPG step on
re-constructed values, this approach yields a probabilistic trophic
metaweb for mammals of Canada based on knowledge of European species,
despite a limited (\(\approx\) 5\%) taxonomic overlap.

Graph embeddings \emph{can} serve as a dimensionality reduction method.
For example, RDPG (Strydom et al., 2022) and t-SVD (truncated Singular
Value Decomposition; Poisot et al., 2021) typically embed networks using
fewer dimensions than the original network (the original network has as
many dimensions as species, and as many informative dimensions as
trophically unique species; Strydom, Dalla Riva, et al., 2021). But this
is not necessarily the case -- indeed, one may perform a PCA (a special
case of SVD) to project the raw data into a subspace that improves the
efficacy of t-SNE (t-distributed stochastic neighbor embedding; Maaten,
2009). There are many dimensionality reductions (Anowar et al., 2021)
that can be applied to an embedded network should the need for
dimensionality reduction (for example for data visualisation) arise. In
brief, many graph embeddings \emph{can} serve as dimensionality
reduction steps, but not all do, neither do all dimensionality reduction
methods provide adequate graph embedding capacities. In the next section
(and fig.~\ref{fig:illustration}), we show how the amount of
dimensionality reduction can affect the quality of the embedding.

\hypertarget{an-illustration-of-metaweb-embedding}{%
\section{An illustration of metaweb
embedding}\label{an-illustration-of-metaweb-embedding}}

In this section, we illustrate the embedding of a collection of
bipartite networks collected by Hadfield et al. (2014), using t-SVD and
RDPG (see Strydom et al., 2022 for the full details). Briefly, an RDPG
decomposes a network into two subspaces (left and right), which are
matrices that when multiplied give an approximation of the original
network. The code to reproduce this example is available as
supplementary material (note, for the sake of comparison, that Strydom,
Catchen, et al., 2021 have an example using embedding through PCA
followed by prediction using a deep neural network on the same dataset).
The resulting (binary) metaweb \(\mathcal{M}\) has 2131 interactions
between 206 parasites and 121 hosts, and its adjacency matrix has full
rank (\emph{i.e.} it represents a space with 121 dimensions). All
analyses were done using Julia (Bezanson et al., 2017) version 1.7.2,
\emph{Makie.jl} (Danisch \& Krumbiegel, 2021), and
\emph{EcologicalNetworks.jl} (Poisot et al., 2019).

\begin{figure}
\hypertarget{fig:illustration}{%
\centering
\includegraphics{figures/illustration.png}
\caption{Illustration of an embedding for an host-parasite metaweb,
using Random Dot Product Graphs. \textbf{A}, decrease in approximation
error as the number of dimensions in the subspaces increases.
\textbf{B}, position of hosts and parasites in the first two dimensions
of their respective subspaces. \textbf{C}, predicted interaction weight
from the RDPG based on the status of the species pair in the metaweb.
\textbf{D}, relationship between the position on the first dimension and
parasite generalism.}\label{fig:illustration}
}
\end{figure}

The embedding of the metaweb holds several pieces of information
(fig.~\ref{fig:illustration}). In panel \textbf{A}, we show that the
\(L_2\) loss (\emph{i.e.} the sum of squared errors) between the
empirical and reconstructed metaweb decreases when the number of
dimensions (rank) of the subspace increases, with an inflection point
around 25 dimensions. As discussed by Runghen et al. (2021), there is
often a trade-off between the number of dimensions to use (more
dimensions are more computationally demanding) and the quality of the
representation. In this instance, accepting \(L_2 = 500\) as an
approximation of the network means that the error for every position in
the metaweb is \(\approx \left(500/(206\times 121)\right)^{1/2}\). In
fig.~\ref{fig:illustration}, panel \textbf{B}, we show the positions of
hosts and parasites on the first two dimensions of the left and right
subspaces. Note that these values largely skew negative, because the
first dimensions capture the coarse structure of the network: most pairs
of species do not interact, and therefore have negative values. In
fig.~\ref{fig:illustration}, panel \textbf{C}, we show the predicted
weight (\emph{i.e.} the result of the multiplication of the RDGP
subspaces at a rank of 25) as a function of whether the interactions are
observed, not-observed, or unknown due to lack of co-occurrence. This
reveals that the observed interactions have higher predicted weights,
although there is some overlap; the usual approach to identify potential
interactions based on this information would be a thresholding analysis,
which is outside the scope of this manuscript (and is done in the papers
cited in this illustration). Note that the values are not bound to the
unit interval, which emphasizes the need for either scaling or clamping
(although thresholding analyses are insensitive to this choice).
Finally, in fig.~\ref{fig:illustration}, panel \textbf{D}, we show that
the embedding, as it captures structural information about the network,
holds ecological information; indeed, the position of the parasite on
the first dimension of the left sub-space is a linear predictor of its
number of hosts.

\hypertarget{the-metaweb-embeds-both-ecological-hypotheses-and-practices}{%
\section{The metaweb embeds both ecological hypotheses and
practices}\label{the-metaweb-embeds-both-ecological-hypotheses-and-practices}}

The goal of metaweb inference is to provide information about the
interactions between species at a large spatial scale. But as Herbert
(1965) rightfully pointed out, ``{[}y{]}ou can't draw neat lines around
planet-wide problems''; any inference of a metaweb at large scales must
contend with several novel, and interwoven, families of problems. In
this section, we list some of the most pressing research priorities
(\emph{i.e.} problems that can be adressed with subsequent data analysis
or simulations), as well as issues related to the application of these
methods at the science-policy interface.

The first open research problem is the taxonomic and spatial limit of
the metaweb to embed and transfer. If the initial metaweb is too narrow
in scope, notably from a taxonomic point of view, the chances of finding
another area with enough related species (through phylogenetic
relatedness or similarity of functional traits) to make a reliable
inference decreases; this would likely be indicated by large confidence
intervals during estimation of the values in the low-rank space, meaning
that the representation of the original graph is difficult to transfer
to the new problem. Alternatively, if the initial metaweb is too large
(taxonomically), then the resulting embeddings would need to represent
interactions between taxonomic groups that are not present in the new
location. This would lead to a much higher variance in the starting
dataset, and to under-dispersion in the target dataset, resulting in the
potential under or over estimation of the strength of new predicted
interactions. The lack of well documented metawebs is currently
preventing the development of more concrete guidelines. The question of
phylogenetic relatedness and distribution is notably relevant if the
metaweb is assembled in an area with mostly endemic species (\emph{e.g.}
a system that has undergone recent radiation or that has remained in
isolation for a long period of time might not have an analogous system
with which to draw knowledge from), and as with every predictive
algorithm, there is room for the application of our best ecological
judgement. Because this problem relates to distribution of species in
the geographic or phylogenetic space, it can certainly be approached
through assessing the performance of embedding transfer in simulated
starting/target species pools.

The second series of problems relate to determining which area should be
used to infer the new metaweb in, as this determines the species pool
that must be used. Metawebs can be constructed by assigning interactions
in a list of species within geographic boundaries. The upside of this
approach is that information at the country level is likely to be
required for biodiversity assessments, as countries set conservation
goals at the national level (Buxton et al., 2021), and as quantitative
instruments are designed to work at these scales (Turak et al., 2017);
specific strategies are often enacted at smaller scales, nested within a
specific country (Ray et al., 2021). But there is no guarantee that
these boundaries are meaningful. In fact, we do not have a satisfying
answer to the question of ``where does a food web stop?''; the most
promising solutions involve examining the spatial consistency of network
area relationships (Fortin et al., 2021; see \emph{e.g.} Galiana et al.,
2018, 2019, 2021), which is impossible in the absence of enough
information about the network itself. This suggests that inferred
metawebs should be further downscaled to allow for \emph{a posteriori}
analyses. The methodology for metaweb downscaling is currently limited,
and it is likely that the sustained effort to characterize the spatial
dependency of food web structure will lead to more prescriptive
guidelines about the need for prediction downscaling.

The final family of problems relates less to ecological methods than to
the praxis of ecological research. Operating under the context of
national divisions, in large parts of the world, reflects nothing more
than the legacy of settler colonialism, which drives a disparity in
available ecological data. Applying any embedding to biased data does
not debias them, but instead embeds these very same biases, propagating
them to the machine learning models using embeddings to make
predictions. Indeed, the use of ecological data is not an apolitical act
(Nost \& Goldstein, 2021), as data infrastructures tend to be designed
to answer questions within national boundaries (therefore placing
contingencies on what is available to be embedded), and their use often
draws upon and reinforces territorial statecraft. As per Machen \& Nost
(2021), this is particularly true when the output of ``algorithmic
thinking'' (\emph{e.g.} relying on machine learning to generate
knowledge) can be re-used for governance (\emph{e.g.} enacting
conservation decisions at the national scale). As information on species
interaction networks structure is increasingly leveraged as a tool to
guide conservation actions (see \emph{e.g.} recent discussions for
food-web based conservation; Eero et al., 2021; Naman et al., 2022;
Stier et al., 2017), the need to appraise and correct biases that are
unwittingly propagated to algorithms when embedded from the original
data is paramount. Predictive approaches deployed at the continental
scale, no matter their intent, originate in the framework that
contributed to the ongoing biodiversity crisis (Adam, 2014) and
reinforced environmental injustice (Choudry, 2013; Domínguez \& Luoma,
2020). Particularly on Turtle Island and other territories that were
traditionally stewarded by Indigenous people, these approaches should be
replaced (or at least guided and framed) by Indigenous principles of
land management (Eichhorn et al., 2019; No'kmaq et al., 2021), as part
of an ``algorithm-in-the-loop'' approach. Human-algorithm interactions
are notoriously difficult and can yield adverse effect (Green \& Chen,
2019; Stevenson \& Doleac, 2021), suggesting the need to systematically
study them for the specific purpose of biodiversity governance, as well
as to improve the algorithmic literacy of decision makers. As we see
artificial intelligence/machine learning being increasingly mobilized to
generate knowledge that is lacking for conservation decisions
(\emph{e.g.} Lamba et al., 2019; Mosebo Fernandes et al., 2020) and
drive policy decisions (Weiskopf et al., 2022), our discussion of these
tools need to go beyond the technical and statistical, and into the
governance consequences they can have.

\textbf{Acknowledgements:} We acknowledge that this study was conducted
on land within the traditional unceded territory of the Saint Lawrence
Iroquoian, Anishinabewaki, Mohawk, Huron-Wendat, and Omàmiwininiwak
nations. TP, TS, DC, and LP received funding from the Canadian Institute
for Ecology \& Evolution. FB is funded by the Institute for Data
Valorization (IVADO). TS, SB, and TP are funded by a donation from the
Courtois Foundation. CB was awarded a Mitacs Elevate Fellowship no.
IT12391, in partnership with fRI Research, and also acknowledges funding
from Alberta Innovates and the Forest Resources Improvement Association
of Alberta. M-JF acknowledges funding from NSERC Discovery Grant and
NSERC CRC. RR is funded by New Zealand's Biological Heritage Ngā Koiora
Tuku Iho National Science Challenge, administered by New Zealand
Ministry of Business, Innovation, and Employment. BM is funded by the
NSERC Alexander Graham Bell Canada Graduate Scholarship and the FRQNT
master's scholarship. LP acknowledges funding from NSERC Discovery Grant
(NSERC RGPIN-2019-05771). TP acknowledges financial support from the
Fondation Courtois, and NSERC through the Discovery Grants and Discovery
Accelerator Supplement programs. MJF is supported by an NSERC PDF and an
RBC Post-Doctoral Fellowship.

\textbf{Conflict of interest:} The authors have no conflict interests to
disclose

\textbf{Authors' contributions:} TS, and TP conceived the ideas
discussed in the manuscript. All authors contributed to writing and
editing the manuscript.

\textbf{Data availability:} There is no data associated with this
manuscript.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Adam2014EleTre}{}%
Adam, R. (2014). \emph{Elephant treaties: The Colonial legacy of the
biodiversity crisis}. UPNE.

\leavevmode\hypertarget{ref-Albouy2019MarFis}{}%
Albouy, C., Archambault, P., Appeltans, W., Araújo, M. B., Beauchesne,
D., Cazelles, K., Cirtwill, A. R., Fortin, M.-J., Galiana, N., Leroux,
S. J., Pellissier, L., Poisot, T., Stouffer, D. B., Wood, S. A., \&
Gravel, D. (2019). The marine fish food web is globally connected.
\emph{Nature Ecology \& Evolution}, \emph{3}(8, 8), 1153--1161.
\url{https://doi.org/10.1038/s41559-019-0950-y}

\leavevmode\hypertarget{ref-Anowar2021ConEmp}{}%
Anowar, F., Sadaoui, S., \& Selim, B. (2021). Conceptual and empirical
comparison of dimensionality reduction algorithms (PCA, KPCA, LDA, MDS,
SVD, LLE, ISOMAP, LE, ICA, t-SNE). \emph{Computer Science Review},
\emph{40}, 100378. \url{https://doi.org/10.1016/j.cosrev.2021.100378}

\leavevmode\hypertarget{ref-Arsov2019NetEmb}{}%
Arsov, N., \& Mirceva, G. (2019). \emph{Network Embedding: An Overview}.
\url{http://arxiv.org/abs/1911.11726}

\leavevmode\hypertarget{ref-Bartomeus2016ComFra}{}%
Bartomeus, I., Gravel, D., Tylianakis, J. M., Aizen, M. A., Dickie, I.
A., \& Bernard-Verdier, M. (2016). A common framework for identifying
linkage rules across different types of interactions. \emph{Functional
Ecology}, \emph{30}(12), 1894--1903.
\url{http://onlinelibrary.wiley.com/doi/10.1111/1365-2435.12666/full}

\leavevmode\hypertarget{ref-Bezanson2017JulFre}{}%
Bezanson, J., Edelman, A., Karpinski, S., \& Shah, V. (2017). Julia: A
Fresh Approach to Numerical Computing. \emph{SIAM Review}, \emph{59}(1),
65--98. \url{https://doi.org/10.1137/141000671}

\leavevmode\hypertarget{ref-Bohan2017NexGlo}{}%
Bohan, D. A., Vacher, C., Tamaddoni-Nezhad, A., Raybould, A., Dumbrell,
A. J., \& Woodward, G. (2017). Next-Generation Global Biomonitoring:
Large-scale, Automated Reconstruction of Ecological Networks.
\emph{Trends in Ecology \& Evolution}.
\url{https://doi.org/10.1016/j.tree.2017.03.001}

\leavevmode\hypertarget{ref-Botella2022AppGra}{}%
Botella, C., Dray, S., Matias, C., Miele, V., \& Thuiller, W. (2022). An
appraisal of graph embeddings for comparing trophic network
architectures. \emph{Methods in Ecology and Evolution}, \emph{13}(1),
203--216. \url{https://doi.org/10.1111/2041-210X.13738}

\leavevmode\hypertarget{ref-Braga2019SpaAna}{}%
Braga, J., Pollock, L. J., Barros, C., Galiana, N., Montoya, J. M.,
Gravel, D., Maiorano, L., Montemaggiori, A., Ficetola, G. F., Dray, S.,
\& Thuiller, W. (2019). Spatial analyses of multi-trophic terrestrial
vertebrate assemblages in Europe. \emph{Global Ecology and
Biogeography}, \emph{28}(11), 1636--1648.
\url{https://doi.org/10.1111/geb.12981}

\leavevmode\hypertarget{ref-BramonMora2018IdeCom}{}%
Bramon Mora, B., Gravel, D., Gilarranz, L. J., Poisot, T., \& Stouffer,
D. B. (2018). Identifying a common backbone of interactions underlying
food webs from different ecosystems. \emph{Nature Communications},
\emph{9}(1), 2603. \url{https://doi.org/10.1038/s41467-018-05056-0}

\leavevmode\hypertarget{ref-Buxton2021KeyInf}{}%
Buxton, R. T., Bennett, J. R., Reid, A. J., Shulman, C., Cooke, S. J.,
Francis, C. M., Nyboer, E. A., Pritchard, G., Binley, A. D., Avery-Gomm,
S., Ban, N. C., Beazley, K. F., Bennett, E., Blight, L. K., Bortolotti,
L. E., Camfield, A. F., Gadallah, F., Jacob, A. L., Naujokaitis-Lewis,
I., \ldots{} Smith, P. A. (2021). Key information needs to move from
knowledge to action for biodiversity conservation in Canada.
\emph{Biological Conservation}, \emph{256}, 108983.
\url{https://doi.org/10.1016/j.biocon.2021.108983}

\leavevmode\hypertarget{ref-Chami2022MacLea}{}%
Chami, I., Abu-El-Haija, S., Perozzi, B., Ré, C., \& Murphy, K. (2022).
Machine Learning on Graphs: A Model and Comprehensive Taxonomy.
\emph{Journal of Machine Learning Research}, \emph{23}(89), 1--64.
\url{http://jmlr.org/papers/v23/20-852.html}

\leavevmode\hypertarget{ref-Chen2017DeeMul}{}%
Chen, D., Xue, Y., Fink, D., Chen, S., \& Gomes, C. P. (2017).
\emph{Deep Multi-species Embedding}. 3639--3646.
\url{https://www.ijcai.org/proceedings/2017/509}

\leavevmode\hypertarget{ref-Chen2017HarHie}{}%
Chen, H., Perozzi, B., Hu, Y., \& Skiena, S. (2017). \emph{HARP:
Hierarchical Representation Learning for Networks}.
\url{http://arxiv.org/abs/1706.07845}

\leavevmode\hypertarget{ref-Choudry2013SavBio}{}%
Choudry, A. (2013). Saving biodiversity, for whom and for what?
Conservation NGOs, complicity, colonialism and conquest in an era of
capitalist globalization. In \emph{NGOization: Complicity,
contradictions and prospects} (pp. 24--44). Bloomsbury Publishing.

\leavevmode\hypertarget{ref-Cieslak2020TdiSto}{}%
Cieslak, M. C., Castelfranco, A. M., Roncalli, V., Lenz, P. H., \&
Hartline, D. K. (2020). T-Distributed Stochastic Neighbor Embedding
(t-SNE): A tool for eco-physiological transcriptomic analysis.
\emph{Marine Genomics}, \emph{51}, 100723.
\url{https://doi.org/10.1016/j.margen.2019.100723}

\leavevmode\hypertarget{ref-Csermely2004StrLin}{}%
Csermely, P. (2004). Strong links are important, but weak links
stabilize them. \emph{Trends in Biochemical Sciences}, \emph{29}(7),
331--334. \url{https://doi.org/10.1016/j.tibs.2004.05.004}

\leavevmode\hypertarget{ref-DallaRiva2016ExpEvo}{}%
Dalla Riva, G. V., \& Stouffer, D. B. (2016). Exploring the evolutionary
signature of food webs' backbones using functional traits. \emph{Oikos},
\emph{125}(4), 446--456. \url{https://doi.org/10.1111/oik.02305}

\leavevmode\hypertarget{ref-Dallas2017PreCry}{}%
Dallas, T., Park, A. W., \& Drake, J. M. (2017). Predicting cryptic
links in host-parasite networks. \emph{PLOS Computational Biology},
\emph{13}(5), e1005557.
\url{https://doi.org/10.1371/journal.pcbi.1005557}

\leavevmode\hypertarget{ref-Danisch2021MakJl}{}%
Danisch, S., \& Krumbiegel, J. (2021). Makie.jl: Flexible
high-performance data visualization for Julia. \emph{Journal of Open
Source Software}, \emph{6}(65), 3349.
\url{https://doi.org/10.21105/joss.03349}

\leavevmode\hypertarget{ref-Dominguez2020DecCon}{}%
Domínguez, L., \& Luoma, C. (2020). Decolonising Conservation Policy:
How Colonial Land and Conservation Ideologies Persist and Perpetuate
Indigenous Injustices at the Expense of the Environment. \emph{Land},
\emph{9}(3, 3), 65. \url{https://doi.org/10.3390/land9030065}

\leavevmode\hypertarget{ref-Dunne2006NetStr}{}%
Dunne, J. A. (2006). The Network Structure of Food Webs. In J. A. Dunne
\& M. Pascual (Eds.), \emph{Ecological networks: Linking structure and
dynamics} (pp. 27--86). Oxford University Press.

\leavevmode\hypertarget{ref-Eero2021UseFoo}{}%
Eero, M., Dierking, J., Humborg, C., Undeman, E., MacKenzie, B. R.,
Ojaveer, H., Salo, T., \& Köster, F. W. (2021). Use of food web
knowledge in environmental conservation and management of living
resources in the Baltic Sea. \emph{ICES Journal of Marine Science},
\emph{78}(8), 2645--2663. \url{https://doi.org/10.1093/icesjms/fsab145}

\leavevmode\hypertarget{ref-Eichhorn2019SteDec}{}%
Eichhorn, M. P., Baker, K., \& Griffiths, M. (2019). Steps towards
decolonising biogeography. \emph{Frontiers of Biogeography},
\emph{12}(1), 1--7. \url{https://doi.org/10.21425/F5FBG44795}

\leavevmode\hypertarget{ref-Eklof2013DimEco}{}%
Eklöf, A., Jacob, U., Kopp, J., Bosch, J., Castro-Urgal, R., Chacoff, N.
P., Dalsgaard, B., de Sassi, C., Galetti, M., Guimarães, P. R.,
Lomáscolo, S. B., Martín González, A. M., Pizo, M. A., Rader, R.,
Rodrigo, A., Tylianakis, J. M., Vázquez, D. P., \& Allesina, S. (2013).
The dimensionality of ecological networks. \emph{Ecology Letters},
\emph{16}(5), 577--583. \url{https://doi.org/10.1111/ele.12081}

\leavevmode\hypertarget{ref-Fortin2021NetEco}{}%
Fortin, M.-J., Dale, M. R. T., \& Brimacombe, C. (2021). Network ecology
in dynamic landscapes. \emph{Proceedings of the Royal Society B:
Biological Sciences}, \emph{288}(1949), rspb.2020.1889, 20201889.
\url{https://doi.org/10.1098/rspb.2020.1889}

\leavevmode\hypertarget{ref-Fricke2022EffDef}{}%
Fricke, E. C., Ordonez, A., Rogers, H. S., \& Svenning, J.-C. (2022).
The effects of defaunation on plants' capacity to track climate change.
\emph{Science}.
\url{https://www.science.org/doi/abs/10.1126/science.abk3510}

\leavevmode\hypertarget{ref-Galiana2021SpaSca}{}%
Galiana, N., Barros, C., Braga, J., Ficetola, G. F., Maiorano, L.,
Thuiller, W., Montoya, J. M., \& Lurgi, M. (2021). The spatial scaling
of food web structure across European biogeographical regions.
\emph{Ecography}, \emph{n/a}(n/a).
\url{https://doi.org/10.1111/ecog.05229}

\leavevmode\hypertarget{ref-Galiana2019GeoVar}{}%
Galiana, N., Hawkins, B. A., \& Montoya, J. M. (2019). The geographical
variation of network structure is scale dependent: Understanding the
biotic specialization of host--parasitoid networks. \emph{Ecography},
\emph{42}(6), 1175--1187. \url{https://doi.org/10.1111/ecog.03684}

\leavevmode\hypertarget{ref-Galiana2022EcoNet}{}%
Galiana, N., Lurgi, M., Bastazini, V. A. G., Bosch, J., Cagnolo, L.,
Cazelles, K., Claramunt-López, B., Emer, C., Fortin, M.-J., Grass, I.,
Hernández-Castellano, C., Jauker, F., Leroux, S. J., McCann, K., McLeod,
A. M., Montoya, D., Mulder, C., Osorio-Canadas, S., Reverté, S.,
\ldots{} Montoya, J. M. (2022). Ecological network complexity scales
with area. \emph{Nature Ecology \& Evolution}, 1--8.
\url{https://doi.org/10.1038/s41559-021-01644-4}

\leavevmode\hypertarget{ref-Galiana2018SpaSca}{}%
Galiana, N., Lurgi, M., Claramunt-López, B., Fortin, M.-J., Leroux, S.,
Cazelles, K., Gravel, D., \& Montoya, J. M. (2018). The spatial scaling
of species interaction networks. \emph{Nature Ecology \& Evolution},
\emph{2}(5), 782--790. \url{https://doi.org/10.1038/s41559-018-0517-3}

\leavevmode\hypertarget{ref-Gibb2021DatPro}{}%
Gibb, R., Albery, G. F., Becker, D. J., Brierley, L., Connor, R.,
Dallas, T. A., Eskew, E. A., Farrell, M. J., Rasmussen, A. L., Ryan, S.
J., Sweeny, A., Carlson, C. J., \& Poisot, T. (2021). Data
Proliferation, Reconciliation, and Synthesis in Viral Ecology.
\emph{BioScience}, \emph{71}(11), 1148--1156.
\url{https://doi.org/10.1093/biosci/biab080}

\leavevmode\hypertarget{ref-Goodfellow2016DeeLea}{}%
Goodfellow, I., Bengio, Y., \& Courville, A. (2016). \emph{Deep
learning}. MIT Press.

\leavevmode\hypertarget{ref-Goyal2018GraEmb}{}%
Goyal, P., \& Ferrara, E. (2018). Graph embedding techniques,
applications, and performance: A survey. \emph{Knowledge-Based Systems},
\emph{151}, 78--94. \url{https://doi.org/10.1016/j.knosys.2018.03.022}

\leavevmode\hypertarget{ref-Gravel2018BriElt}{}%
Gravel, D., Baiser, B., Dunne, J. A., Kopelke, J.-P., Martinez, N. D.,
Nyman, T., Poisot, T., Stouffer, D. B., Tylianakis, J. M., Wood, S. A.,
\& Roslin, T. (2018). Bringing Elton and Grinnell together: A
quantitative framework to represent the biogeography of ecological
interaction networks. \emph{Ecography}, \emph{0}(0).
\url{https://doi.org/10.1111/ecog.04006}

\leavevmode\hypertarget{ref-Green2019DisInt}{}%
Green, B., \& Chen, Y. (2019). Disparate Interactions: An
Algorithm-in-the-Loop Analysis of Fairness in Risk Assessments.
\emph{Proceedings of the Conference on Fairness, Accountability, and
Transparency}, 90--99. \url{https://doi.org/10.1145/3287560.3287563}

\leavevmode\hypertarget{ref-Grover2016NodSca}{}%
Grover, A., \& Leskovec, J. (2016). Node2vec: Scalable Feature Learning
for Networks. \emph{Proceedings of the 22nd ACM SIGKDD International
Conference on Knowledge Discovery and Data Mining}, 855--864.
\url{https://doi.org/10.1145/2939672.2939754}

\leavevmode\hypertarget{ref-Grunig2020CroFor}{}%
Grünig, M., Mazzi, D., Calanca, P., Karger, D. N., \& Pellissier, L.
(2020). Crop and forest pest metawebs shift towards increased linkage
and suitability overlap under climate change. \emph{Communications
Biology}, \emph{3}(1, 1), 1--10.
\url{https://doi.org/10.1038/s42003-020-0962-9}

\leavevmode\hypertarget{ref-Gupta2021GraNeu}{}%
Gupta, A., Matta, P., \& Pant, B. (2021). Graph neural network: Current
state of Art, challenges and applications. \emph{Materials Today:
Proceedings}, \emph{46}, 10927--10932.
\url{https://doi.org/10.1016/j.matpr.2021.01.950}

\leavevmode\hypertarget{ref-Hadfield2014TalTwo}{}%
Hadfield, J. D., Krasnov, B. R., Poulin, R., \& Nakagawa, S. (2014). A
Tale of Two Phylogenies: Comparative Analyses of Ecological
Interactions. \emph{The American Naturalist}, \emph{183}(2), 174--187.
\url{https://doi.org/10.1086/674445}

\leavevmode\hypertarget{ref-Herbert1965Dun}{}%
Herbert, F. (1965). \emph{Dune} (1st ed.). Chilton Book Company.

\leavevmode\hypertarget{ref-Hinton2002StoNei}{}%
Hinton, G., \& Roweis, S. T. (2002). Stochastic neighbor embedding.
\emph{NIPS}, \emph{15}, 833--840.

\leavevmode\hypertarget{ref-Hoffmann2019MacLea}{}%
Hoffmann, J., Bar-Sinai, Y., Lee, L. M., Andrejevic, J., Mishra, S.,
Rubinstein, S. M., \& Rycroft, C. H. (2019). Machine learning in a
data-limited regime: Augmenting experiments with synthetic data uncovers
order in crumpled sheets. \emph{Science Advances}, \emph{5}(4),
eaau6792. \url{https://doi.org/10.1126/sciadv.aau6792}

\leavevmode\hypertarget{ref-Hortal2015SevSho}{}%
Hortal, J., de Bello, F., Diniz-Filho, J. A. F., Lewinsohn, T. M., Lobo,
J. M., \& Ladle, R. J. (2015). Seven Shortfalls that Beset Large-Scale
Knowledge of Biodiversity. \emph{Annual Review of Ecology, Evolution,
and Systematics}, \emph{46}(1), 523--549.
\url{https://doi.org/10.1146/annurev-ecolsys-112414-054400}

\leavevmode\hypertarget{ref-Jordano2016SamNet}{}%
Jordano, P. (2016). Sampling networks of ecological interactions.
\emph{Functional Ecology}, \emph{30}(12), 1883--1893.
\url{https://doi.org/10.1111/1365-2435.12763}

\leavevmode\hypertarget{ref-Lamba2019DeeLea}{}%
Lamba, A., Cassey, P., Segaran, R. R., \& Koh, L. P. (2019). Deep
learning for environmental conservation. \emph{Current Biology},
\emph{29}(19), R977--R982.
\url{https://doi.org/10.1016/j.cub.2019.08.016}

\leavevmode\hypertarget{ref-Maaten2009LeaPar}{}%
Maaten, L. van der. (2009). Learning a Parametric Embedding by
Preserving Local Structure. \emph{Proceedings of the Twelth
International Conference on Artificial Intelligence and Statistics},
384--391. \url{https://proceedings.mlr.press/v5/maaten09a.html}

\leavevmode\hypertarget{ref-Machen2021ThiAlg}{}%
Machen, R., \& Nost, E. (2021). Thinking algorithmically: The making of
hegemonic knowledge in climate governance. \emph{Transactions of the
Institute of British Geographers}, \emph{46}(3), 555--569.
\url{https://doi.org/10.1111/tran.12441}

\leavevmode\hypertarget{ref-McLeod2021SamAsy}{}%
McLeod, A., Leroux, S. J., Gravel, D., Chu, C., Cirtwill, A. R., Fortin,
M.-J., Galiana, N., Poisot, T., \& Wood, S. A. (2021). Sampling and
asymptotic network properties of spatial multi-trophic networks.
\emph{Oikos}, \emph{n/a}(n/a). \url{https://doi.org/10.1111/oik.08650}

\leavevmode\hypertarget{ref-Melnyk2020GraGra}{}%
Melnyk, K., Klus, S., Montavon, G., \& Conrad, T. O. F. (2020).
GraphKKE: Graph Kernel Koopman embedding for human microbiome analysis.
\emph{Applied Network Science}, \emph{5}(1), 96.
\url{https://doi.org/10.1007/s41109-020-00339-2}

\leavevmode\hypertarget{ref-Morales-Castilla2015InfBio}{}%
Morales-Castilla, I., Matias, M. G., Gravel, D., \& Araújo, M. B.
(2015). Inferring biotic interactions from proxies. \emph{Trends in
Ecology \& Evolution}, \emph{30}(6), 347--356.
\url{https://doi.org/10.1016/j.tree.2015.03.014}

\leavevmode\hypertarget{ref-MoseboFernandes2020MacLea}{}%
Mosebo Fernandes, A. C., Quintero Gonzalez, R., Lenihan-Clarke, M. A.,
Leslie Trotter, E. F., \& Jokar Arsanjani, J. (2020). Machine Learning
for Conservation Planning in a Changing Climate. \emph{Sustainability},
\emph{12}(18, 18), 7657. \url{https://doi.org/10.3390/su12187657}

\leavevmode\hypertarget{ref-Murphy2022ProMac}{}%
Murphy, K. P. (2022). \emph{Probabilistic machine learning: An
introduction}. MIT Press. \href{https://probml.ai}{probml.ai}

\leavevmode\hypertarget{ref-Naman2022FooWeb}{}%
Naman, S. M., White, S. M., Bellmore, J. R., McHugh, P. A., Kaylor, M.
J., Baxter, C. V., Danehy, R. J., Naiman, R. J., \& Puls, A. L. (2022).
Food web perspectives and methods for riverine fish conservation.
\emph{WIREs Water}, \emph{n/a}(n/a), e1590.
\url{https://doi.org/10.1002/wat2.1590}

\leavevmode\hypertarget{ref-Narayanan2017GraLea}{}%
Narayanan, A., Chandramohan, M., Venkatesan, R., Chen, L., Liu, Y., \&
Jaiswal, S. (2017). \emph{Graph2vec: Learning Distributed
Representations of Graphs}. \url{http://arxiv.org/abs/1707.05005}

\leavevmode\hypertarget{ref-Neutel2002StaRea}{}%
Neutel, A.-M., Heesterbeek, J. A. P., \& de Ruiter, P. C. (2002).
Stability in Real Food Webs: Weak Links in Long Loops. \emph{Science},
\emph{296}(5570), 1120--1123.
\url{https://doi.org/10.1126/science.1068326}

\leavevmode\hypertarget{ref-Nokmaq2021AwaSle}{}%
No'kmaq, M., Marshall, A., Beazley, K. F., Hum, J., joudry, shalan,
Papadopoulos, A., Pictou, S., Rabesca, J., Young, L., \& Zurba, M.
(2021). {``Awakening the sleeping giant''}: Re-Indigenization principles
for transforming biodiversity conservation in Canada and beyond.
\emph{FACETS}, \emph{6}(1), 839--869.

\leavevmode\hypertarget{ref-Nost2021PolEco}{}%
Nost, E., \& Goldstein, J. E. (2021). A political ecology of data.
\emph{Environment and Planning E: Nature and Space}, 25148486211043503.
\url{https://doi.org/10.1177/25148486211043503}

\leavevmode\hypertarget{ref-OConnor2020UnvFoo}{}%
O'Connor, L. M. J., Pollock, L. J., Braga, J., Ficetola, G. F.,
Maiorano, L., Martinez‐Almoyna, C., Montemaggiori, A., Ohlmann, M., \&
Thuiller, W. (2020). Unveiling the food webs of tetrapods across Europe
through the prism of the Eltonian niche. \emph{Journal of Biogeography},
\emph{47}(1), 181--192. \url{https://doi.org/10.1111/jbi.13773}

\leavevmode\hypertarget{ref-Pedersen2017SigCol}{}%
Pedersen, E. J., Thompson, P. L., Ball, R. A., Fortin, M.-J., Gouhier,
T. C., Link, H., Moritz, C., Nenzen, H., Stanley, R. R. E., Taranu, Z.
E., Gonzalez, A., Guichard, F., \& Pepin, P. (2017). Signatures of the
collapse and incipient recovery of an overexploited marine ecosystem.
\emph{Royal Society Open Science}, \emph{4}(7), 170215.
\url{https://doi.org/10.1098/rsos.170215}

\leavevmode\hypertarget{ref-Perozzi2014DeeOnl}{}%
Perozzi, B., Al-Rfou, R., \& Skiena, S. (2014). DeepWalk: Online
learning of social representations. \emph{Proceedings of the 20th ACM
SIGKDD International Conference on Knowledge Discovery and Data Mining},
701--710. \url{https://doi.org/10.1145/2623330.2623732}

\leavevmode\hypertarget{ref-Pichler2020MacLea}{}%
Pichler, M., Boreux, V., Klein, A.-M., Schleuning, M., \& Hartig, F.
(2020). Machine learning algorithms to infer trait-matching and predict
species interactions in ecological networks. \emph{Methods in Ecology
and Evolution}, \emph{11}(2), 281--293.
\url{https://doi.org/10.1111/2041-210X.13329}

\leavevmode\hypertarget{ref-Poisot2019EcoJl}{}%
Poisot, T., Belisle, Z., Hoebeke, L., Stock, M., \& Szefer, P. (2019).
EcologicalNetworks.jl - analysing ecological networks. \emph{Ecography}.
\url{https://doi.org/10.1111/ecog.04310}

\leavevmode\hypertarget{ref-Poisot2016StrPro}{}%
Poisot, T., Cirtwill, A. R., Cazelles, K., Gravel, D., Fortin, M.-J., \&
Stouffer, D. B. (2016). The structure of probabilistic networks.
\emph{Methods in Ecology and Evolution}, \emph{7}(3), 303--312.
\url{https://doi.org/10.1111/2041-210X.12468}

\leavevmode\hypertarget{ref-Poisot2021ImpMam}{}%
Poisot, T., Ouellet, M.-A., Mollentze, N., Farrell, M. J., Becker, D.
J., Albery, G. F., Gibb, R. J., Seifert, S. N., \& Carlson, C. J.
(2021). \emph{Imputing the mammalian virome with linear filtering and
singular value decomposition}. \url{http://arxiv.org/abs/2105.14973}

\leavevmode\hypertarget{ref-Poisot2015SpeWhy}{}%
Poisot, T., Stouffer, D. B., \& Gravel, D. (2015). Beyond species: Why
ecological interaction networks vary through space and time.
\emph{Oikos}, \emph{124}(3), 243--251.
\url{https://doi.org/10.1111/oik.01719}

\leavevmode\hypertarget{ref-Ramasamy2015ComSpe}{}%
Ramasamy, D., \& Madhow, U. (2015). Compressive spectral embedding:
Sidestepping the SVD. In C. Cortes, N. Lawrence, D. Lee, M. Sugiyama, \&
R. Garnett (Eds.), \emph{Advances in neural information processing
systems} (Vol. 28). Curran Associates, Inc.
\url{https://proceedings.neurips.cc/paper/2015/file/4f6ffe13a5d75b2d6a3923922b3922e5-Paper.pdf}

\leavevmode\hypertarget{ref-Ray2021BioCri}{}%
Ray, J. C., Grimm, J., \& Olive, A. (2021). The biodiversity crisis in
Canada: Failures and challenges of federal and sub-national strategic
and legal frameworks. \emph{FACETS}, \emph{6}, 1044--1068.
\url{https://doi.org/10.1139/facets-2020-0075}

\leavevmode\hypertarget{ref-Runghen2021ExpNod}{}%
Runghen, R., Stouffer, D. B., \& Dalla Riva, G. V. (2021).
\emph{Exploiting node metadata to predict interactions in large networks
using graph embedding and neural networks}.
\url{https://doi.org/10.1101/2021.06.10.447991}

\leavevmode\hypertarget{ref-Saravia2021EcoNet}{}%
Saravia, L. A., Marina, T. I., Kristensen, N. P., De Troch, M., \& Momo,
F. R. (2021). Ecological network assembly: How the regional metaweb
influences local food webs. \emph{Journal of Animal Ecology},
\emph{n/a}(n/a). \url{https://doi.org/10.1111/1365-2656.13652}

\leavevmode\hypertarget{ref-Stevenson2021AlgRis}{}%
Stevenson, M. T., \& Doleac, J. L. (2021). \emph{Algorithmic Risk
Assessment in the Hands of Humans} {[}SSRN Scholarly Paper{]}.
\url{https://doi.org/10.2139/ssrn.3489440}

\leavevmode\hypertarget{ref-Stier2017IntExp}{}%
Stier, A. C., Samhouri, J. F., Gray, S., Martone, R. G., Mach, M. E.,
Halpern, B. S., Kappel, C. V., Scarborough, C., \& Levin, P. S. (2017).
Integrating Expert Perceptions into Food Web Conservation and
Management. \emph{Conservation Letters}, \emph{10}(1), 67--76.
\url{https://doi.org/10.1111/conl.12245}

\leavevmode\hypertarget{ref-Strydom2022FooWeb}{}%
Strydom, T., Bouskila, S., Banville, F., Barros, C., Caron, D., Farrell,
M. J., Fortin, M.-J., Hemming, V., Mercier, B., Pollock, L. J., Runghen,
R., Dalla Riva, G. V., \& Poisot, T. (2022). Food web reconstruction
through phylogenetic transfer of low-rank network representation.
\emph{Methods in Ecology and Evolution}, \emph{n/a}(n/a).
\url{https://doi.org/10.1111/2041-210X.13835}

\leavevmode\hypertarget{ref-Strydom2021RoaPre}{}%
Strydom, T., Catchen, M. D., Banville, F., Caron, D., Dansereau, G.,
Desjardins-Proulx, P., Forero-Muñoz, N. R., Higino, G., Mercier, B.,
Gonzalez, A., Gravel, D., Pollock, L., \& Poisot, T. (2021). A roadmap
towards predicting species interaction networks (across space and time).
\emph{Philosophical Transactions of the Royal Society B: Biological
Sciences}, \emph{376}(1837), 20210063.
\url{https://doi.org/10.1098/rstb.2021.0063}

\leavevmode\hypertarget{ref-Strydom2021SvdEnt}{}%
Strydom, T., Dalla Riva, G. V., \& Poisot, T. (2021). SVD Entropy
Reveals the High Complexity of Ecological Networks. \emph{Frontiers in
Ecology and Evolution}, \emph{9}.
\url{https://doi.org/10.3389/fevo.2021.623141}

\leavevmode\hypertarget{ref-Tang2015LinLar}{}%
Tang, J., Qu, M., Wang, M., Zhang, M., Yan, J., \& Mei, Q. (2015). LINE:
Large-scale Information Network Embedding. \emph{Proceedings of the 24th
International Conference on World Wide Web}, 1067--1077.
\url{https://doi.org/10.1145/2736277.2741093}

\leavevmode\hypertarget{ref-Turak2017UsiEss}{}%
Turak, E., Brazill-Boast, J., Cooney, T., Drielsma, M., DelaCruz, J.,
Dunkerley, G., Fernandez, M., Ferrier, S., Gill, M., Jones, H., Koen,
T., Leys, J., McGeoch, M., Mihoub, J.-B., Scanes, P., Schmeller, D., \&
Williams, K. (2017). Using the essential biodiversity variables
framework to measure biodiversity change at national scale.
\emph{Biological Conservation}, \emph{213}, 264--271.
\url{https://doi.org/10.1016/j.biocon.2016.08.019}

\leavevmode\hypertarget{ref-Wang2016StrDee}{}%
Wang, D., Cui, P., \& Zhu, W. (2016). Structural Deep Network Embedding.
\emph{Proceedings of the 22nd ACM SIGKDD International Conference on
Knowledge Discovery and Data Mining}, 1225--1234.
\url{https://doi.org/10.1145/2939672.2939753}

\leavevmode\hypertarget{ref-Wang2021JoiEmb}{}%
Wang, S., Arroyo, J., Vogelstein, J. T., \& Priebe, C. E. (2021). Joint
Embedding of Graphs. \emph{IEEE Transactions on Pattern Analysis and
Machine Intelligence}, \emph{43}(4), 1324--1336.
\url{https://doi.org/10.1109/TPAMI.2019.2948619}

\leavevmode\hypertarget{ref-Wardeh2021PreMam}{}%
Wardeh, M., Baylis, M., \& Blagrove, M. S. C. (2021). Predicting
mammalian hosts in which novel coronaviruses can be generated.
\emph{Nature Communications}, \emph{12}(1, 1), 780.
\url{https://doi.org/10.1038/s41467-021-21034-5}

\leavevmode\hypertarget{ref-Weiskopf2022IncUpt}{}%
Weiskopf, S. R., Harmáčková, Z. V., Johnson, C. G., Londoño-Murcia, M.
C., Miller, B. W., Myers, B. J. E., Pereira, L., Arce-Plata, M. I.,
Blanchard, J. L., Ferrier, S., Fulton, E. A., Harfoot, M., Isbell, F.,
Johnson, J. A., Mori, A. S., Weng, E., \& Rosa, I. M. D. (2022).
Increasing the uptake of ecological model results in policy decisions to
improve biodiversity outcomes. \emph{Environmental Modelling \&
Software}, \emph{149}, 105318.
\url{https://doi.org/10.1016/j.envsoft.2022.105318}

\leavevmode\hypertarget{ref-Wood2015EffSpa}{}%
Wood, S. A., Russell, R., Hanson, D., Williams, R. J., \& Dunne, J. A.
(2015). Effects of spatial scale of sampling on food web structure.
\emph{Ecology and Evolution}, \emph{5}(17), 3769--3782.
\url{https://doi.org/10.1002/ece3.1640}

\leavevmode\hypertarget{ref-Xu2020UndGra}{}%
Xu, M. (2020). \emph{Understanding graph embedding methods and their
applications}. \url{http://arxiv.org/abs/2012.08019}

\leavevmode\hypertarget{ref-Yan2005GraEmb}{}%
Yan, S., Xu, D., Zhang, B., \& Zhang, H.-J. (2005). Graph embedding: A
general framework for dimensionality reduction. \emph{2005 IEEE Computer
Society Conference on Computer Vision and Pattern Recognition
(CVPR'05)}, \emph{2}, 830--837 vol. 2.
\url{https://doi.org/10.1109/CVPR.2005.170}

\leavevmode\hypertarget{ref-Young2007RanDot}{}%
Young, S. J., \& Scheinerman, E. R. (2007). Random Dot Product Graph
Models for Social Networks. In A. Bonato \& F. R. K. Chung (Eds.),
\emph{Algorithms and Models for the Web-Graph} (pp. 138--149). Springer.
\url{https://doi.org/10.1007/978-3-540-77004-6_11}

\leavevmode\hypertarget{ref-Zhou2020GraNeu}{}%
Zhou, J., Cui, G., Hu, S., Zhang, Z., Yang, C., Liu, Z., Wang, L., Li,
C., \& Sun, M. (2020). Graph neural networks: A review of methods and
applications. \emph{AI Open}, \emph{1}, 57--81.
\url{https://doi.org/10.1016/j.aiopen.2021.01.001}

\end{CSLReferences}

\end{document}