### 2023-01-25 Strydom_et_al_2023_graph_embedding_draft.tex
Deploying to gh-pages from @ PoisotLab/ms_metaweb_perspectives@347a9937e2eb5fc9f006894aadc63420e927d746 🚀

%!TEX TS-program = xelatex
\documentclass[11pt]{article}

\usepackage[english]{babel}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{stix2}
\usepackage[scaled]{helvet}
\usepackage[scaled]{inconsolata}

\usepackage{lastpage}

\usepackage{gensymb}

\usepackage{setspace}

\usepackage{ccicons}

\usepackage[hang,flushmargin]{footmisc}

\usepackage{geometry}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}

\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\makeatletter
\newcounter{tableno}
\newenvironment{tablenos:no-prefix-table-caption}{
  \caption@ifcompatibility{}{
    \let\oldthetable\thetable
    \let\oldtheHtable\theHtable
    \renewcommand{\thetable}{tableno:\thetableno}
    \renewcommand{\theHtable}{tableno:\thetableno}
    \stepcounter{tableno}
    \captionsetup{labelformat=empty}
  }
}{
  \caption@ifcompatibility{}{
    \captionsetup{labelformat=default}
    \let\thetable\oldthetable
    \let\theHtable\oldtheHtable
    \addtocounter{table}{-1}
  }
}
\makeatother

\usepackage{array}
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\let\PBS=\PreserveBackslash

\usepackage[breaklinks=true]{hyperref}
\hypersetup{colorlinks,%
citecolor=blue,%
filecolor=blue,%
linkcolor=blue,%
urlcolor=blue}
\usepackage{url}

\usepackage{caption}
\setcounter{secnumdepth}{0}
\usepackage{cleveref}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}

\usepackage{longtable}
\usepackage{booktabs}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for \widthof, \maxof
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\maxof{\widthof{#1}}{\csllabelwidth}}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\geometry{verbose,letterpaper,tmargin=2.2cm,bmargin=2.2cm,lmargin=2.2cm,rmargin=2.2cm}

\usepackage{lineno}
\usepackage[nolists,noheads]{endfloat}

\pagestyle{plain}

\tolerance=1
\emergencystretch=\maxdimen
\hyphenpenalty=10000
\hbadness=10000

\doublespacing

\fancypagestyle{normal}
{
  \fancyhf{}
  \fancyfoot[R]{\footnotesize\sffamily\thepage\ of \pageref*{LastPage}}
}
\begin{document}
\raggedright
\thispagestyle{empty}
{\Large\bfseries\sffamily Graph embedding and transfer learning can help
predict species interaction networks despite data limitations}
\vskip 5em

%
\href{https://orcid.org/0000-0001-6067-1349}{Tanya\,Strydom}%
%
\,\textsuperscript{1,2,‡}\quad %
\href{https://orcid.org/0000-0003-0193-5441}{Salomé\,Bouskila}%
%
\,\textsuperscript{1}\quad %
\href{https://orcid.org/0000-0001-9051-0597}{Francis\,Banville}%
%
\,\textsuperscript{1,2,3}\quad %
\href{https://orcid.org/0000-0003-4036-977X}{Ceres\,Barros}%
%
\,\textsuperscript{4}\quad %
\href{https://orcid.org/0000-0002-2151-6693}{Dominique\,Caron}%
%
\,\textsuperscript{2,5}\quad %
\href{https://orcid.org/0000-0003-0452-6993}{Maxwell J\,Farrell}%
%
\,\textsuperscript{6}\quad %
\href{https://orcid.org/0000-0002-9935-1366}{Marie-Josée\,Fortin}%
%
\,\textsuperscript{6}\quad %
\href{https://orcid.org/0000-0003-3220-6161}{Victoria\,Hemming}%
%
\,\textsuperscript{7}\quad %
\href{https://orcid.org/0000-0002-4104-9463}{Benjamin\,Mercier}%
%
\,\textsuperscript{2,3}\quad %
\href{https://orcid.org/0000-0002-6004-4027}{Laura J.\,Pollock}%
%
\,\textsuperscript{2,5}\quad %
\href{https://orcid.org/0000-0001-5785-8321}{Rogini\,Runghen}%
%
\,\textsuperscript{8}\quad %
\href{https://orcid.org/0000-0002-3454-0633}{Giulio V.\,Dalla Riva}%
%
\,\textsuperscript{9}\quad %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{1,2,‡}

\textsuperscript{1}\,Département de Sciences Biologiques, Université de
Montréal, Montréal, Canada\quad \textsuperscript{2}\,Quebec Centre for
Biodiversity Science, Montréal,
Canada\quad \textsuperscript{3}\,Département de Biologie, Université de
Sherbrooke, Sherbrooke, Canada\quad \textsuperscript{4}\,Department of
Forest Resources Management, University of British Columbia, Vancouver,
B.C., Canada\quad \textsuperscript{5}\,Department of Biology, McGill
University, Montréal, Canada\quad \textsuperscript{6}\,Department of
Ecology \& Evolutionary Biology, University of Toronto, Toronto,
Canada\quad \textsuperscript{7}\,Department of Forest and Conservation
Sciences, University of British Columbia, Vancouver,
Canada\quad \textsuperscript{8}\,Centre for Integrative Ecology, School
of Biological Sciences, University of Canterbury, Canterbury, New
Zealand\quad \textsuperscript{9}\,School of Mathematics and Statistics,
University of Canterbury, Canterbury, New Zealand

\textsuperscript{‡}\,Equal contributions\\

\textbf{Correspondance to:}\\
Timothée Poisot --- \texttt{timothee.poisot@umontreal.ca}\\

\vfill

\begin{enumerate}
    \item Metawebs, (networks of potential interactions within a species
pool) are a powerful abstraction to understand how large-scale species
interaction networks are structured.%
    \item Because metawebs are typically expressed at large spatial and
taxonomic scales, assembling them is a tedious and costly process;
predictive methods can help circumvent the limitations in data
deficiencies, by providing `draft' metawebs.%
    \item One way to improve our ability to predict metawebs is to
maximize available information by using graph embeddings, instead of the
list of species interactions. Graph embedding is an emerging field in
machine learning that holds great potential for ecological problems.%
    \item In this perspective, we outline how the challenges associated
with inferring metawebs line-up with the advantages of graph embeddings;
as well as discuss how the choice of the species pool has consequences
on the reconstructed network, but also embeds hypotheses about which
human-made boundaries are ecologically meaningful.%
\end{enumerate}

\vfill
This work is released by its authors under a CC-BY 4.0 license\hfill\ccby\\
Last revision: \emph{\today}

\clearpage
\thispagestyle{empty}

\vfill

\vfill

\clearpage
\linenumbers
\pagestyle{normal}

Being able to infer \emph{potential} interactions could serve as a
significant breakthrough in our ability to start thinking about species
interaction networks over large spatial scales (Hortal et al., 2015).
Understanding species interactions holds enormous potential to not only
understand and more rapidly learn about species interactions and
metawebs, but also how changes in management of a single species may
impact non-target species. In a recent overview of the field of
ecological network prediction, Strydom, Catchen, et al. (2021)
identified two challenges of interest to the prediction of interactions
at large scales. First, there is a relative scarcity of relevant data in
most places globally -- paradoxically, this restricts our ability to
infer interactions for locations where inference is perhaps the least
required (and leaves us unable to make inference in regions without
interaction data); second, accurate predictors are important for
accurate predictions, and the lack of methods that can leverage a small
amount of \emph{accurate} data is a serious impediment to our predictive
ability. In most places, our most reliable biodiversity knowledge is
that of a species pool (\emph{i.e.} a set of potentially interacting
species in a given area): through the analysis of databases like the
Global Biodiversity Information Facility (GBIF) or the International
Union for the Conservation of Nature (IUCN), it is possible to construct
a list of species in a region of interest; but inferring the potential
interactions between these species is difficult.

Following the definition of Dunne (2006), a metaweb is the ecological
network analogue to the species pool; specifically, it inventories all
\emph{potential} interactions between species for a spatially delimited
area (and so captures the \(\gamma\) diversity of interactions). The
metaweb is not a prediction of the network at a specific point within
the spatial area it covers: it will have a different structure, notably
by having a larger connectance (see \emph{e.g.} Wood et al., 2015) and
complexity (see \emph{e.g.} Galiana et al., 2022), from any of these
local networks. These local networks (which capture the \(\alpha\)
diversity of interactions) are a subset of the metaweb's species and
realized interactions, and have been called ``metaweb realizations''
(Poisot et al., 2015). Differences between local networks and their
metawebs are due to chance, species abundance and co-occurrence, local
environmental conditions, and local distribution of functional traits,
among others. Specifically, although co-occurrence can be driven by
interactions (Cazelles et al., 2016), co-occurrence alone is not a
predictor of interactions (Blanchet et al., 2020; Thurman et al., 2019),
and therefore lack of co-occurrence cannot be used to rule out lack of a
feasible interaction. Yet, recent results by Saravia et al. (2021)
strongly suggest that the local (metaweb) realizations only respond
weakly to local conditions: instead, they reflect constraints inherited
by the structure of their metaweb. This establishes the metaweb
structure as the core goal of predictive network ecology, as it is a
required information to accurately produce downscaled, local
predictions.

Because the metaweb represents the joint effect of functional,
phylogenetic, and macroecological processes (Morales-Castilla et al.,
2015), it holds valuable ecological information. Specifically, it
represents the ``upper bounds'' on what the composition of the local
networks, given a local species pool, can be (see \emph{e.g.} McLeod et
al., 2021); this information can help evaluate the ability of ecological
assemblages to withstand the effects of, for example, climate change
(Fricke et al., 2022). These local networks may be reconstructed given
an appropriate knowledge of local species composition and provide
information on the structure of food webs at finer spatial scales. This
has been done for example for tree-galler-parasitoid systems (Gravel et
al., 2018), fish trophic interactions (Albouy et al., 2019), tetrapod
trophic interactions (J. Braga et al., 2019; O'Connor et al., 2020), and
crop-pest networks (Grünig et al., 2020). In this contribution, we
highlight the power in viewing (and constructing) metawebs as
\emph{probabilistic} objects in the context of rare interactions,
discuss how a family of machine learning tools (graph embeddings and
transfer learning) can be used to overcome data limitations to metaweb
inference, and highlight how the use of metawebs introduces important
questions for the field of network ecology.

\hypertarget{a-metaweb-is-an-inherently-probabilistic-object}{%
\section{A metaweb is an inherently probabilistic
object}\label{a-metaweb-is-an-inherently-probabilistic-object}}

Treating interactions as probabilistic (as opposed to binary) events is
a more nuanced and realistic way to represent them. Dallas et al. (2017)
suggested that most links in ecological networks are cryptic,
\emph{i.e.} uncommon or hard to observe. This argument echoes Jordano
(2016): sampling ecological interactions is difficult because it
requires first the joint observation of two species, and then the
observation of their interaction. In addition, it is generally expected
weak or rare links to be more prevalent in networks than common or
strong links (Csermely, 2004), compared to strong, persistent links;
this is notably the case in food chains, wherein many weaker links are
key to the stability of a system (Neutel et al., 2002). In the light of
these observations, we expect to see an over-representation of
low-probability (rare) interactions under a model that accurately
predicts interaction probabilities.

Yet the original metaweb definition, and indeed most past uses of
metawebs, was based on the presence/absence of interactions. Moving
towards \emph{probabilistic} metawebs, by representing interactions as
Bernoulli events (see \emph{e.g.} Poisot et al., 2016), offers the
opportunity to weigh these rare interactions appropriately. The inherent
plasticity of interactions is important to capture: there have been
documented instances of food webs undergoing rapid collapse/recovery
cycles over short periods of time (\emph{e.g.} Pedersen et al., 2017).
Furthermore, because the structure of the metaweb cannot be known in
advance, it is important to rely on predictive tools that do not assume
a specific network topology for link prediction (Gaucher et al., 2021),
but are able to work on generalizations of the network. These
considerations emphasize why metaweb predictions should focus on
quantitative (preferentially probabilistic) predictions, and this should
constrain the suite of appropriate models used to predict them.

It is important to recall that a metaweb is intended as a catalogue of
all potential (feasible) interactions, which is then filtered for a
given application (Morales-Castilla et al., 2015). It is therefore
important to separate the interactions that happen ``almost surely''
(repeated observational data), ``almost never'' (repeated lack of
evidence \emph{or} evidence that the link is forbidden through
\emph{e.g.} trait mis-match), and interactions with a probability that
lays somewhere in between (Catchen et al., 2023). In a sense, that most
ecological interactions are elusive can call for a slightly different
approach to sampling: once the common interactions are documented, the
effort required in documenting each rare interaction will increase
exponentially (Jordano, 2016). Recent proposals in other fields relying
on machine learning approaches emphasize the idea that algorithms meant
to predict, through the assumption that they approximate the process
generating the data, can also act as data generators (Hoffmann et al.,
2019). High quality observational data can be used to infer core rules
underpinning network structure, and be supplemented with synthetic data
coming from predictive models trained on them, thereby increasing the
volume of information available for analysis. Indeed, Strydom, Catchen,
et al. (2021) suggested that knowing the metaweb may render the
prediction of local networks easier, because it fixes an ``upper bound''
on which interactions can exist. In this context, a probabilistic
metaweb represents an aggregation of informative priors on the
biological feasibility of interactions, which is usually hard to obtain
yet has possibly the most potential to boost our predictive ability
(Bartomeus, 2013; Bartomeus et al., 2016). This would represent a
departure from simple rules expressed at the network scale (\emph{e.g.}
Williams \& Martinez, 2000) to a view of network prediction based on
learning the rules that underpin interactions \emph{and} their
variability (Gupta et al., 2022).

\begin{figure}
\hypertarget{fig:embedding}{%
\centering
\includegraphics{figures/conceptual_2.png}
\caption{The embedding process (\textbf{A}) can help to identify links
(interactions) that may have been missed within the original community
(represented by the orange dashed lines, \textbf{B}). Transfer learning
allows for the prediction links (interactions) when novel species
(\textbf{C}) are included alongside the original community (\textbf{D}).
This is achieved by learning using other relevant predictors
(\emph{e.g.} traits) in conjunction with the known interactions to infer
latent values (\textbf{E}). Ultimately this allows us to predict links
(interactions) for species external to the original sample as well as
within sample links (\textbf{F}). Within this context the predicted (and
original) networks as well as the ecological predictors used (green
boxes) are products that can be quantified through field measurements,
whereas the embedded as well as imputed matrices (purple box) are
representative of a decomposition of the interaction
networks}\label{fig:embedding}
}
\end{figure}

\hypertarget{graph-embedding-offers-promises-for-the-inference-of-potential-interactions}{%
\section{Graph embedding offers promises for the inference of potential
interactions}\label{graph-embedding-offers-promises-for-the-inference-of-potential-interactions}}

Graph (or Network) embedding (fig.~\ref{fig:embedding}) is a family of
machine learning techniques, whose main task is to learn a mapping
function from a discrete graph to a continuous domain (Chami et al.,
2022; \textbf{arsov\_network\_2019?}). Their main goal is to learn a low
dimensional vector representations of the graph (embeddings), such that
its key properties (\emph{e.g.} local or global structures) are retained
in the embedding space (Yan et al., 2005). The embedding space may, but
will not necessarily, have lower dimensionality than the graph.
Ecological networks are promising candidates for the routine application
of embeddings, as they tend to possess a shared structural backbone (see
\emph{e.g.} Bramon Mora et al., 2018), which hints at structural
invariants in empirical data. Assuming that these structural invariants
are common enough, they would dominate the structure of networks, and
therefore be adequately captured by the first (lower) dimensions of an
embedding, without the need to measure derived aspects of their
structure (\emph{e.g.} motifs, paths, modularity, \ldots).

Indeed, food webs are inherently low-dimensional objects, and can be
adequately represented with less than ten dimensions (J. Braga et al.,
2019; M. P. Braga et al., 2021; Eklöf et al., 2013). Simulation results
by Botella et al. (2022) suggests that there is no dominant method to
identify architectural similarities between networks: multiple
approaches need to be tested and compared to the network descriptor of
interest on a problem-specific basis. This matches previous results on
graph embedding, wherein different embedding algorithms yield different
network embeddings (Goyal \& Ferrara, 2018), calling for a careful
selection of the problem-specific approach to use. In
tbl.~\ref{tbl:methods}, we present a selection of common graph and node
embedding methods, alongside examples of their use to predict
interactions or statistical associations between species. These methods
rely largely on linear algebra or pseudo-random walks on graphs. All
forms of embeddings presented in the table share the common property of
summarizing their objects into (sets of) dense feature vectors, that
capture the overall network structure, pairwise information on nodes,
and emergent aspects of the network, in a compressed way (\emph{i.e.}
with some information loss, as we later discuss in the illustration).
Node embeddings tend to focus on maintaining pairwise relationships
(\emph{i.e.} species interactions), while graph embeddings focus on
maintaining the network structure (\emph{i.e.} emergent properties).
Nevertheless, some graph embedding techniques (like RDPG, see
\emph{e.g.} Wu et al., 2021) will provide high-quality node-level
embeddings while also preserving network structure.

One prominent family of approaches we do not discuss in the present
manuscript is Graph Neural Networks {[}GNN; Zhou et al. (2020){]}. GNN
are, in a sense, a method to embed a graph into a dense subspace, but
belong to the family of deep learning methods, which has its own set of
practices (see \emph{e.g.} Goodfellow et al., 2016). An important issue
with methods based on deep learning is that, because their parameter
space is immense, the sample size of the data fed into them must be
similarly large (typically thousands of instances). This is a
requirement for the model to converge correctly during training, but
this assumption is unlikely to be met given the size of datasets
currently available for metawebs (or single time/location species
interaction networks). This data volume requirement is mostly absent
from the techniques we list below. Furthermore, GNN still have some
challenges related to their shallow structure, and concerns related to
scalability (see Gupta et al., 2021 for a review), which are mostly
absent from the methods listed in tbl.~\ref{tbl:methods}. Assuming that
the uptake of next-generation biomonitoring techniques does indeed
deliver larger datasets on species interactions (Bohan et al., 2017),
there is a potential for GNN to become an applicable
embedding/predictive technique in the coming years.

\hypertarget{tbl:methods}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1016}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1172}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1875}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3438}}@{}}
\caption{\label{tbl:methods}Overview of some common graph embedding
approaches, by type of embedded objects, alongside examples of their use
in the prediction of species interactions. These methods have not yet
been routinely used to predict species interactions; most examples that
we identified were either statistical associations, or analogues to
joint species distribution models. \(^a\): application is concerned with
\emph{statistical} interactions, which are not necessarilly direct
biotic interactions; \(^b\):application is concerned with joint-SDM-like
approach, which is also very close to statistical associations as
opposed to direct biotic interactions. Given the need to evaluate
different methods on a problem-specific basis, the fact that a lot of
methods have not been used on network problems is an opportunity for
benchmarking and method development. Note that the row for PCA also
applies to kernel/probabilistic PCA, which are variations on the more
general method of SVD. Note further that tSNE has been included because
it is frequently used to embed graphs, including of species
associations/interactions, despite not being strictly speaking, a graph
embedding technique (see \emph{e.g.} Chami et al., 2022)}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Method
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Object
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Technique
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Reference
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Application
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Method
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Object
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Technique
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Reference
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Application
\end{minipage} \\
\midrule()
\endhead
tSNE & nodes & statistical divergence & Hinton \& Roweis (2002) &
Cieslak et al. (2020) \(^a\) Gibb et al. (2021) \\
LINE & nodes & stochastic gradient descent & Tang et al. (2015) & \\
SDNE & nodes & gradient descent & D. Wang et al. (2016) & \\
node2vec & nodes & stochastic gradient descent & Grover \& Leskovec
(2016) & \\
HARP & nodes & meta-strategy & H. Chen et al. (2017) & \\
DMSE & joint nodes & deep neural network & D. Chen et al. (2017) & D.
Chen et al. (2017) \(^b\) \\
graph2vec & sub-graph & skipgram network & Narayanan et al. (2017) & \\
RDPG & graph & SVD & Young \& Scheinerman (2007) & Dalla Riva \&
Stouffer (2016); Poisot et al. (2021) \\
GLEE & graph & Laplacian eigenmap & Torres et al. (2020) & \\
DeepWalk & graph & stochastic gradient descent & Perozzi et al. (2014) &
Wardeh et al. (2021) \\
GraphKKE & graph & stochastic differential equation & Melnyk et al.
(2020) & Melnyk et al. (2020) \(^a\) \\
FastEmbed & graph & eigen decomposition & Ramasamy \& Madhow (2015) & \\
PCA & graph & eigen decomposition & Surendran (2013) & Strydom, Catchen,
et al. (2021) \\
Joint methods & multiple graphs & multiple strategies & S. Wang et al.
(2021) & \\
\bottomrule()
\end{longtable}

The popularity of graph embedding techniques in machine learning is more
than the search for structural invariants: graphs are discrete objects,
and machine learning techniques tend to handle continuous data better.
Bringing a sparse graph into a continuous, dense vector space (Xu, 2021)
opens up a broader variety of predictive algorithms, notably of the sort
that are able to predict events as probabilities (Murphy, 2022).
Furthermore, the projection of the graph itself is a representation that
can be learned; Runghen et al. (2021), for example, used a neural
network to learn the embedding of a network in which not all
interactions were known, based on the nodes' metadata. This example has
many parallels in ecology (see fig.~\ref{fig:embedding} \textbf{C}), in
which node metadata can be represented by phylogeny, abundance, or
functional traits. Using phylogeny as a source of information assumes
(or strives to capture) the action of evolutionary processes on network
structure, which at least for food webs have been well documented (M. P.
Braga et al., 2021; Dalla Riva \& Stouffer, 2016; Eklöf \& Stouffer,
2016; Stouffer et al., 2012; Stouffer et al., 2007); similarly, the use
of functional traits assumes that interactions can be inferred from the
knowledge of trait-matching rules, which is similarly well supported in
the empirical literature (Bartomeus, 2013; Bartomeus et al., 2016;
Goebel et al., 2023; Gravel et al., 2013). Relating this information to
an embedding rather than a list of networks measures would allow to
capture their effect on the more fundamental aspects of network
structure; conversely, the absence of a phylogenetic or functional
signal may suggest that evolutionary/trait processes are not strong
drivers of network structure, therefore opening a new way to perform
hypothesis testing.

Before moving further, it is important to clarify the epistemic status
of node values derived from embeddings: specifically, they are
\emph{not} functional traits, and therefore should not be discussed in
terms of effects or responses. As per the framework of Malaterre et al.
(2019), these values neither derive from, nor result in, changes in
organismal performance, and should therefore not be used to quantify
\emph{e.g.} functional diversity. This holds true even when there are
correlations between latent values and functional traits: although these
enable an ecological discussion of how traits condition the structure of
the network, the existence of a statistical relationship does not
elevate the latent values to the status of functional traits.

Rather than directly predicting biological rules (see \emph{e.g.}
Pichler et al., 2020 for an overview), which may be confounded by the
sparse nature of graph data, learning embeddings works in the
low-dimensional space that maximizes information about the network
structure. This approach is further justified by the observation, for
example, that the macro-evolutionary history of a network is adequately
represented by some graph embeddings {[}Random dot product graphs
(RDPG); see Dalla Riva \& Stouffer (2016){]}. In a recent publication,
Strydom et al. (2022) have used an embedding (based on RDPG) to project
a metaweb of trophic interactions between European mammals, and
transferred this information to mammals of Canada, using the
phylogenetic distance between related clades to infer the values in the
latent sub-space into which the European metaweb was projected. By
performing the RDPG step on re-constructed values, this approach yields
a probabilistic trophic metaweb for mammals of Canada based on knowledge
of European species, despite a limited (\(\approx\) 5\%) taxonomic
overlap.

Graph embeddings \emph{can} serve as a dimensionality reduction method.
For example, RDPG (Strydom et al., 2022) and t-SVD {[}truncated Singular
Value Decomposition; Poisot et al. (2021){]} typically embed networks
using fewer dimensions than the original network {[}the original network
has as many dimensions as species, and as many informative dimensions as
trophically unique species; Strydom, Dalla Riva, et al. (2021){]}. But
this is not necessarily the case -- indeed, one may perform a PCA (a
special case of SVD) to project the raw data into a subspace that
improves the efficacy of t-SNE {[}t-distributed stochastic neighbor
embedding; Maaten (2009){]}. There are many dimensionality reductions
(Anowar et al., 2021) that can be applied to an embedded network should
the need for dimensionality reduction (for example for data
visualisation) arise. In brief, many graph embeddings \emph{can} serve
as dimensionality reduction steps, but not all do, neither do all
dimensionality reduction methods provide adequate graph embedding
capacities. In the next section (and fig.~\ref{fig:illustration}), we
show how the amount of dimensionality reduction can affect the quality
of the embedding.

\hypertarget{an-illustration-of-metaweb-embedding}{%
\section{An illustration of metaweb
embedding}\label{an-illustration-of-metaweb-embedding}}

In this section, we illustrate the embedding of a collection of
bipartite networks collected by Hadfield et al. (2014), using t-SVD and
RDPG. Briefly, an RDPG decomposes a network into two subspaces (left and
right), which are matrices that when multiplied give an approximation of
the original network. RDPG has the particularly desirable properties of
being a graph embedding technique that produces relevant node-level
feature vectors, and provides good approximations of graphs with varied
structures (Athreya et al., 2017). The code to reproduce this example is
available as supplementary material (note, for the sake of comparison,
that Strydom, Catchen, et al., 2021 have an example using embedding
through PCA followed by prediction using a deep neural network on the
same dataset). The resulting (binary) metaweb \(\mathcal{M}\) has 2131
interactions between 206 parasites and 121 hosts, and its adjacency
matrix has full rank (\emph{i.e.} it represents a space with 121
dimensions). All analyses were done using Julia (Bezanson et al., 2017)
version 1.7.2, \emph{Makie.jl} (Danisch \& Krumbiegel, 2021), and
\emph{EcologicalNetworks.jl} (Poisot et al., 2019).

\begin{figure}
\hypertarget{fig:illustration1}{%
\centering
\includegraphics{figures/illustration-part1.png}
\caption{Validation of an embedding for a host-parasite metaweb, using
Random Dot Product Graphs. \textbf{A}, decrease in approximation error
as the number of dimensions in the subspaces increases. \textbf{B},
increase in cumulative variance explained as the number of ranks
considered increases; in \textbf{A} and \textbf{B}, the dot represents
the point of inflexion in the curve (at rank 39) estimated using the
finite differences method. \textbf{C}, position of hosts and parasites
in the space of latent variables on the first and second dimensions of
their respective subspaces (the results have been clamped to the unit
interval). \textbf{D}, predicted interaction weight from the RDPG based
on the status of the species pair in the
metaweb.}\label{fig:illustration1}
}
\end{figure}

In fig.~\ref{fig:illustration1}, we focus on some statistical checks of
the embedding. In panel \textbf{A}, we show that the averaged \(L_2\)
loss (\emph{i.e.} the sum of squared errors) between the empirical and
reconstructed metaweb decreases when the number of dimensions (rank) of
the subspace increases, with an inflection at 39 dimensions (out of 120
initially) according to the finite differences method. As discussed by
Runghen et al. (2021), there is often a trade-off between the number of
dimensions to use (more dimensions are more computationally demanding)
and the quality of the representation. In panel \textbf{B}, we show the
increase in cumulative variance explained at each rank, and visualize
that using 39 ranks explains about 70\% of the variance in the empirical
metaweb. This is a different information from the \(L_2\) loss (which is
averaged across interactions), as it works on the eigenvalues of the
embedding, and therefore captures higher-level features of the network.
In panel \textbf{C}, we show positions of hosts and parasites on the
first two dimensions of the left and right subspaces. Note that these
values largely skew negative, because the first dimensions capture the
coarse structure of the network: most pairs of species do not interact,
and therefore have negative values. Finally in panel \textbf{D}, we show
the predicted weight (\emph{i.e.} the result of the multiplication of
the RDGP subspaces at a rank of 39) as a function of whether the
interactions are observed, not-observed, or unknown due to lack of
co-occurrence. This reveals that the observed interactions have higher
predicted weights, although there is some overlap; the usual approach to
identify potential interactions based on this information would be a
thresholding analysis, which is outside the scope of this manuscript
(and is done in the papers cited in this illustration). Because the
values returned from RDPG are not bound to the unit interval, we
performed a clamping of the weights to the unit space, showing a
one-inflation in documented interactions, and a zero-inflation in other
species pairs. This last figure crosses from the statistical into the
ecological, by showing that species pairs with no documented
co-occurrence have weights that are not distinguishable from species
pairs with no documented interactions, suggesting that (as befits a
host-parasite model) the ability to interact is a strong predictor of
co-occurrence.

\begin{figure}
\hypertarget{fig:illustration2}{%
\centering
\includegraphics{figures/illustration-part2.png}
\caption{Ecological analysis of an embedding for a host-parasite
metaweb, using Random Dot Product Graphs. \textbf{A}, relationship
between the number of parasites and position along the first axis of the
right-subspace for all hosts, showing that the embedding captures
elements of network structure at the species scale. \textbf{B}, weak
relationship between the body mass of hosts (in grams) and the position
alongside the same dimension. \textbf{C}, weak relationship between
bodymass of hosts and parasite richness. \textbf{D}, distribution of
positions alongside the same axis for hosts grouped by taxonomic
family.}\label{fig:illustration2}
}
\end{figure}

The results of fig.~\ref{fig:illustration1} show that we can extract an
embedding of the metaweb that captures enough variance to be relevant;
specifically, this is true both of \(L_2\) loss (indicating that RDPG is
able to capture pairwise processes) and the cumulative variance
explained (indicating that RDPG is able to capture network-level
structure). Therefore, in fig.~\ref{fig:illustration2}, we relate the
values of latent variables for hosts to different ecologically-relevant
data. In panel \textbf{A}, we show that host with a higher value on the
first dimension have fewer parasites. This relates to the body size of
hosts in the \emph{PanTHERIA} database (Jones et al., 2009), as shown in
panel \textbf{B}: interestingly, the position on the first axis is only
weakly correlated to body mass of the host; this matches well establihed
results showing that body size/mass is not always a direct predictor of
parasite richness in terrestrial mammals (Morand \& Poulin, 1998), a
result we observe in panel \textbf{C}. Finally, in panel \textbf{D}, we
can see how different taxonomic families occupy different positions on
the first axis, with \emph{e.g.} Sciuridae being biased towards higher
values. These results show how we can look for ecological informations
in the output of network embeddings, which can further be refined into
the selection of predictors for transfer learning.

\hypertarget{the-metaweb-embeds-both-ecological-hypotheses-and-practices}{%
\section{The metaweb embeds both ecological hypotheses and
practices}\label{the-metaweb-embeds-both-ecological-hypotheses-and-practices}}

The goal of metaweb inference is to provide information about the
interactions between species at a large spatial scale. But as Herbert
(1965) rightfully pointed out, ``{[}y{]}ou can't draw neat lines around
planet-wide problems''; any inference of a metaweb at large scales must
contend with several novel, and interwoven, families of problems. In
this section, we list some of the most pressing research priorities
(\emph{i.e.} problems that can be adressed with subsequent data analysis
or simulations), as well as issues related to the application of these
methods at the science-policy interface.

\hypertarget{identifying-the-properties-of-the-network-to-embed}{%
\subsection{Identifying the properties of the network to
embed}\label{identifying-the-properties-of-the-network-to-embed}}

If the initial metaweb is too narrow in scope, notably from a taxonomic
point of view, the chances of finding another area with enough related
species (through phylogenetic relatedness or similarity of functional
traits) to make a reliable inference decreases; this would likely be
indicated by large confidence intervals during estimation of the values
in the low-rank space, meaning that the representation of the original
graph is difficult to transfer to the new problem. Alternatively, if the
initial metaweb is too large (taxonomically), then the resulting
embeddings would need to represent interactions between taxonomic groups
that are not present in the new location. This would lead to a much
higher variance in the starting dataset, and to under-dispersion in the
target dataset, resulting in the potential under or over estimation of
the strength of new predicted interactions. The lack of well documented
metawebs is currently preventing the development of more concrete
guidelines. The question of phylogenetic relatedness and distribution is
notably relevant if the metaweb is assembled in an area with mostly
endemic species (\emph{e.g.} a system that has undergone recent
radiation or that has remained in isolation for a long period of time
might not have an analogous system with which to draw knowledge from),
and as with every predictive algorithm, there is room for the
application of our best ecological judgement. Because this problem
relates to distribution of species in the geographic or phylogenetic
space, it can certainly be approached through assessing the performance
of embedding transfer in simulated starting/target species pools.

\hypertarget{identifying-the-scope-of-the-prediction-to-perform}{%
\subsection{Identifying the scope of the prediction to
perform}\label{identifying-the-scope-of-the-prediction-to-perform}}

The area used to infer the new metaweb in determines the species pool
that must be used to perform this embedding. Metawebs can be constructed
by assigning interactions in a list of species within geographic
boundaries. The upside of this approach is that information at the
country level is likely to be required for biodiversity assessments, as
countries set conservation goals at the national level (Buxton et al.,
2021), and as quantitative instruments are designed to work at these
scales (\textbf{turak\_measuring\_2017?}); specific strategies are often
enacted at smaller scales, nested within a specific country (Ray et al.,
2021). But there is no guarantee that these boundaries are meaningful.
In fact, we do not have a satisfying answer to the question of ``where
does an ecological network stop?''. Recent results by Martins et al.
(2022) suggest that networks are shaped within eco-regions, with abrupt
structural transitions from an eco-region to the next. Should this trend
hold generally, this would provide an ecologically-relevant scale at
which metawebs can be downscaled and predicted. Other solutions leverage
network-area relationships to identify areas in which networks are
structurally similar (see \emph{e.g.} Fortin et al., 2021; Galiana et
al., 2022, 2018), which requires ample pre-existing information about
the network. This suggests that inferred metawebs should be further
downscaled to allow for \emph{a posteriori} analyses; Llewelyn et al.
(2022) provide compelling evidence for the fact that when a place is
rich in endemic species, even at smaller spatial scales, the transfer of
information about interactions becomes more challenging, re-emphasizing
the need to identify guidelines for when and where network predictions
can be mapped.

\hypertarget{minding-legacies-shaping-ecological-datasets}{%
\subsection{Minding legacies shaping ecological
datasets}\label{minding-legacies-shaping-ecological-datasets}}

Operating under the context of national divisions, in large parts of the
world, reflects nothing more than the legacy of settler colonialism,
which drives a disparity in available ecological data. Applying any
embedding to biased data does not debias them, but instead embeds these
very same biases, propagating them to the machine learning models using
embeddings to make predictions. Indeed, the use of ecological data
itself is not an apolitical act (Nost \& Goldstein, 2021), as data
infrastructures tend to be designed to answer questions within national
boundaries (therefore placing contingencies on what is available to be
embedded). Furthermore, their use often draws upon, and reinforces,
territorial statecraft (see \emph{e.g.} Barrett, 2005). As per Machen \&
Nost (2021), these biases are particularly important to consider when
the output of ``algorithmic thinking'' (\emph{i.e.} relying on machine
learning to generate knowledge or as a substitute to human
decision-making) can be re-purposed for governance (\emph{e.g.} enacting
conservation decisions on the basis of model prediction). As information
on species interaction networks structure is increasingly leveraged as a
tool to guide conservation actions (see \emph{e.g.} Eero et al., 2021;
Naman et al., 2022; Stier et al., 2017), the need to appraise and
correct biases that are unwittingly propagated to algorithms when
embedded from the original data is paramount. These considerations are
even more urgent in the specific context of biodiversity data, as
long-term colonial legacies still shape taxonomic composition to this
day (Lenzner et al., 2022; Raja, 2022), and where much shorter-term
changes in taxonomic and genetic richness of wildlife emerged through
environmental racism (Schmidt \& Garroway, 2022).

\hypertarget{conclusion-metaweb-prediction-in-context}{%
\section{Conclusion: metaweb prediction in
context}\label{conclusion-metaweb-prediction-in-context}}

Predictive approaches, regardless of the scale at which they are
deployed and the intent of their deployment, originate in the framework
that contributed to the ongoing biodiversity crisis (Adam, 2014) and
reinforced environmental injustice (Choudry, 2013; Domínguez \& Luoma,
2020). The risk of embedding this legacy in our models is great,
especially when the impact of this legacy on species pools is being
increasingly documented. This problem can be addressed by re-framing the
way we interact with models, especially in the context where these
models are intended to support conservation actions. Particularly on
territories that were traditionally stewarded by Indigenous people,
these approaches should be replaced (or at least guided and framed) by
Indigenous principles of land management (Eichhorn et al., 2019; No'kmaq
et al., 2021), as part of an ``algorithm-in-the-loop'' approach.
Human-algorithm interactions are notoriously difficult and can yield
adverse effect (Green \& Chen, 2019; Stevenson \& Doleac, 2021),
suggesting the need to systematically study them for the specific
purpose of biodiversity governance, as well as to improve the
algorithmic literacy of decision makers. As we see artificial
intelligence/machine learning being increasingly mobilized to generate
knowledge that is lacking for conservation decisions (\emph{e.g.} Lamba
et al., 2019; Mosebo Fernandes et al., 2020) and drive policy decisions
(Weiskopf et al., 2022), our discussion of these tools need to go beyond
the technical and statistical, and into the governance consequences they
can have.

\textbf{Acknowledgements:} We acknowledge that this study was conducted
on land within the traditional unceded territory of the Saint Lawrence
Iroquoian, Anishinabewaki, Mohawk, Huron-Wendat, and Omàmiwininiwak
nations. TP, TS, DC, and LP received funding from the Canadian Institute
for Ecology \& Evolution. FB is funded by the Institute for Data
Valorization (IVADO). TS, SB, and TP are funded by a donation from the
Courtois Foundation. CB was awarded a Mitacs Elevate Fellowship no.
IT12391, in partnership with fRI Research, and also acknowledges funding
from Alberta Innovates and the Forest Resources Improvement Association
of Alberta. M-JF acknowledges funding from NSERC Discovery Grant and
NSERC CRC. RR is funded by New Zealand's Biological Heritage Ngā Koiora
Tuku Iho National Science Challenge, administered by New Zealand
Ministry of Business, Innovation, and Employment. BM is funded by the
NSERC Alexander Graham Bell Canada Graduate Scholarship and the FRQNT
master's scholarship. LP acknowledges funding from NSERC Discovery Grant
(NSERC RGPIN-2019-05771). TP acknowledges financial support from the
Fondation Courtois, and NSERC through the Discovery Grants and Discovery
Accelerator Supplement programs. MJF is supported by an NSERC PDF and an
RBC Post-Doctoral Fellowship.

\textbf{Conflict of interest:} The authors have no conflict interests to
disclose

\textbf{Authors' contributions:} TS, and TP conceived the ideas
discussed in the manuscript. All authors contributed to writing and
editing the manuscript.

\textbf{Data availability:} There is no data associated with this
manuscript.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Adam2014Elephant}{}}%
Adam, R. (2014). \emph{Elephant treaties: The Colonial legacy of the
biodiversity crisis}. UPNE.

\leavevmode\vadjust pre{\hypertarget{ref-Albouy2019Marine}{}}%
Albouy, C., Archambault, P., Appeltans, W., Araújo, M. B., Beauchesne,
D., Cazelles, K., Cirtwill, A. R., Fortin, M.-J., Galiana, N., Leroux,
S. J., Pellissier, L., Poisot, T., Stouffer, D. B., Wood, S. A., \&
Gravel, D. (2019). The marine fish food web is globally connected.
\emph{Nature Ecology \& Evolution}, \emph{3}(8, 8), 1153--1161.
\url{https://doi.org/10.1038/s41559-019-0950-y}

\leavevmode\vadjust pre{\hypertarget{ref-Anowar2021Conceptual}{}}%
Anowar, F., Sadaoui, S., \& Selim, B. (2021). Conceptual and empirical
comparison of dimensionality reduction algorithms (PCA, KPCA, LDA, MDS,
SVD, LLE, ISOMAP, LE, ICA, t-SNE). \emph{Computer Science Review},
\emph{40}, 100378. \url{https://doi.org/10.1016/j.cosrev.2021.100378}

\leavevmode\vadjust pre{\hypertarget{ref-Athreya2017Statistical}{}}%
Athreya, A., Fishkind, D. E., Levin, K., Lyzinski, V., Park, Y., Qin,
Y., Sussman, D. L., Tang, M., Vogelstein, J. T., \& Priebe, C. E.
(2017). \emph{Statistical inference on random dot product graphs: A
survey} (No. arXiv:1709.05454). arXiv.
\url{http://arxiv.org/abs/1709.05454}

\leavevmode\vadjust pre{\hypertarget{ref-Barrett2005Environment}{}}%
Barrett, S. (2005). \emph{Environment and Statecraft: The Strategy of
Environmental Treaty-Making} (1st ed.). Oxford University PressOxford.
\url{https://doi.org/10.1093/0199286094.001.0001}

\leavevmode\vadjust pre{\hypertarget{ref-Bartomeus2013Understanding}{}}%
Bartomeus, I. (2013). Understanding linkage rules in plant-pollinator
networks by using hierarchical models that incorporate pollinator
detectability and plant traits. \emph{PloS One}, \emph{8}(7), e69200.
\url{http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0069200}

\leavevmode\vadjust pre{\hypertarget{ref-Bartomeus2016Common}{}}%
Bartomeus, I., Gravel, D., Tylianakis, J. M., Aizen, M. A., Dickie, I.
A., \& Bernard-Verdier, M. (2016). A common framework for identifying
linkage rules across different types of interactions. \emph{Functional
Ecology}, \emph{30}(12), 1894--1903.
\url{http://onlinelibrary.wiley.com/doi/10.1111/1365-2435.12666/full}

\leavevmode\vadjust pre{\hypertarget{ref-Bezanson2017Julia}{}}%
Bezanson, J., Edelman, A., Karpinski, S., \& Shah, V. B. (2017). Julia:
A Fresh Approach to Numerical Computing. \emph{SIAM Review},
\emph{59}(1), 65--98. \url{https://doi.org/10.1137/141000671}

\leavevmode\vadjust pre{\hypertarget{ref-Blanchet2020Cooccurrence}{}}%
Blanchet, F. G., Cazelles, K., \& Gravel, D. (2020). Co-occurrence is
not evidence of ecological interactions. \emph{Ecology Letters}.

\leavevmode\vadjust pre{\hypertarget{ref-Bohan2017Nextgeneration}{}}%
Bohan, D. A., Vacher, C., Tamaddoni-Nezhad, A., Raybould, A., Dumbrell,
A. J., \& Woodward, G. (2017). Next-Generation Global Biomonitoring:
Large-scale, Automated Reconstruction of Ecological Networks.
\emph{Trends in Ecology \& Evolution}.
\url{https://doi.org/10.1016/j.tree.2017.03.001}

\leavevmode\vadjust pre{\hypertarget{ref-Botella2022Appraisal}{}}%
Botella, C., Dray, S., Matias, C., Miele, V., \& Thuiller, W. (2022). An
appraisal of graph embeddings for comparing trophic network
architectures. \emph{Methods in Ecology and Evolution}, \emph{13}(1),
203--216. \url{https://doi.org/10.1111/2041-210X.13738}

\leavevmode\vadjust pre{\hypertarget{ref-Braga2019Spatial}{}}%
Braga, J., Pollock, L. J., Barros, C., Galiana, N., Montoya, J. M.,
Gravel, D., Maiorano, L., Montemaggiori, A., Ficetola, G. F., Dray, S.,
\& Thuiller, W. (2019). Spatial analyses of multi-trophic terrestrial
vertebrate assemblages in Europe. \emph{Global Ecology and
Biogeography}, \emph{28}(11), 1636--1648.
\url{https://doi.org/10.1111/geb.12981}

\leavevmode\vadjust pre{\hypertarget{ref-Braga2021Phylogenetic}{}}%
Braga, M. P., Janz, N., Nylin, S., Ronquist, F., \& Landis, M. J.
(2021). Phylogenetic reconstruction of ancestral ecological networks
through time for pierid butterflies and their host plants. \emph{Ecology
Letters}, \emph{n/a}(n/a). \url{https://doi.org/10.1111/ele.13842}

\leavevmode\vadjust pre{\hypertarget{ref-BramonMora2018Identifying}{}}%
Bramon Mora, B., Gravel, D., Gilarranz, L. J., Poisot, T., \& Stouffer,
D. B. (2018). Identifying a common backbone of interactions underlying
food webs from different ecosystems. \emph{Nature Communications},
\emph{9}(1), 2603. \url{https://doi.org/10.1038/s41467-018-05056-0}

\leavevmode\vadjust pre{\hypertarget{ref-Buxton2021Key}{}}%
Buxton, R. T., Bennett, J. R., Reid, A. J., Shulman, C., Cooke, S. J.,
Francis, C. M., Nyboer, E. A., Pritchard, G., Binley, A. D., Avery-Gomm,
S., Ban, N. C., Beazley, K. F., Bennett, E., Blight, L. K., Bortolotti,
L. E., Camfield, A. F., Gadallah, F., Jacob, A. L., Naujokaitis-Lewis,
I., \ldots{} Smith, P. A. (2021). Key information needs to move from
knowledge to action for biodiversity conservation in Canada.
\emph{Biological Conservation}, \emph{256}, 108983.
\url{https://doi.org/10.1016/j.biocon.2021.108983}

\leavevmode\vadjust pre{\hypertarget{ref-Catchen2023Missing}{}}%
Catchen, M., Poisot, T., Pollock, L., \& Gonzalez, A. (2023). \emph{The
missing link: Discerning true from false negatives when sampling species
interaction networks} (Preprint No. 4929). EcoEvoRXiV.
\url{https://doi.org/10.32942/X2DW22}

\leavevmode\vadjust pre{\hypertarget{ref-Cazelles2016Theory}{}}%
Cazelles, K., Araújo, M. B., Mouquet, N., \& Gravel, D. (2016). A theory
for species co-occurrence in interaction networks. \emph{Theoretical
Ecology}, \emph{9}(1), 39--48.
\url{https://doi.org/10.1007/s12080-015-0281-9}

\leavevmode\vadjust pre{\hypertarget{ref-Chami2022Machine}{}}%
Chami, I., Abu-El-Haija, S., Perozzi, B., Ré, C., \& Murphy, K. (2022).
Machine Learning on Graphs: A Model and Comprehensive Taxonomy.
\emph{Journal of Machine Learning Research}, \emph{23}(89), 1--64.
\url{http://jmlr.org/papers/v23/20-852.html}

\leavevmode\vadjust pre{\hypertarget{ref-Chen2017Deep}{}}%
Chen, D., Xue, Y., Fink, D., Chen, S., \& Gomes, C. P. (2017).
\emph{Deep Multi-species Embedding}. 3639--3646.
\url{https://www.ijcai.org/proceedings/2017/509}

\leavevmode\vadjust pre{\hypertarget{ref-Chen2017Harp}{}}%
Chen, H., Perozzi, B., Hu, Y., \& Skiena, S. (2017). \emph{HARP:
Hierarchical Representation Learning for Networks}.
\url{http://arxiv.org/abs/1706.07845}

\leavevmode\vadjust pre{\hypertarget{ref-Choudry2013Saving}{}}%
Choudry, A. (2013). Saving biodiversity, for whom and for what?
Conservation NGOs, complicity, colonialism and conquest in an era of
capitalist globalization. In \emph{NGOization: Complicity,
contradictions and prospects} (pp. 24--44). Bloomsbury Publishing.

\leavevmode\vadjust pre{\hypertarget{ref-Cieslak2020Tdistributed}{}}%
Cieslak, M. C., Castelfranco, A. M., Roncalli, V., Lenz, P. H., \&
Hartline, D. K. (2020). T-Distributed Stochastic Neighbor Embedding
(t-SNE): A tool for eco-physiological transcriptomic analysis.
\emph{Marine Genomics}, \emph{51}, 100723.
\url{https://doi.org/10.1016/j.margen.2019.100723}

\leavevmode\vadjust pre{\hypertarget{ref-Csermely2004Strong}{}}%
Csermely, P. (2004). Strong links are important, but weak links
stabilize them. \emph{Trends in Biochemical Sciences}, \emph{29}(7),
331--334. \url{https://doi.org/10.1016/j.tibs.2004.05.004}

\leavevmode\vadjust pre{\hypertarget{ref-DallaRiva2016Exploring}{}}%
Dalla Riva, G. V., \& Stouffer, D. B. (2016). Exploring the evolutionary
signature of food webs' backbones using functional traits. \emph{Oikos},
\emph{125}(4), 446--456. \url{https://doi.org/10.1111/oik.02305}

\leavevmode\vadjust pre{\hypertarget{ref-Dallas2017Predicting}{}}%
Dallas, T., Park, A. W., \& Drake, J. M. (2017). Predicting cryptic
links in host-parasite networks. \emph{PLOS Computational Biology},
\emph{13}(5), e1005557.
\url{https://doi.org/10.1371/journal.pcbi.1005557}

\leavevmode\vadjust pre{\hypertarget{ref-Danisch2021Makie}{}}%
Danisch, S., \& Krumbiegel, J. (2021). Makie.jl: Flexible
high-performance data visualization for Julia. \emph{Journal of Open
Source Software}, \emph{6}(65), 3349.
\url{https://doi.org/10.21105/joss.03349}

\leavevmode\vadjust pre{\hypertarget{ref-Dominguez2020Decolonising}{}}%
Domínguez, L., \& Luoma, C. (2020). Decolonising Conservation Policy:
How Colonial Land and Conservation Ideologies Persist and Perpetuate
Indigenous Injustices at the Expense of the Environment. \emph{Land},
\emph{9}(3, 3), 65. \url{https://doi.org/10.3390/land9030065}

\leavevmode\vadjust pre{\hypertarget{ref-Dunne2006Network}{}}%
Dunne, J. A. (2006). The Network Structure of Food Webs. In J. A. Dunne
\& M. Pascual (Eds.), \emph{Ecological networks: Linking structure and
dynamics} (pp. 27--86). Oxford University Press.

\leavevmode\vadjust pre{\hypertarget{ref-Eero2021Use}{}}%
Eero, M., Dierking, J., Humborg, C., Undeman, E., MacKenzie, B. R.,
Ojaveer, H., Salo, T., \& Köster, F. W. (2021). Use of food web
knowledge in environmental conservation and management of living
resources in the Baltic Sea. \emph{ICES Journal of Marine Science},
\emph{78}(8), 2645--2663. \url{https://doi.org/10.1093/icesjms/fsab145}

\leavevmode\vadjust pre{\hypertarget{ref-Eichhorn2019Steps}{}}%
Eichhorn, M. P., Baker, K., \& Griffiths, M. (2019). Steps towards
decolonising biogeography. \emph{Frontiers of Biogeography},
\emph{12}(1), 1--7. \url{https://doi.org/10.21425/F5FBG44795}

\leavevmode\vadjust pre{\hypertarget{ref-Eklof2013Dimensionality}{}}%
Eklöf, A., Jacob, U., Kopp, J., Bosch, J., Castro-Urgal, R., Chacoff, N.
P., Dalsgaard, B., de Sassi, C., Galetti, M., Guimarães, P. R.,
Lomáscolo, S. B., Martín González, A. M., Pizo, M. A., Rader, R.,
Rodrigo, A., Tylianakis, J. M., Vázquez, D. P., \& Allesina, S. (2013).
The dimensionality of ecological networks. \emph{Ecology Letters},
\emph{16}(5), 577--583. \url{https://doi.org/10.1111/ele.12081}

\leavevmode\vadjust pre{\hypertarget{ref-Eklof2016Phylogenetic}{}}%
Eklöf, A., \& Stouffer, D. B. (2016). The phylogenetic component of food
web structure and intervality. \emph{Theoretical Ecology}, \emph{9}(1),
107--115. \url{https://doi.org/10.1007/s12080-015-0273-9}

\leavevmode\vadjust pre{\hypertarget{ref-Fortin2021Network}{}}%
Fortin, M.-J., Dale, M. R. T., \& Brimacombe, C. (2021). Network ecology
in dynamic landscapes. \emph{Proceedings of the Royal Society B:
Biological Sciences}, \emph{288}(1949), rspb.2020.1889, 20201889.
\url{https://doi.org/10.1098/rspb.2020.1889}

\leavevmode\vadjust pre{\hypertarget{ref-Fricke2022Effects}{}}%
Fricke, E. C., Ordonez, A., Rogers, H. S., \& Svenning, J.-C. (2022).
The effects of defaunation on plants' capacity to track climate change.
\emph{Science}.
\url{https://www.science.org/doi/abs/10.1126/science.abk3510}

\leavevmode\vadjust pre{\hypertarget{ref-Galiana2022Ecological}{}}%
Galiana, N., Lurgi, M., Bastazini, V. A. G., Bosch, J., Cagnolo, L.,
Cazelles, K., Claramunt-López, B., Emer, C., Fortin, M.-J., Grass, I.,
Hernández-Castellano, C., Jauker, F., Leroux, S. J., McCann, K., McLeod,
A. M., Montoya, D., Mulder, C., Osorio-Canadas, S., Reverté, S.,
\ldots{} Montoya, J. M. (2022). Ecological network complexity scales
with area. \emph{Nature Ecology \& Evolution}, 1--8.
\url{https://doi.org/10.1038/s41559-021-01644-4}

\leavevmode\vadjust pre{\hypertarget{ref-Galiana2018Spatial}{}}%
Galiana, N., Lurgi, M., Claramunt-López, B., Fortin, M.-J., Leroux, S.,
Cazelles, K., Gravel, D., \& Montoya, J. M. (2018). The spatial scaling
of species interaction networks. \emph{Nature Ecology \& Evolution},
\emph{2}(5), 782--790. \url{https://doi.org/10.1038/s41559-018-0517-3}

\leavevmode\vadjust pre{\hypertarget{ref-Gaucher2021Outlier}{}}%
Gaucher, S., Klopp, O., \& Robin, G. (2021). Outlier detection in
networks with missing links. \emph{Computational Statistics \& Data
Analysis}, \emph{164}, 107308.
\url{https://doi.org/10.1016/j.csda.2021.107308}

\leavevmode\vadjust pre{\hypertarget{ref-Gibb2021Data}{}}%
Gibb, R., Albery, G. F., Becker, D. J., Brierley, L., Connor, R.,
Dallas, T. A., Eskew, E. A., Farrell, M. J., Rasmussen, A. L., Ryan, S.
J., Sweeny, A., Carlson, C. J., \& Poisot, T. (2021). Data
Proliferation, Reconciliation, and Synthesis in Viral Ecology.
\emph{BioScience}, \emph{71}(11), 1148--1156.
\url{https://doi.org/10.1093/biosci/biab080}

\leavevmode\vadjust pre{\hypertarget{ref-Goebel2023Body}{}}%
Goebel, L. G. A., Vitorino, B. D., Frota, A. V. B., \& Santos-Filho, M.
dos. (2023). Body mass determines the role of mammal species in a
frugivore-large fruit interaction network in a Neotropical savanna.
\emph{Journal of Tropical Ecology}, \emph{39}, e12.
\url{https://doi.org/10.1017/S0266467422000505}

\leavevmode\vadjust pre{\hypertarget{ref-Goodfellow2016Deep}{}}%
Goodfellow, I., Bengio, Y., \& Courville, A. (2016). \emph{Deep
learning}. MIT Press.

\leavevmode\vadjust pre{\hypertarget{ref-Goyal2018Graph}{}}%
Goyal, P., \& Ferrara, E. (2018). Graph embedding techniques,
applications, and performance: A survey. \emph{Knowledge-Based Systems},
\emph{151}, 78--94. \url{https://doi.org/10.1016/j.knosys.2018.03.022}

\leavevmode\vadjust pre{\hypertarget{ref-Gravel2018Bringing}{}}%
Gravel, D., Baiser, B., Dunne, J. A., Kopelke, J.-P., Martinez, N. D.,
Nyman, T., Poisot, T., Stouffer, D. B., Tylianakis, J. M., Wood, S. A.,
\& Roslin, T. (2018). Bringing Elton and Grinnell together: A
quantitative framework to represent the biogeography of ecological
interaction networks. \emph{Ecography}, \emph{0}(0).
\url{https://doi.org/10.1111/ecog.04006}

\leavevmode\vadjust pre{\hypertarget{ref-Gravel2013Inferring}{}}%
Gravel, D., Poisot, T., Albouy, C., Velez, L., \& Mouillot, D. (2013).
Inferring food web structure from predator-prey body size relationships.
\emph{Methods in Ecology and Evolution}, \emph{4}(11), 1083--1090.
\url{https://doi.org/10.1111/2041-210X.12103}

\leavevmode\vadjust pre{\hypertarget{ref-Green2019Disparate}{}}%
Green, B., \& Chen, Y. (2019). Disparate Interactions: An
Algorithm-in-the-Loop Analysis of Fairness in Risk Assessments.
\emph{Proceedings of the Conference on Fairness, Accountability, and
Transparency}, 90--99. \url{https://doi.org/10.1145/3287560.3287563}

\leavevmode\vadjust pre{\hypertarget{ref-Grover2016Node2vec}{}}%
Grover, A., \& Leskovec, J. (2016). Node2vec: Scalable Feature Learning
for Networks. \emph{Proceedings of the 22nd ACM SIGKDD International
Conference on Knowledge Discovery and Data Mining}, 855--864.
\url{https://doi.org/10.1145/2939672.2939754}

\leavevmode\vadjust pre{\hypertarget{ref-Grunig2020Crop}{}}%
Grünig, M., Mazzi, D., Calanca, P., Karger, D. N., \& Pellissier, L.
(2020). Crop and forest pest metawebs shift towards increased linkage
and suitability overlap under climate change. \emph{Communications
Biology}, \emph{3}(1, 1), 1--10.
\url{https://doi.org/10.1038/s42003-020-0962-9}

\leavevmode\vadjust pre{\hypertarget{ref-Gupta2022Simultaneously}{}}%
Gupta, A., Furrer, R., \& Petchey, O. L. (2022). Simultaneously
estimating food web connectance and structure with uncertainty.
\emph{Ecology and Evolution}, \emph{12}(3), e8643.
\url{https://doi.org/10.1002/ece3.8643}

\leavevmode\vadjust pre{\hypertarget{ref-Gupta2021Graph}{}}%
Gupta, A., Matta, P., \& Pant, B. (2021). Graph neural network: Current
state of Art, challenges and applications. \emph{Materials Today:
Proceedings}, \emph{46}, 10927--10932.
\url{https://doi.org/10.1016/j.matpr.2021.01.950}

\leavevmode\vadjust pre{\hypertarget{ref-Hadfield2014Tale}{}}%
Hadfield, J. D., Krasnov, B. R., Poulin, R., \& Nakagawa, S. (2014). A
Tale of Two Phylogenies: Comparative Analyses of Ecological
Interactions. \emph{The American Naturalist}, \emph{183}(2), 174--187.
\url{https://doi.org/10.1086/674445}

\leavevmode\vadjust pre{\hypertarget{ref-Herbert1965Dune}{}}%
Herbert, F. (1965). \emph{Dune} (1st ed.). Chilton Book Company.

\leavevmode\vadjust pre{\hypertarget{ref-Hinton2002Stochastic}{}}%
Hinton, G., \& Roweis, S. T. (2002). Stochastic neighbor embedding.
\emph{NIPS}, \emph{15}, 833--840.

\leavevmode\vadjust pre{\hypertarget{ref-Hoffmann2019Machine}{}}%
Hoffmann, J., Bar-Sinai, Y., Lee, L. M., Andrejevic, J., Mishra, S.,
Rubinstein, S. M., \& Rycroft, C. H. (2019). Machine learning in a
data-limited regime: Augmenting experiments with synthetic data uncovers
order in crumpled sheets. \emph{Science Advances}, \emph{5}(4),
eaau6792. \url{https://doi.org/10.1126/sciadv.aau6792}

\leavevmode\vadjust pre{\hypertarget{ref-Hortal2015Seven}{}}%
Hortal, J., de Bello, F., Diniz-Filho, J. A. F., Lewinsohn, T. M., Lobo,
J. M., \& Ladle, R. J. (2015). Seven Shortfalls that Beset Large-Scale
Knowledge of Biodiversity. \emph{Annual Review of Ecology, Evolution,
and Systematics}, \emph{46}(1), 523--549.
\url{https://doi.org/10.1146/annurev-ecolsys-112414-054400}

\leavevmode\vadjust pre{\hypertarget{ref-Jones2009Pantheria}{}}%
Jones, K. E., Bielby, J., Cardillo, M., Fritz, S. A., O'Dell, J., Orme,
C. D. L., Safi, K., Sechrest, W., Boakes, E. H., Carbone, C., Connolly,
C., Cutts, M. J., Foster, J. K., Grenyer, R., Habib, M., Plaster, C. A.,
Price, S. A., Rigby, E. A., Rist, J., \ldots{} Purvis, A. (2009).
PanTHERIA: A species‐level database of life history, ecology, and
geography of extant and recently extinct mammals: Ecological Archives
E090‐184. \emph{Ecology}, \emph{90}(9), 2648--2648.
\url{https://doi.org/10.1890/08-1494.1}

\leavevmode\vadjust pre{\hypertarget{ref-Jordano2016Sampling}{}}%
Jordano, P. (2016). Sampling networks of ecological interactions.
\emph{Functional Ecology}, \emph{30}(12), 1883--1893.
\url{https://doi.org/10.1111/1365-2435.12763}

\leavevmode\vadjust pre{\hypertarget{ref-Lamba2019Deep}{}}%
Lamba, A., Cassey, P., Segaran, R. R., \& Koh, L. P. (2019). Deep
learning for environmental conservation. \emph{Current Biology},
\emph{29}(19), R977--R982.
\url{https://doi.org/10.1016/j.cub.2019.08.016}

\leavevmode\vadjust pre{\hypertarget{ref-Lenzner2022Naturalized}{}}%
Lenzner, B., Latombe, G., Schertler, A., Seebens, H., Yang, Q., Winter,
M., Weigelt, P., van Kleunen, M., Pyšek, P., Pergl, J., Kreft, H.,
Dawson, W., Dullinger, S., \& Essl, F. (2022). Naturalized alien floras
still carry the legacy of European colonialism. \emph{Nature Ecology \&
Evolution}, 1--10. \url{https://doi.org/10.1038/s41559-022-01865-1}

\leavevmode\vadjust pre{\hypertarget{ref-Llewelyn2022Predicting}{}}%
Llewelyn, J., Strona, G., Dickman, C. R., Greenville, A. C., Wardle, G.
M., Lee, M. S. Y., Doherty, S., Shabani, F., Saltré, F., \& Bradshaw, C.
J. A. (2022). \emph{Predicting predator-prey interactions in terrestrial
endotherms using random forest} {[}Preprint{]}. Ecology.
\url{https://doi.org/10.1101/2022.09.02.506446}

\leavevmode\vadjust pre{\hypertarget{ref-Maaten2009Learning}{}}%
Maaten, L. van der. (2009). Learning a Parametric Embedding by
Preserving Local Structure. \emph{Proceedings of the Twelth
International Conference on Artificial Intelligence and Statistics},
384--391. \url{https://proceedings.mlr.press/v5/maaten09a.html}

\leavevmode\vadjust pre{\hypertarget{ref-Machen2021Thinking}{}}%
Machen, R., \& Nost, E. (2021). Thinking algorithmically: The making of
hegemonic knowledge in climate governance. \emph{Transactions of the
Institute of British Geographers}, \emph{46}(3), 555--569.
\url{https://doi.org/10.1111/tran.12441}

\leavevmode\vadjust pre{\hypertarget{ref-Malaterre2019Functional}{}}%
Malaterre, C., Dussault, A. C., Mermans, E., Barker, G., Beisner, B. E.,
Bouchard, F., Desjardins, E., Handa, I. T., Kembel, S. W., Lajoie, G.,
Maris, V., Munson, A. D., Odenbaugh, J., Poisot, T., Shapiro, B. J., \&
Suttle, C. A. (2019). Functional Diversity: An Epistemic Roadmap.
\emph{BioScience}, \emph{69}(10), 800--811.
\url{https://doi.org/10.1093/biosci/biz089}

\leavevmode\vadjust pre{\hypertarget{ref-Martins2022Global}{}}%
Martins, L. P., Stouffer, D. B., Blendinger, P. G., Böhning-Gaese, K.,
Buitrón-Jurado, G., Correia, M., Costa, J. M., Dehling, D. M., Donatti,
C. I., Emer, C., Galetti, M., Heleno, R., Jordano, P., Menezes, Í.,
Morante-Filho, J. C., Muñoz, M. C., Neuschulz, E. L., Pizo, M. A.,
Quitián, M., \ldots{} Tylianakis, J. M. (2022). Global and regional
ecological boundaries explain abrupt spatial discontinuities in avian
frugivory interactions. \emph{Nature Communications}, \emph{13}(1, 1),
6943. \url{https://doi.org/10.1038/s41467-022-34355-w}

\leavevmode\vadjust pre{\hypertarget{ref-McLeod2021Sampling}{}}%
McLeod, A., Leroux, S. J., Gravel, D., Chu, C., Cirtwill, A. R., Fortin,
M.-J., Galiana, N., Poisot, T., \& Wood, S. A. (2021). Sampling and
asymptotic network properties of spatial multi-trophic networks.
\emph{Oikos}, \emph{n/a}(n/a). \url{https://doi.org/10.1111/oik.08650}

\leavevmode\vadjust pre{\hypertarget{ref-Melnyk2020Graphkke}{}}%
Melnyk, K., Klus, S., Montavon, G., \& Conrad, T. O. F. (2020).
GraphKKE: Graph Kernel Koopman embedding for human microbiome analysis.
\emph{Applied Network Science}, \emph{5}(1), 96.
\url{https://doi.org/10.1007/s41109-020-00339-2}

\leavevmode\vadjust pre{\hypertarget{ref-Morales-Castilla2015Inferring}{}}%
Morales-Castilla, I., Matias, M. G., Gravel, D., \& Araújo, M. B.
(2015). Inferring biotic interactions from proxies. \emph{Trends in
Ecology \& Evolution}, \emph{30}(6), 347--356.
\url{https://doi.org/10.1016/j.tree.2015.03.014}

\leavevmode\vadjust pre{\hypertarget{ref-Morand1998Density}{}}%
Morand, S., \& Poulin, R. (1998). Density, body mass and parasite
species richness of terrestrial mammals. \emph{Evolutionary Ecology},
\emph{12}(6), 717--727. \url{https://doi.org/10.1023/A:1006537600093}

\leavevmode\vadjust pre{\hypertarget{ref-MoseboFernandes2020Machine}{}}%
Mosebo Fernandes, A. C., Quintero Gonzalez, R., Lenihan-Clarke, M. A.,
Leslie Trotter, E. F., \& Jokar Arsanjani, J. (2020). Machine Learning
for Conservation Planning in a Changing Climate. \emph{Sustainability},
\emph{12}(18, 18), 7657. \url{https://doi.org/10.3390/su12187657}

\leavevmode\vadjust pre{\hypertarget{ref-Murphy2022Probabilistic}{}}%
Murphy, K. P. (2022). \emph{Probabilistic machine learning: An
introduction}. MIT Press. \href{https://probml.ai}{probml.ai}

\leavevmode\vadjust pre{\hypertarget{ref-Naman2022Food}{}}%
Naman, S. M., White, S. M., Bellmore, J. R., McHugh, P. A., Kaylor, M.
J., Baxter, C. V., Danehy, R. J., Naiman, R. J., \& Puls, A. L. (2022).
Food web perspectives and methods for riverine fish conservation.
\emph{WIREs Water}, \emph{n/a}(n/a), e1590.
\url{https://doi.org/10.1002/wat2.1590}

\leavevmode\vadjust pre{\hypertarget{ref-Narayanan2017Graph2vec}{}}%
Narayanan, A., Chandramohan, M., Venkatesan, R., Chen, L., Liu, Y., \&
Jaiswal, S. (2017). \emph{Graph2vec: Learning Distributed
Representations of Graphs}. \url{http://arxiv.org/abs/1707.05005}

\leavevmode\vadjust pre{\hypertarget{ref-Neutel2002Stability}{}}%
Neutel, A.-M., Heesterbeek, J. A. P., \& de Ruiter, P. C. (2002).
Stability in Real Food Webs: Weak Links in Long Loops. \emph{Science},
\emph{296}(5570), 1120--1123.
\url{https://doi.org/10.1126/science.1068326}

\leavevmode\vadjust pre{\hypertarget{ref-Nokmaq2021Awakening}{}}%
No'kmaq, M., Marshall, A., Beazley, K. F., Hum, J., joudry, shalan,
Papadopoulos, A., Pictou, S., Rabesca, J., Young, L., \& Zurba, M.
(2021). {``Awakening the sleeping giant''}: Re-Indigenization principles
for transforming biodiversity conservation in Canada and beyond.
\emph{FACETS}, \emph{6}(1), 839--869.

\leavevmode\vadjust pre{\hypertarget{ref-Nost2021Political}{}}%
Nost, E., \& Goldstein, J. E. (2021). A political ecology of data.
\emph{Environment and Planning E: Nature and Space}, 25148486211043503.
\url{https://doi.org/10.1177/25148486211043503}

\leavevmode\vadjust pre{\hypertarget{ref-OConnor2020Unveiling}{}}%
O'Connor, L. M. J., Pollock, L. J., Braga, J., Ficetola, G. F.,
Maiorano, L., Martinez‐Almoyna, C., Montemaggiori, A., Ohlmann, M., \&
Thuiller, W. (2020). Unveiling the food webs of tetrapods across Europe
through the prism of the Eltonian niche. \emph{Journal of Biogeography},
\emph{47}(1), 181--192. \url{https://doi.org/10.1111/jbi.13773}

\leavevmode\vadjust pre{\hypertarget{ref-Pedersen2017Signatures}{}}%
Pedersen, E. J., Thompson, P. L., Ball, R. A., Fortin, M.-J., Gouhier,
T. C., Link, H., Moritz, C., Nenzen, H., Stanley, R. R. E., Taranu, Z.
E., Gonzalez, A., Guichard, F., \& Pepin, P. (2017). Signatures of the
collapse and incipient recovery of an overexploited marine ecosystem.
\emph{Royal Society Open Science}, \emph{4}(7), 170215.
\url{https://doi.org/10.1098/rsos.170215}

\leavevmode\vadjust pre{\hypertarget{ref-Perozzi2014Deepwalk}{}}%
Perozzi, B., Al-Rfou, R., \& Skiena, S. (2014). DeepWalk: Online
learning of social representations. \emph{Proceedings of the 20th ACM
SIGKDD International Conference on Knowledge Discovery and Data Mining},
701--710. \url{https://doi.org/10.1145/2623330.2623732}

\leavevmode\vadjust pre{\hypertarget{ref-Pichler2020Machine}{}}%
Pichler, M., Boreux, V., Klein, A.-M., Schleuning, M., \& Hartig, F.
(2020). Machine learning algorithms to infer trait-matching and predict
species interactions in ecological networks. \emph{Methods in Ecology
and Evolution}, \emph{11}(2), 281--293.
\url{https://doi.org/10.1111/2041-210X.13329}

\leavevmode\vadjust pre{\hypertarget{ref-Poisot2019Ecologicalnetworks}{}}%
Poisot, T., Belisle, Z., Hoebeke, L., Stock, M., \& Szefer, P. (2019).
EcologicalNetworks.jl - analysing ecological networks. \emph{Ecography}.
\url{https://doi.org/10.1111/ecog.04310}

\leavevmode\vadjust pre{\hypertarget{ref-Poisot2016Structure}{}}%
Poisot, T., Cirtwill, A. R., Cazelles, K., Gravel, D., Fortin, M.-J., \&
Stouffer, D. B. (2016). The structure of probabilistic networks.
\emph{Methods in Ecology and Evolution}, \emph{7}(3), 303--312.
\url{https://doi.org/10.1111/2041-210X.12468}

\leavevmode\vadjust pre{\hypertarget{ref-Poisot2021Imputing}{}}%
Poisot, T., Ouellet, M.-A., Mollentze, N., Farrell, M. J., Becker, D.
J., Albery, G. F., Gibb, R. J., Seifert, S. N., \& Carlson, C. J.
(2021). \emph{Imputing the mammalian virome with linear filtering and
singular value decomposition}. \url{http://arxiv.org/abs/2105.14973}

\leavevmode\vadjust pre{\hypertarget{ref-Poisot2015Species}{}}%
Poisot, T., Stouffer, D. B., \& Gravel, D. (2015). Beyond species: Why
ecological interaction networks vary through space and time.
\emph{Oikos}, \emph{124}(3), 243--251.
\url{https://doi.org/10.1111/oik.01719}

\leavevmode\vadjust pre{\hypertarget{ref-Raja2022Colonialism}{}}%
Raja, N. B. (2022). Colonialism shaped today's biodiversity.
\emph{Nature Ecology \& Evolution}, 1--2.
\url{https://doi.org/10.1038/s41559-022-01903-y}

\leavevmode\vadjust pre{\hypertarget{ref-Ramasamy2015Compressive}{}}%
Ramasamy, D., \& Madhow, U. (2015). Compressive spectral embedding:
Sidestepping the SVD. In C. Cortes, N. Lawrence, D. Lee, M. Sugiyama, \&
R. Garnett (Eds.), \emph{Advances in neural information processing
systems} (Vol. 28). Curran Associates, Inc.
\url{https://proceedings.neurips.cc/paper/2015/file/4f6ffe13a5d75b2d6a3923922b3922e5-Paper.pdf}

\leavevmode\vadjust pre{\hypertarget{ref-Ray2021Biodiversity}{}}%
Ray, J. C., Grimm, J., \& Olive, A. (2021). The biodiversity crisis in
Canada: Failures and challenges of federal and sub-national strategic
and legal frameworks. \emph{FACETS}, \emph{6}, 1044--1068.
\url{https://doi.org/10.1139/facets-2020-0075}

\leavevmode\vadjust pre{\hypertarget{ref-Runghen2021Exploiting}{}}%
Runghen, R., Stouffer, D. B., \& Dalla Riva, G. V. (2021).
\emph{Exploiting node metadata to predict interactions in large networks
using graph embedding and neural networks}.
\url{https://doi.org/10.1101/2021.06.10.447991}

\leavevmode\vadjust pre{\hypertarget{ref-Saravia2021Ecological}{}}%
Saravia, L. A., Marina, T. I., Kristensen, N. P., De Troch, M., \& Momo,
F. R. (2021). Ecological network assembly: How the regional metaweb
influences local food webs. \emph{Journal of Animal Ecology},
\emph{n/a}(n/a). \url{https://doi.org/10.1111/1365-2656.13652}

\leavevmode\vadjust pre{\hypertarget{ref-Schmidt2022Systemic}{}}%
Schmidt, C., \& Garroway, C. J. (2022). Systemic racism alters wildlife
genetic diversity. \emph{Proceedings of the National Academy of
Sciences}, \emph{119}(43), e2102860119.
\url{https://doi.org/10.1073/pnas.2102860119}

\leavevmode\vadjust pre{\hypertarget{ref-Stevenson2021Algorithmic}{}}%
Stevenson, M. T., \& Doleac, J. L. (2021). \emph{Algorithmic Risk
Assessment in the Hands of Humans} (SSRN Scholarly Paper No. 3489440).
\url{https://doi.org/10.2139/ssrn.3489440}

\leavevmode\vadjust pre{\hypertarget{ref-Stier2017Integrating}{}}%
Stier, A. C., Samhouri, J. F., Gray, S., Martone, R. G., Mach, M. E.,
Halpern, B. S., Kappel, C. V., Scarborough, C., \& Levin, P. S. (2017).
Integrating Expert Perceptions into Food Web Conservation and
Management. \emph{Conservation Letters}, \emph{10}(1), 67--76.
\url{https://doi.org/10.1111/conl.12245}

\leavevmode\vadjust pre{\hypertarget{ref-Stouffer2007Evidence}{}}%
Stouffer, D. B., Camacho, J., Jiang, W., \& Nunes Amaral, L. A. (2007).
Evidence for the existence of a robust pattern of prey selection in food
webs. \emph{Proceedings of the Royal Society B: Biological Sciences},
\emph{274}(1621), 1931--1940.
\url{https://doi.org/10.1098/rspb.2007.0571}

\leavevmode\vadjust pre{\hypertarget{ref-Stouffer2012Evolutionary}{}}%
Stouffer, D. B., Sales-Pardo, M., Sirer, M. I., \& Bascompte, J. (2012).
Evolutionary Conservation of Species' Roles in Food Webs.
\emph{Science}, \emph{335}(6075), 1489--1492.
\url{https://doi.org/10.1126/science.1216556}

\leavevmode\vadjust pre{\hypertarget{ref-Strydom2022Food}{}}%
Strydom, T., Bouskila, S., Banville, F., Barros, C., Caron, D., Farrell,
M. J., Fortin, M.-J., Hemming, V., Mercier, B., Pollock, L. J., Runghen,
R., Dalla Riva, G. V., \& Poisot, T. (2022). Food web reconstruction
through phylogenetic transfer of low-rank network representation.
\emph{Methods in Ecology and Evolution}, \emph{n/a}(n/a).
\url{https://doi.org/10.1111/2041-210X.13835}

\leavevmode\vadjust pre{\hypertarget{ref-Strydom2021Roadmap}{}}%
Strydom, T., Catchen, M. D., Banville, F., Caron, D., Dansereau, G.,
Desjardins-Proulx, P., Forero-Muñoz, N. R., Higino, G., Mercier, B.,
Gonzalez, A., Gravel, D., Pollock, L., \& Poisot, T. (2021). A roadmap
towards predicting species interaction networks (across space and time).
\emph{Philosophical Transactions of the Royal Society B: Biological
Sciences}, \emph{376}(1837), 20210063.
\url{https://doi.org/10.1098/rstb.2021.0063}

\leavevmode\vadjust pre{\hypertarget{ref-Strydom2021Svd}{}}%
Strydom, T., Dalla Riva, G. V., \& Poisot, T. (2021). SVD Entropy
Reveals the High Complexity of Ecological Networks. \emph{Frontiers in
Ecology and Evolution}, \emph{9}.
\url{https://doi.org/10.3389/fevo.2021.623141}

\leavevmode\vadjust pre{\hypertarget{ref-Surendran2013Graph}{}}%
Surendran, S. (2013). Graph Embedding and Dimensionality Reduction - A
Survey. \emph{International Journal of Computer Science \& Engineering
Technology}, \emph{4}(1).
\url{https://www.semanticscholar.org/paper/Graph-Embedding-and-Dimensionality-Reduction-A-Surendran/3f413d591e4b2b876e033eeb9390e232ad4826ca}

\leavevmode\vadjust pre{\hypertarget{ref-Tang2015Line}{}}%
Tang, J., Qu, M., Wang, M., Zhang, M., Yan, J., \& Mei, Q. (2015). LINE:
Large-scale Information Network Embedding. \emph{Proceedings of the 24th
International Conference on World Wide Web}, 1067--1077.
\url{https://doi.org/10.1145/2736277.2741093}

\leavevmode\vadjust pre{\hypertarget{ref-Thurman2019Testing}{}}%
Thurman, L. L., Barner, A. K., Garcia, T. S., \& Chestnut, T. (2019).
Testing the link between species interactions and co-occurrence in a
trophic network. \emph{Ecography}, \emph{0}.
\url{https://doi.org/10.1111/ecog.04360}

\leavevmode\vadjust pre{\hypertarget{ref-Torres2020Glee}{}}%
Torres, L., Chan, K. S., \& Eliassi-Rad, T. (2020). GLEE: Geometric
Laplacian Eigenmap Embedding. \emph{Journal of Complex Networks},
\emph{8}(2), cnaa007. \url{https://doi.org/10.1093/comnet/cnaa007}

\leavevmode\vadjust pre{\hypertarget{ref-Wang2016Structural}{}}%
Wang, D., Cui, P., \& Zhu, W. (2016). Structural Deep Network Embedding.
\emph{Proceedings of the 22nd ACM SIGKDD International Conference on
Knowledge Discovery and Data Mining}, 1225--1234.
\url{https://doi.org/10.1145/2939672.2939753}

\leavevmode\vadjust pre{\hypertarget{ref-Wang2021Joint}{}}%
Wang, S., Arroyo, J., Vogelstein, J. T., \& Priebe, C. E. (2021). Joint
Embedding of Graphs. \emph{IEEE Transactions on Pattern Analysis and
Machine Intelligence}, \emph{43}(4), 1324--1336.
\url{https://doi.org/10.1109/TPAMI.2019.2948619}

\leavevmode\vadjust pre{\hypertarget{ref-Wardeh2021Predicting}{}}%
Wardeh, M., Baylis, M., \& Blagrove, M. S. C. (2021). Predicting
mammalian hosts in which novel coronaviruses can be generated.
\emph{Nature Communications}, \emph{12}(1, 1), 780.
\url{https://doi.org/10.1038/s41467-021-21034-5}

\leavevmode\vadjust pre{\hypertarget{ref-Weiskopf2022Increasing}{}}%
Weiskopf, S. R., Harmáčková, Z. V., Johnson, C. G., Londoño-Murcia, M.
C., Miller, B. W., Myers, B. J. E., Pereira, L., Arce-Plata, M. I.,
Blanchard, J. L., Ferrier, S., Fulton, E. A., Harfoot, M., Isbell, F.,
Johnson, J. A., Mori, A. S., Weng, E., \& Rosa, I. M. D. (2022).
Increasing the uptake of ecological model results in policy decisions to
improve biodiversity outcomes. \emph{Environmental Modelling \&
Software}, \emph{149}, 105318.
\url{https://doi.org/10.1016/j.envsoft.2022.105318}

\leavevmode\vadjust pre{\hypertarget{ref-Williams2000Simple}{}}%
Williams, R. J., \& Martinez, N. D. (2000). Simple rules yield complex
food webs. \emph{Nature}, \emph{404}(6774), 180--183.
\url{https://doi.org/10.1038/35004572}

\leavevmode\vadjust pre{\hypertarget{ref-Wood2015Effects}{}}%
Wood, S. A., Russell, R., Hanson, D., Williams, R. J., \& Dunne, J. A.
(2015). Effects of spatial scale of sampling on food web structure.
\emph{Ecology and Evolution}, \emph{5}(17), 3769--3782.
\url{https://doi.org/10.1002/ece3.1640}

\leavevmode\vadjust pre{\hypertarget{ref-Wu2021Maximum}{}}%
Wu, D., Palmer, D. R., \& Deford, D. R. (2021). \emph{Maximum a
Posteriori Inference of Random Dot Product Graphs via Conic Programming}
(No. arXiv:2101.02180). arXiv. \url{http://arxiv.org/abs/2101.02180}

\leavevmode\vadjust pre{\hypertarget{ref-Xu2021Understanding}{}}%
Xu, M. (2021). Understanding Graph Embedding Methods and Their
Applications. \emph{SIAM Review}, \emph{63}(4), 825--853.
\url{https://doi.org/10.1137/20M1386062}

\leavevmode\vadjust pre{\hypertarget{ref-Yan2005Graph}{}}%
Yan, S., Xu, D., Zhang, B., \& Zhang, H.-J. (2005). Graph embedding: A
general framework for dimensionality reduction. \emph{2005 IEEE Computer
Society Conference on Computer Vision and Pattern Recognition
(CVPR'05)}, \emph{2}, 830--837 vol. 2.
\url{https://doi.org/10.1109/CVPR.2005.170}

\leavevmode\vadjust pre{\hypertarget{ref-Young2007Random}{}}%
Young, S. J., \& Scheinerman, E. R. (2007). Random Dot Product Graph
Models for Social Networks. In A. Bonato \& F. R. K. Chung (Eds.),
\emph{Algorithms and Models for the Web-Graph} (pp. 138--149). Springer.
\url{https://doi.org/10.1007/978-3-540-77004-6_11}

\leavevmode\vadjust pre{\hypertarget{ref-Zhou2020Graph}{}}%
Zhou, J., Cui, G., Hu, S., Zhang, Z., Yang, C., Liu, Z., Wang, L., Li,
C., \& Sun, M. (2020). Graph neural networks: A review of methods and
applications. \emph{AI Open}, \emph{1}, 57--81.
\url{https://doi.org/10.1016/j.aiopen.2021.01.001}

\end{CSLReferences}

\end{document}