### 2023-08-11 index.qmd
move to quarto

---
title: Graph embedding and transfer learning can help predict potential species interaction networks despite data limitations
authors:
  - name: Tanya Strydom
    affiliation:
        - ref: udem
        - ref: qcbs
    orcid: 0000-0001-6067-1349
    email: tanya.strydom@umontreal.ca
  - name: Salomé Bouskila
    affiliation:
        - ref: udem
    roles: writing
    corresponding: false
    orcid: 0000-0003-0193-5441
    email: salome.bouskila@umontreal.ca
  - name: Francis Banville
    affiliation:
        - ref: udem
        - ref: qcbs
        - ref: udes
    orcid: 0000-0001-9051-0597
    email: francis.banville@umontreal.ca
  - name: Ceres Barros
    affiliation:
        - Department of Forest Resources Management, University of British Columbia, Vancouver, B.C., Canada
    orcid: 0000-0003-4036-977X
    email: cbarros@mail.ubc.ca
affiliations:
    - id: udem
      name: Département de Sciences Biologiques, Université de Montréal, Montréal, Canada
    - id: qcbs
      name: Québec Centre for Biodiversity Science, Montréal, Canada
    - id: udes
      name: Département de Biologie, Université de Sherbrooke, Sherbrooke, Canada
bibliography: references.bib
---

The ability to infer *potential* interactions could serve as a significant
breakthrough in our ability to conceptualize species interaction networks over
large spatial scales [@Hortal2015Seven]. Reliable inferences would not only
boost our understanding of the structure of species interaction networks, but
also increase the amount of information that can be used for biodiversity
management. In a recent overview of the field of ecological network prediction,
@Strydom2021Roadmap identified two challenges of interest to the prediction of
interactions at large scales. First, there is a relative scarcity of relevant
data in most places globally -- which, due to the limitations in most predictive
methods, restricts the ability to infer interactions to locations where it is
least required (*i.e.* regions where we already have interaction data) leaving
us unable to make inference in data scarce regions (where we most need it);
second, accurate predictors are important for accurate predictions, and the lack
of methods that can leverage a small amount of *accurate* data is a serious
impediment to our predictive ability. In most places, our most reliable
biodiversity knowledge is that of a species pool where a set of potentially
interacting species in a given area could occur: through the analysis of
databases like the Global Biodiversity Information Facility (GBIF) or the
International Union for the Conservation of Nature (IUCN), it is possible to
construct a list of species for a region of interest; however inferring the
potential interactions between these species still remains a challenge.

Following the definition of @Dunne2006Network, a metaweb is the ecological
network analogue to the species pool; specifically, it inventories all
*potential* interactions between species for a spatially delimited area (and so
captures the $\gamma$ diversity of interactions). The metaweb itself is not a
prediction of local networks at specific locations within the spatial area it
covers: it will have a different structure, notably by having a larger
connectance [see *e.g.* @Wood2015Effects] and complexity [see *e.g.*
@Galiana2022Ecological], than any of these local networks. These local networks
(which capture the $\alpha$ diversity of interactions) are a subset of the
metaweb's species and its realized interactions, and have been called "metaweb
realizations" [@Poisot2015Species]. Differences between local networks and their
metawebs are due to chance, species abundance and co-occurrence, local
environmental conditions, and local distribution of functional traits, among
others. Specifically, although co-occurrence can be driven by interactions
[@Cazelles2016Theory], co-occurrence alone is not a predictor of interactions
[@Blanchet2020Cooccurrence; @Thurman2019Testing], and therefore the lack of
co-occurrence cannot be used to infer the lack of a feasible interaction. Yet,
recent results by @Saravia2021Ecological strongly suggested that local (metaweb)
realizations only respond weakly to local conditions: instead, they reflect
constraints inherited by the structure of their metaweb. This sets up the core
goal of predictive network ecology as the prediction of metaweb structure, as it
is required to accurately produce downscaled, local predictions.

Because the metaweb represents the joint effect of functional, phylogenetic, and
macroecological processes [@Morales-Castilla2015Inferring], it holds valuable
ecological information. Specifically, it represents the "upper bounds" on what
the composition of the local networks, given a local species pool, can be [see
*e.g.* @McLeod2021Sampling]; this information can help evaluate the ability of
ecological assemblages to withstand the effects of, for example, climate change
[@Fricke2022Effects]. These local networks may be reconstructed given an
appropriate knowledge of local species composition and provide information on
the structure of food webs at finer spatial scales. This has been done for
example for tree-galler-parasitoid systems [@Gravel2018Bringing], fish trophic
interactions [@Albouy2019Marine], tetrapod trophic interactions
[@Braga2019Spatial; @OConnor2020Unveiling], and crop-pest networks
[@Grunig2020Crop]. In this contribution, we highlight the power of viewing (and
constructing) metawebs as *probabilistic* objects in the context of
low-probability interactions, discuss how a family of machine learning tools
(graph embeddings and transfer learning) can be used to overcome data
limitations to metaweb inference, and highlight how the use of metawebs
introduces important questions for the field of network ecology.

# A metaweb is an inherently probabilistic object

Treating interactions as probabilistic (as opposed to binary) events is a more
nuanced and realistic way to represent them. @Dallas2017Predicting suggested
that most interactions (links) in ecological networks are cryptic, *i.e.*
uncommon or hard to observe. This argument echoes @Jordano2016Sampling: sampling
ecological interactions is difficult because it requires first the joint
observation of two species, and then the observation of their interaction. In
addition, it is generally expected that weak or rare interactions will be more
prevalent in networks than common or strong interactions [@Csermely2004Strong],
compared to strong, persistent interactions; this is notably the case in food
chains, wherein many weaker interactions are key to the stability of a system
[@Neutel2002Stability]. In the light of these observations, we expect to see an
over-representation of low-probability (hereafter rare) interactions under a
model that accurately predicts interaction probabilities.

Yet, the original metaweb definition, and indeed most past uses of
metawebs, was based on the presence/absence of interactions. Moving towards
*probabilistic* metawebs, by representing interactions as Bernoulli events [see
*e.g.* @Poisot2016Structure], offers the opportunity to weigh these rare
interactions appropriately. The inherent plasticity of interactions is important
to capture: there have been documented instances of food webs undergoing rapid
collapse/recovery cycles over short periods of time [*e.g.*
@Pedersen2017Signatures]. Furthermore, because the structure of the metaweb
cannot be known in advance, it is important to rely on predictive tools that do
not assume a specific network topology for link prediction
[@Gaucher2021Outlier], but are able to work on generalizations of the network.
These considerations emphasize why metaweb predictions should focus on
quantitative (preferentially probabilistic) predictions, and this should
constrain the suite of models that are appropriate for prediction.

It is important to recall that a metaweb is intended as a catalogue of all
potential (feasible) interactions, which is then filtered for a given
application [@Morales-Castilla2015Inferring]. It is therefore important to
separate the interactions that happen "almost surely" (repeated observational
data), "almost never" (repeated lack of evidence *or* evidence that the link is
forbidden through *e.g.* trait mis-match), and interactions with a probability
that lays somewhere in between [@Catchen2023Missing]. In a sense, that most
ecological interactions are elusive can call for a slightly different approach
to sampling: once the common interactions are documented, the effort required in
documenting each rare interaction will increase exponentially
[@Jordano2016Sampling]. Recent proposals in other fields relying on machine
learning approaches emphasize the idea that algorithms meant to predict, through
the assumption that they approximate the process generating the data, can also
act as data generators [@Hoffmann2019Machine]. High quality observational data
can be used to infer core rules underpinning network structure, and be
supplemented with synthetic data coming from predictive models trained on them,
thereby increasing the volume of information available for analysis. Indeed,
@Strydom2021Roadmap suggested that knowing the metaweb may render the prediction
of local networks easier, because it fixes an "upper bound" on which
interactions can exist. In this context, a probabilistic metaweb represents an
aggregation of informative priors on the biological feasibility of interactions,
which is usually hard to obtain yet has possibly the most potential to boost our
predictive ability of local networks [@Bartomeus2013Understanding;
@Bartomeus2016Common]. This would represent a departure from simple rules
expressed at the network scale [*e.g.* @Williams2000Simple] to a view of network
prediction based on learning the rules that underpin interactions *and* their
variability [@Gupta2022Simultaneously].

![The embedding process (**A**) can help to identify links (interactions) that
may have been missed within the original community (represented by the orange
dashed arrows, **B**). Transfer learning (**D**) allows for the prediction links
(interactions) even when novel species (**C**) are included alongside the
original community. This is achieved by learning using other relevant predictors
(*e.g.* traits) in conjunction with the known interactions to infer latent
values (**E**). Ultimately this allows us to predict links (interactions) for
species external from the original sample (blue dashed arrows) as well as
missing within sample links (**F**). Within this context the predicted (and
original) networks as well as the ecological predictors used (green boxes) are
products that can be quantified through measurements in the field, whereas the
embedded as well as imputed matrices (purple box) are representative of a
decomposition of the interaction matrices onto the embedding
space](figures/conceptual_2.png){#fig:embedding}

# Graph embedding offers promises for the inference of potential interactions

Graph (or network) embedding (@fig:embedding) is a family of machine learning
techniques, whose main task is to learn a mapping function from a discrete graph
to a continuous domain [@Arsov2019Network; @Chami2022Machine]. Their main goal
is to learn a low dimensional vector representation of the graph (embeddings),
such that its key properties (*e.g.* local or global structures) are retained in
the embedding space [@Yan2005Graph]. The embedding space may, but will not
necessarily, have lower dimensionality than the graph. Ecological networks are
promising candidates for the routine application of embeddings, as they tend to
possess a shared structural backbone [see *e.g.* @BramonMora2018Identifying],
which hints at structural invariants in empirical data. Assuming that these
structural invariants are common enough, they would dominate the structure of
networks, and therefore be adequately captured by the first (lower) dimensions
of an embedding, without the need to measure derived aspects of their structure
(*e.g.* motifs, paths, modularity, ...).

## Graph embedding produces latent variables (but not traits)

Before moving further, it is important to clarify the epistemic status of node
values derived from embeddings: specifically, they are *not* functional traits,
and therefore should not be interpreted in terms of effects or responses. As per
the framework of @Malaterre2019Functional, these values neither derive from, nor
result in, changes in organismal performance, and should therefore not be used
to quantify *e.g.* functional diversity. This holds true even when there are
correlations between latent values and functional traits: although these enable
an ecological discussion of how traits condition the structure of the network,
the existence of a statistical relationship does not elevate the latent values
to the status of functional traits.

Rather than directly predicting biological rules [see *e.g.* @Pichler2020Machine
for an overview], which may be confounded by the sparse nature of graph data,
learning embeddings works in the low-dimensional space that maximizes
information about the network structure. This approach is further justified by
the observation, for example, that the macro-evolutionary history of a network
is adequately represented by some graph embeddings [Random dot product graphs
(RDPG); see @DallaRiva2016Exploring]. In a recent publication, @Strydom2022Food
have used an embedding (based on RDPG) to project a metaweb of trophic
interactions between European mammals, and transferred this information to
mammals of Canada, using the phylogenetic distance between related clades to
infer the values in the latent subspace into which the European metaweb was
projected. By performing the RDPG step on re-constructed values, this approach
yields a probabilistic trophic metaweb for mammals of Canada based on knowledge
of European species, despite a limited ($\approx$ 5%) taxonomic overlap, and
illustrates how the values derived from an embedding can be used for prediction
without being "traits" of the species they represent.

## Ecological networks are good candidates for embedding

Food webs are inherently low-dimensional objects, and can be adequately
represented with less than ten dimensions [@Braga2021Phylogenetic;
@Eklof2013Dimensionality; @Braga2019Spatial]. Simulation results by
@Botella2022Appraisal suggested that there is no dominant method to identify
architectural similarities between networks: multiple approaches need to be
tested and compared to the network descriptor of interest on a problem-specific
basis. This matches previous results on graph embedding, wherein different
embedding algorithms yield different network embeddings [@Goyal2018Graph],
calling for a careful selection of the problem-specific approach to use. In
@tbl:methods, we present a selection of common graph and node embedding methods,
alongside examples of their use to predict interactions or statistical
associations between species. These methods rely largely on linear algebra or
pseudo-random walks on graphs. All forms of embeddings presented in @tbl:methods
share the common property of summarizing their objects into (sets of) dense
feature vectors, that capture the overall network structure, pairwise
information on nodes, and emergent aspects of the network, in a compressed way
(*i.e.* with some information loss, as we later discuss in the illustration).
Node embeddings tend to focus on maintaining pairwise relationships (*i.e.*
species interactions), while graph embeddings focus on maintaining the network
structure (*i.e.* emergent properties). Nevertheless, some graph embedding
techniques [like RDPG, see *e.g.* @Wu2021Maximum] will provide high-quality
node-level embeddings while also preserving network structure.

::: {#box-gnn .callout-note}
## Graph Neural Networks {#box-gnn}

One prominent family of approaches we do not discuss in the present manuscript
is Graph Neural Networks [GNN; @Zhou2020Graph]. GNN are, in a sense, a method to
embed a graph into a dense subspace, but belong to the family of deep learning
methods, which has its own set of practices [see *e.g.* @Goodfellow2016Deep]. An
important issue with methods based on deep learning is that, because their
parameter space is immense, the sample size of the data fed into them must be
similarly large (typically thousands of instances). This is a requirement for
the model to converge correctly during training, but this assumption is unlikely
to be met given the size of datasets currently available for metawebs (or single
time/location species interaction networks). This data volume requirement is
mostly absent from the techniques we list below. Furthermore, GNN still have
some challenges related to their shallow structure, and concerns related to
scalability [see @Gupta2021Graph for a review], which are mostly absent from the
methods listed in @tbl:methods. Assuming that the uptake of next-generation
biomonitoring techniques does indeed deliver larger datasets on species
interactions [@Bohan2017Nextgeneration], there is nevertheless the potential for
GNN to become an applicable embedding/predictive technique in the coming years.
:::

Graph embeddings *can* serve as a dimensionality reduction method. For example,
RDPG [@Strydom2022Food] and t-SVD [truncated Singular Value Decomposition;
@Poisot2021Imputing] typically embed networks using fewer dimensions than the
original network [the original network has as many dimensions as species, and as
many informative dimensions as trophically unique species; @Strydom2021Svd].
However, this is not necessarily the case -- indeed, one may perform a PCA (a
special case of SVD) to project the raw data into a subspace that improves the
efficacy of t-SNE [t-distributed stochastic neighbor embedding;
@Maaten2009Learning]. There are many dimensionality reductions
[@Anowar2021Conceptual] that can be applied to an embedded network should the
need for dimensionality reduction (for example for data visualization) arise. In
brief, many graph embeddings *can* serve as dimensionality reduction steps, but
not all do, neither do all dimensionality reduction methods provide adequate
graph embedding capacities. In the next section (and @fig:embedding), we show
how the amount of dimensionality reduction can affect the quality of the
embedding.

| Method        | Object          | Technique                        | Reference                | Application                                                                                                       |
| ------------- | --------------- | -------------------------------- | ------------------------ | ----------------------------------------------------------------------------------------------------------------- |
| tSNE          | nodes           | statistical divergence           | @Hinton2002Stochastic    | [@Cieslak2020Tdistributed, species-environment responses $^a$] [@Gibb2021Data, host-virus network representation] |
| LINE          | nodes           | stochastic gradient descent      | @Tang2015Line            |                                                                                                                   |
| SDNE          | nodes           | gradient descent                 | @Wang2016Structural      |                                                                                                                   |
| node2vec      | nodes           | stochastic gradient descent      | @Grover2016Node2vec      |                                                                                                                   |
| HARP          | nodes           | meta-strategy                    | @Chen2017Harp            |                                                                                                                   |
| DMSE          | joint nodes     | deep neural network              | @Chen2017Deep            | [@Chen2017Deep, species-environment interactions $^b$]                                                            |
| graph2vec     | sub-graph       | skipgram network                 | @Narayanan2017Graph2vec  |                                                                                                                   |
| RDPG          | graph           | SVD                              | @Young2007Random         | [@DallaRiva2016Exploring, trophic interactions] [@Poisot2021Imputing, host-virus network prediction]              |
| GLEE          | graph           | Laplacian eigenmap               | @Torres2020Glee          |                                                                                                                   |
| DeepWalk      | graph           | stochastic gradient descent      | @Perozzi2014Deepwalk     | [@Wardeh2021Predicting, host-virus interactions]                                                                  |
| GraphKKE      | graph           | stochastic differential equation | @Melnyk2020Graphkke      | [@Melnyk2020Graphkke, microbiome species associations $^a$]                                                       |
| FastEmbed     | graph           | eigen decomposition              | @Ramasamy2015Compressive |                                                                                                                   |
| PCA           | graph           | eigen decomposition              | @Surendran2013Graph      | [@Strydom2021Roadmap, host-parasite interactions]                                                                 |
| Joint methods | multiple graphs | multiple strategies              | @Wang2021Joint           |                                                                                                                   |

: Overview of some common graph embedding approaches, by type of embedded
objects, alongside examples of their use in the prediction of species
interactions. These methods have not yet been routinely used to predict species
interactions; most examples that we identified were either statistical
associations, or analogues to joint species distribution models. $^a$:
application is concerned with *statistical* interactions, which are not
necessarilly direct biotic interactions; $^b$:application is concerned with
joint-SDM-like approach, which is also very close to statistical associations as
opposed to direct biotic interactions. Given the need to evaluate different
methods on a problem-specific basis, the fact that a lot of methods have not
been used on network problems is an opportunity for benchmarking and method
development. Note that the row for PCA also applies to kernel/probabilistic PCA,
which are variations on the more general method of SVD. Note further that tSNE
has been included because it is frequently used to embed graphs, including of
species associations/interactions, despite not being strictly speaking, a graph
embedding technique [see *e.g.* @Chami2022Machine]. {#tbl:methods}

The popularity of graph embedding techniques in machine learning is more than
the search for structural invariants: graphs are discrete objects, and machine
learning techniques tend to handle continuous data better. Bringing a sparse
graph into a continuous, dense vector space [@Xu2021Understanding] opens up a
broader variety of predictive algorithms, notably of the sort that are able to
predict events as probabilities [@Murphy2022Probabilistic]. Furthermore, the
projection of the graph itself is a representation that can be learned;
@Runghen2021Exploiting, for example, used a neural network to learn the
embedding of a network in which not all interactions were known, based on the
nodes' metadata. This example has many parallels in ecology (see @fig:embedding
**C**), in which node metadata can be represented by phylogeny, abundance, or
functional traits. Using phylogeny as a source of information assumes (or
strives to capture) the action of evolutionary processes on network structure,
which at least for food webs have been well documented [@Braga2021Phylogenetic;
@DallaRiva2016Exploring; @Eklof2016Phylogenetic; @Stouffer2007Evidence;
@Stouffer2012Evolutionary]; similarly, the use of functional traits assumes that
interactions can be inferred from the knowledge of trait-matching rules, which
is similarly well supported in the empirical literature
[@Bartomeus2013Understanding; @Bartomeus2016Common; @Goebel2023Body;
@Gravel2013Inferring]. Relating this information to an embedding rather than a
list of network measures would allow to capture their effect on the more
fundamental aspects of network structure; conversely, the absence of a
phylogenetic or functional signal may suggest that evolutionary/trait processes
are not strong drivers of network structure, therefore opening a new way to
perform hypothesis testing.

# An illustration of metaweb embedding

In this section, we illustrate the embedding of a collection of bipartite
networks collected by @Hadfield2014Tale, using t-SVD and RDPG. Briefly, an RDPG
decomposes a network into two subspaces (left and right), which are matrices
that when multiplied give an approximation of the original network. RDPG has the
particularly desirable properties of being a graph embedding technique that
produces relevant node-level feature vectors, and provides good approximations
of graphs with varied structures [@Athreya2017Statistical]. The code to
reproduce this example is available as supplementary material [note, for the
sake of comparison, that @Strydom2021Roadmap have an example using embedding
through PCA followed by prediction using a deep neural network on the same
dataset]. The resulting (binary) metaweb $\mathcal{M}$ has 2131 interactions
between 206 parasites and 121 hosts, and its adjacency matrix has full rank
(*i.e.* it represents a space with 121 dimensions). All analyses were done using
Julia [@Bezanson2017Julia] version 1.7.2, *Makie.jl* [@Danisch2021Makie], and
*EcologicalNetworks.jl* [@Poisot2019Ecologicalnetworks].

![Validation of an embedding for a host-parasite metaweb, using Random Dot
Product Graphs. **A**, decrease in approximation error as the number of
dimensions in the subspaces increases. **B**, increase in cumulative variance
explained as the number of ranks considered increases; in **A** and **B**, the
dot represents the point of inflexion in the curve (at rank 39) estimated using
the finite differences method. **C**, position of hosts and parasites in the
space of latent variables on the first and second dimensions of their respective
subspaces (the results have been clamped to the unit interval). **D**, predicted
interaction weight from the RDPG based on the status of the species pair in the
metaweb.](figures/illustration-part1.png){#fig:illustration1}

In @fig:illustration1, we focus on some statistical checks of the embedding. In
panel **A**, we show that the averaged $L_2$ loss (*i.e.* the sum of squared
errors) between the empirical and reconstructed metaweb decreases when the
number of dimensions (rank) of the subspace increases, with an inflection at 39
dimensions (out of 120 initially) according to the finite differences method. As
discussed by @Runghen2021Exploiting, there is often a trade-off between the
number of dimensions to use (more dimensions are more computationally demanding)
and the quality of the representation. In panel **B**, we show the increase in
cumulative variance explained at each rank, and visualize that using 39 ranks
explains about 70% of the variance in the empirical metaweb. This is a different
information from the $L_2$ loss (which is averaged across interactions), as it
works on the eigenvalues of the embedding, and therefore captures higher-level
features of the network. In panel **C**, we show positions of hosts and
parasites on the first two dimensions of the left and right subspaces. Note that
these values largely skew negative, because the first dimensions capture the
coarse structure of the network: most pairs of species do not interact, and
therefore have negative values. Finally in panel **D**, we show the predicted
weight (*i.e.* the result of the multiplication of the RDGP subspaces at a rank
of 39) as a function of whether the interactions are observed, not-observed, or
unknown due to lack of co-occurrence in the original dataset. This reveals that
the observed interactions have higher predicted weights, although there is some
overlap; the usual approach to identify potential interactions based on this
information would be a thresholding analysis, which is outside the scope of this
manuscript (and is done in the papers cited in this illustration). Because the
values returned from RDPG are not bound to the unit interval, we performed a
clamping of the weights to the unit space, showing a one-inflation in documented
interactions, and a zero-inflation in other species pairs. This last figure
crosses from the statistical into the ecological, by showing that species pairs
with no documented co-occurrence have weights that are not distinguishable from
species pairs with no documented interactions, suggesting that (as befits a
host-parasite model) the ability to interact is a strong predictor of
co-occurrence.

![Ecological analysis of an embedding for a host-parasite metaweb, using Random
Dot Product Graphs. **A**, relationship between the number of parasites and
position along the first axis of the right-subspace for all hosts, showing that
the embedding captures elements of network structure at the species scale.
**B**, weak relationship between the body mass of hosts (in grams) and the
position alongside the same dimension. **C**, weak relationship between body
mass of hosts and parasite richness. **D**, distribution of positions alongside
the same axis for hosts grouped by taxonomic
family.](figures/illustration-part2.png){#fig:illustration2}

The results of @fig:illustration1 show that we can extract an embedding of the
metaweb that captures enough variance to be relevant; specifically, this is true
for both $L_2$ loss (indicating that RDPG is able to capture pairwise processes)
and the cumulative variance explained (indicating that RDPG is able to capture
network-level structure). Therefore, in @fig:illustration2, we relate the values
of latent variables for hosts to different ecologically-relevant data. In panel
**A**, we show that host with a higher value on the first dimension have fewer
parasites. This relates to the body size of hosts in the *PanTHERIA* database
[@Jones2009Pantheria], as shown in panel **B**: interestingly, the position on
the first axis is only weakly correlated to body mass of the host; this matches
well established results showing that body size/mass is not always a direct
predictor of parasite richness in terrestrial mammals [@Morand1998Density], a
result we observe in panel **C**. Finally, in panel **D**, we can see how
different taxonomic families occupy different positions on the first axis, with
*e.g.* Sciuridae being biased towards higher values. These results show how we
can look for ecological informations in the output of network embeddings, which
can further be refined into the selection of predictors for transfer learning.

# The metaweb merges ecological hypotheses and practices

Metaweb inference seeks to provide information about the interactions between
species at a large spatial scale, typically a scale large enough to be
considered of biogeographic relevance (indeed, many of the examples covered in
the introduction span areas larger than a country, some of them global). But as
@Herbert1965Dune rightfully pointed out, "[y]ou can't draw neat lines around
planet-wide problems"; any inference of a metaweb must therefore contend with
several novel, interwoven, families of problems. In this section, we outline
three that we think are particularly important, and can discuss how they may
addressed with subsequent data analysis or simulations, and how they emerge in
the specific context of using embeddings; some of these issues are related to
the application of these methods at the science-policy interface.

## Identifying the properties of the network to embed

If the initial metaweb is too narrow in scope, notably from a taxonomic point of
view, the chances of finding another area with enough related species (through
phylogenetic relatedness or similarity of functional traits) to make a reliable
inference decreases. This is because transfer requires similarity
(@fig:embedding). A diagnostic for the lack of similar species would likely be
large confidence intervals during estimation of the values in the low-rank
space. In other words, the representation of the original graph is difficult to
transfer to the new problem. Alternatively, if the initial metaweb is too large
(taxonomically), then the resulting embeddings would need to represent
interactions between taxonomic groups that are not present in the new location.
This would lead to a much higher variance in the starting dataset, and to
under-dispersion in the target dataset, resulting in the potential under or over
estimation of the strength of new predicted interactions.
@Llewelyn2022Predicting provided compelling evidence for these situations by
showing that, even at small spatial scales, the transfer of information about
interactions becomes more challenging when areas rich with endemic species are
considered. The lack of well documented metawebs is currently preventing the
development of more concrete guidelines. The question of phylogenetic
relatedness and distribution is notably relevant if the metaweb is assembled in
an area with mostly endemic species (*e.g.* a system that has undergone recent
radiation or that has remained in isolation for a long period of time might not
have an analogous system with which to draw knowledge from), and as with every
predictive algorithm, there is room for the application of our best ecological
judgement. Because this problem relates to distribution of species in the
geographic or phylogenetic space, it can certainly be approached through
assessing the performance of embedding transfer in simulated starting/target
species pools.

## Identifying the scope of the prediction to perform

The area for which we seek to predict the metaweb should determine the species
pool on which the embedding is performed. Metawebs can be constructed by
assigning interactions in a list of species within specific regions. The upside
of this approach is that information relevant for the construction of this
dataset is likely to exist, as countries usually set conservation goals at the
national level [@Buxton2021Key], and as quantitative instruments are
consequently designed to work at these scales [@Turak2017Using]; specific
strategies are often enacted at smaller scales, nested within a specific country
[@Ray2021Biodiversity]. However, there is no guarantee that these arbitrary
boundaries are meaningful. In fact, we do not have a satisfying answer to the
question of "where does an ecological network stop?", the answer to which would
dictate the spatial span to embed/predict. Recent results by @Martins2022Global
suggested that networks are shaped within eco-regions, with abrupt structural
transitions from an eco-region to the next. Should this trend hold generally,
this would provide an ecologically-relevant scale at which metawebs can be
downscaled and predicted. Other solutions could leverage network-area
relationships to identify areas in which networks are structurally similar [see
*e.g.* @Fortin2021Network; @Galiana2018Spatial; @Galiana2022Ecological]. Both of
these solutions require ample pre-existing information about the network in
space. Nevertheless, the inclusion of species for which we have data but that
are not in the right spatial extent *may* improve the performance of approaches
based on embedding and transfer, *if* they increase the similarity between the
target and destination network. This proposal can specifically be evaluated by
adding nodes to the network to embed, and assessing the performance of
predictive models [see *e.g.* @Llewelyn2022Predicting].

# Conclusion: metawebs, predictions, and people

Predictive approaches in ecology, regardless of the scale at which they are
deployed and the intent of their deployment, originate in the framework that
contributed to the ongoing biodiversity crisis [@Adam2014Elephant] and
reinforced environmental injustice [@Choudry2013Saving;
@Dominguez2020Decolonising]. The risk of embedding this legacy in our models is
real, especially when the impact of this legacy on species pools is being
increasingly documented. This problem can be addressed by re-framing the way we
interact with models, especially when models are intended to support
conservation actions. Particularly on territories that were traditionally
stewarded by Indigenous people, we must interrogate how predictive approaches
and the biases that underpin them can be put to task in accompanying Indigenous
principles of land management [@Eichhorn2019Steps; @Nokmaq2021Awakening]. The
discussion of "algorithm-in-the-loop" approaches that is now pervasive in the
machine learning community provides examples of why this is important.
Human-algorithm interactions are notoriously difficult and can yield adverse
effects [@Green2019Disparate; @Stevenson2021Algorithmic], suggesting the need to
systematically study them for the specific purpose of, here, biodiversity
governance. Improving the algorithmic literacy of decision makers is part of the
solution [*e.g.* @Lamba2019Deep; @MoseboFernandes2020Machine], as we can
reasonably expect that model outputs will be increasingly used to drive policy
decisions [@Weiskopf2022Increasing]. Our discussion of these approaches need to
go beyond the technical and statistical, and into the governance consequences
they can have. To embed data also embeds historical and contemporary biases that
acted on these data, both because they shaped the ecological processes
generating them, and the global processes leading to their measurement and
publication. For a domain as vast as species interaction networks, these biases
exist at multiple scales along the way, and a challenge for prediction is not
only to develop (or adopt) new quantitative tools, but to assess the behavior of
these tools in the proper context.

::: {#box-legacy .callout-note}
## Minding legacies shaping ecological datasets {#box-legacy}

In large parts of the world, boundaries that delineate geographic regions are
merely a reflection the legacy of settler colonialism, which drives global
disparity in capacity to collect and publish ecological data. Applying any
embedding to biased data does not debias them, but rather embeds these biases,
propagating them to the models using embeddings to make predictions.
Furthermore, the use of ecological data itself is not an apolitical act
[@Nost2021Political]: data infrastructures tend to be designed to answer
questions within national boundaries (therefore placing contingencies on what is
available to be embedded), their use often drawing upon, and reinforcing,
territorial statecraft [see *e.g.* @Barrett2005Environment]. As per
@Machen2021Thinking, these biases are particularly important to consider when
knowledge generated algorithmically is used to supplement or replace human
decision-making, especially for governance (*e.g.* enacting conservation
decisions on the basis of model prediction). As information on networks is
increasingly leveraged for conservation actions [see *e.g.* @Eero2021Use;
@Naman2022Food; @Stier2017Integrating], the need to appraise and correct biases
that are unwittingly propagated to algorithms when embedded from the original
data is immense. These considerations are even more urgent in the specific
context of biodiversity data. Long-term colonial legacies still shape taxonomic
composition to this day [@Lenzner2022Naturalized; @Raja2022Colonialism], and
much shorter-term changes in taxonomic and genetic richness of wildlife emerged
through environmental racism [@Schmidt2022Systemic]. Thus, the set of species
found at a specific location is not only as the result of a response to
ecological processes separate from human influence, but also the result of
human-environment interaction as well as the result legislative/political
histories.
:::

**Acknowledgements:** We acknowledge that this study was conducted on land
within the traditional unceded territory of the Saint Lawrence Iroquoian,
Anishinabewaki, Mohawk, Huron-Wendat, and Omàmiwininiwak nations. TP, TS, DC,
and LP received funding from the Canadian Institute for Ecology & Evolution. FB
is funded by the Institute for Data Valorization (IVADO). TS, SB, and TP are
funded by a donation from the Courtois Foundation. CB was awarded a Mitacs
Elevate Fellowship no. IT12391, in partnership with fRI Research, and also
acknowledges funding from Alberta Innovates and the Forest Resources Improvement
Association of Alberta. M-JF acknowledges funding from NSERC Discovery Grant and
NSERC CRC. RR is funded by New Zealand’s Biological Heritage Ngā Koiora Tuku Iho
National Science Challenge, administered by New Zealand Ministry of Business,
Innovation, and Employment. BM is funded by the NSERC Alexander Graham Bell
Canada Graduate Scholarship and the FRQNT master’s scholarship. LP acknowledges
funding from NSERC Discovery Grant (NSERC RGPIN-2019-05771). TP acknowledges
financial support from the Fondation Courtois, and NSERC through the Discovery
Grants and Discovery Accelerator Supplement programs. MJF is supported by an
NSERC PDF and an RBC Post-Doctoral Fellowship.

**Conflict of interest:** The authors have no conflict of interests to disclose

**Authors' contributions:** TS, and TP conceived the ideas discussed in the
manuscript. All authors contributed to writing and editing the manuscript.

**Data availability:** There is no data associated with this manuscript.

# References